---
title: "Statistical analysis: The Effects of Iron Overload on Hemptotic Stemcell Transplant (HSCT) Outcomes in Sickle Cell Patients"
author: "Nhat Nguyen \n PEDSnet Coordinating Center"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    page-layout: full
    embed-resources: true
    smooth-scroll: true
    fig-cap: TRUE
    df-print: paged
    toc: TRUE
    toc-depth: 5
    toc-collapsed: TRUE
    toc-float: TRUE
    toc-location: left
    number-sections: TRUE
    number-depth: 3
    output-file: Gibson_sickle_cell_statisical_analysis
    output-ext: "html"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 6, fig.align = "center")
# Default taken from R notebook behavior: when knitting, wd will always be location of notebook
base_dir <- '..'
Sys.setenv(PEDSNET_SKIP_REQUEST_EXECUTION=1)
tryCatch(source('../site/run.R'),
         error = function (e) message(e)) # May not be able to make db connection

# Set to "local" to read data from ../results, or another value to read from db
data_source <- "not_local" #if_else(config('execution_mode') == 'distribution', 'local', 'not_local')

require(tidyr)
require(knitr)
require(kableExtra)
require(stringr)
require(tibble)
require(ggplot2)
require(table1)
require(tidyverse)
library(survival)
library(survminer)
library(ggsurvfit)
library(gtsummary)
library(cowplot)
```

```{r funtioncs, include=FALSE}
get_codesets <- function(codeset_name) {
  read_csv(paste0('../specs/', codeset_name, '.csv'), col_types = c('c', 'c', 'c', 'c'), show_col_types = FALSE) %>%
  mutate(concept_id = as.character((concept_id)), concept_code = as.character((concept_code))) %>% collect()
}

get_results <- function(tbl_name) {
  if (data_source == 'local') {
    rslt <- read.csv(paste0('../local/', tbl_name, '.csv'))
  }
  else {
    rslt <- results_tbl(tbl_name) %>% collect()
  }
  rslt
}

prettify_kable <- function(data) {
  rslt <- data %>% kable(digits = 4, format.args = list(big.mark = ','))
  if (knitr::is_html_output()) rslt <- rslt %>%
      kable_styling(bootstrap_options = c("striped", "hover")) %>%
      column_spec(1, bold = T, border_right = T)
  rslt
}

my.render.cont <- function(x) {
    with(stats.apply.rounding(stats.default(x), digits=2), c("",
        "Median [Min, Max]"=sprintf("%s [%s, %s])", MEDIAN, MIN, MAX)))
}

cox_sum <- function(cox_sum){
  cox_sum <- summary(cox_sum)
  res <- cbind.data.frame(cox_sum$conf.int, cox_sum$coefficients)

  res <- res[c("exp(coef)", "lower .95", "upper .95", "Pr(>|z|)")] %>% 
              rename(HR = "exp(coef)", lower = "lower .95", upper = "upper .95", p_val = "Pr(>|z|)") %>%
              mutate(`95 CI` = paste0("(",round(lower, 2), " - ", round(upper, 2), ")")) %>%
              mutate(p_val = case_when(p_val <= 0.0001~ paste0(round(p_val, 5), "***"), 
                                      p_val <= 0.001~ paste0(round(p_val, 5), "**"), 
                                      p_val <= 0.01~ paste0(round(p_val, 5), "*"), 
                                      p_val <= 0.05~ paste0(round(p_val, 5), "."), 
                                      TRUE ~ as.character(round(p_val, 5)))) %>%
              mutate(HR = round(HR, 2)) %>%
              select("HR", "95 CI", "p_val")
  return(res)
}

cox_fit_plot <- function(data, time_var, status_var, covariates, iron_level_var, plot_labels, sub_group){
  data <- data %>% filter(scd_type == sub_group)

  # Create the formula 
  formula_str <- paste("Surv(", time_var, "/365.25,", status_var, ") ~", paste(covariates, collapse = " + "))
  formula <- as.formula(formula_str)

  # Fit the Cox proportional hazards model
  cox_model <- coxph(formula, data = data)

  iron_levels <- data %>% distinct(!!sym(iron_level_var)) 

  # visualiza the impact on iron levels
  iron_levels_df <- with(data,
               data.frame(iron_levels, 
                          age_at_ce = rep(mean(age_at_ce, na.rm = TRUE), length(iron_levels)),
                          gender = rep("Female", length(iron_levels)),
                          transplant_type_combined = rep("matched related",length(iron_levels)),
                          has_busulfan = rep("No", length(iron_levels))
                          )
               )

  # print(iron_levels_df)
  # iron_levels_df <- iron_levels_df %>% rename(iron_level_var = test)
  
  plot_fit <- survfit(cox_model, newdata = iron_levels_df)
  p <- ggsurvplot(plot_fit, data = iron_levels_df, 
                  conf.int = FALSE, ggtheme = theme_minimal(),
                  legend.labs = iron_levels %>% pull(), #c("Low", "High"),
                  legend.title = plot_labels[1], #"Pre-transplant LIC levels", 
                  xlab = "Years", ylab = paste(plot_labels[2], plot_labels[1], sub_group, sep = " "), #, #"Overall survival probability",
                  #xlim = c(0, 5), ylim = c(0.5, 1.0)
                  ) 
  return(list(fit = cox_model, p = p))
}
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
tribble( ~Item, ~Value,
         'Title', config('title'),
         'PI', config('requester_name'),
         'CDM Version', as.character(config('cdm_version'))) %>%
  prettify_kable()

```
```{r data, include= FALSE}
cohort <- results_tbl("study_cohorts") %>%
                    pivot_longer(cols = aim_2a_2:aim_3_2, names_to = "aim", values_to = "which_aim") %>% 
                    filter(which_aim == TRUE) %>% 
                    collect() %>%
                    mutate(age_at_ce = round(as.numeric(difftime(transplant_date, birth_date, units = "days"))/365.25)) %>%
                    # mutate(transplant_type = case_when(grepl("allogeneic", transplant_concept_name, ignore.case = TRUE) ~ "allogeneic",
                    #                           grepl("autologous", transplant_concept_name, ignore.case = TRUE) ~ "autologous",
                    #                           TRUE ~ "unknown")) %>% 
                    select(person_id, aim, site, age_at_ce , gender, scd_concept_name, scd_type, transplant_concept_name, transplant_type, transplant_date)
cohort_ct <- cohort %>% distinct_ct()
```
# Data analysis plan:
- This is a retrospective cohort study comparing outcomes between patients with pre/post-HSCT iron overload to patients without pre/post-HSCT iron overload. Analyze the impact of pre/post-HSCT elevated ferritin and LIC on short-term hematologic and transplant-related outcome: 
  - overall survival
  - graft-versus-host disease (GvHD)
  - graft failure
  - incidence of veno-occlusive disease (VOD)
  - bacteremia
  - immune reconstitution
  - time to neutrophil and platelet engraftment. 

## Aim_2a: 
- GRFS (Graft failure free, GVHD-free survival survival): Time-to-event analysis with Cox proportional hazards model, Kaplan-Meier curves with log-rank test 
- GVHD: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test.  
- Disease relapse: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test.  
- VOD: Multivariable logistic regression with stratification for busulfan conditioning, use interaction term for busulfan and iron score 
- Time to neutrophil and platelet engraftment: time-to-event analysis using a competing risks analysis. Competing risks include death or graft failure prior to engraftment. 
- Bacteremia within 1 year after HSCT: Time-to-event analysis using a competing risks analysis. Competing risk includes death. 
- Immune Reconstitution: time-to-event analysis using a competing risks analysis. Competing risks include death and graft failure requiring a second transplant. 

## Aim_2b: 
- GRFS: Time-to-event analysis with Cox proportional hazards model, Kaplan-Meier curves with log-rank test 
- Overall survival: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test.  
- Disease relapse: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test.  
- GVHD: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test. 

# Background:
- Armand et al 2007:
  - n = 590, follow-up times = 8yrs since transplant
  - Disease: acute leukemia, MDS, chronic myelogenous leukemia CML, non-Hodgkin lymphoma NHL. 
  - Ferritin_threshold = 2034 ng/mL: significant negative association between ferritin levels and OS and disease-free survival (DFS) for MDS and acute leukemia. For MDS, an independent significant association between higher ferritin levels and increased treatment-related mortality (TRM). No significant effects observed for CML or NHL
- Armand et al 2012: 
  - n = 45
  - Disease: acute leukemia (acute myelogenous leukemia (AML), acute lymphoblastic leukemia (ALL)) and myelodyplastic syndrome (MDS)
  - LIC_threshold = 5 (g/gdw): higher LIC level is associated with higher 1-yr OS and lower 2-yr OS but not significant. Using threshold of 3 or 7 did not change the results. 
  - Ferrintin_threshold = 2000 ng/ml: higher baseline ferritin level is significantly associated with lower OS at both 1-yr and 2-yr post-transplant. However, neither the association between baseline ferritin and cummulative incidence of relapse (CIR) nor non-relapse mortality (NMR) is significant. 
  - The authors suggested that ferritin rather than LIC is prognostic in this case not because it measures iron stores more accurately but because it's an acute phase reactants (such as albumin or C-reactie protein) which reflects inflammatory status from comorbid conditions, advance diseases, etc..  
- Armand et al 2014:
  - Meta-analysis from 4 published cohorts n = 274 with AML and MDS
  - They looked at the correlation between SF and LIC, and SF of 2500 ng/mL corresponds to an LIC around 6mg/gdw
  - 2 thresholds for each variables LIC 5 & 7 g/gdw, and SF 1000 and 2500 ng/mL. 
  - No siginificant association between SF or LIC and OS or TRM observed for thresholds of 7 and 1000, respectively in all 4 cohorts. 
  - In the pooled cohort, no siginificant association between LIC and OS or TRM observed for LIC threshold of 7. The association between SF and OS is only significant at 1000ng/mL but not at 2500ng/mL. The authors suggested that the drop in OS was not atributted to the elevated iron level but rather the underlying disease that causes the iron level elevation. If this were true, a threshold of 1000ng/mL for SF is a better choice since it allows us to better capture the individuals at risk. 
- Lee et al 2009:
  - Retrospective study, n = 101
  - Diseases: ALL, AML, MDS, aplastic anemia AA. 
  - SF_threshold = 1000 ng/mL: significant association between elevated SF and OS, DFS, and TMR. 
  - Had a third group: patients whose SF dropped to below 1000ng/mL after iron-chelating therapy (IC group)
- Wermke et al 2012:
  - n = 88
  - Diseases: AML and MDS
  - Significant correlation between SF and LIC (r = 0.782). 
  - SF_threshold = 2500 ng/mL and LIC_threshold = 125ng/mL: the results are very interesting: 
    - Elevated LIC but not ferritin is significantly associated with lower OS and NRM. 
- Important covariates:
  - History of transplant: we would not have this info if the patients changed their care sites. However, chart reviewers should have caught this? Is it in the cohort eligibility? 

# Overall survival

- match status: Haploidentical, matched related, matched unrelated, mismatched related and mismatched unrelated:
    - matched related: match_status = 8/8 or 10/10 AND donor_relation = "Related"
    - matched unrelated: match_status = 8/8 or 10/10 AND donor_relation = "Unrelated"
    - mismatched related: match_status = 8/10, 9/10, or 7/8 AND donor_relation = "Related"
    - mismatched unrelated: match_status = 8/10, 9/10, or 7/8 AND donor_relation = "Unrelated"
    - haploididentical (related)
    - autologous/genetherapy
    - Allogeneic: Cord Blood

## Aim 2a-2: Overall survival stratified by <span style="color:red">pre-transplant ferritin levels</span>
```{r cohort_covars_survival_analysis, echo= FALSE, message=FALSE}
stat_dataset <- results_tbl("survival_analysis_dataset") %>% collect() %>%
                  mutate(transplant_type_combined = relevel(factor(transplant_type_combined), ref = "matched related"))
                  # , levels = c("matched related", "matched unrelated", "mismatched related", "mismatched unrelated", "Haploidentical", "Allogeneic: Cord Blood", "autologous/genetherapy", "Unknown")))

ferritin_data <- results_tbl("cr_ferritin_data") %>%
                    select(-no_ferritin_pre, - no_ferritin_post) %>% collect()

stat_dataset_ferritin <-  stat_dataset %>%
            left_join(ferritin_data, by = c("person_id", "transplant_date"))                    
          
aim_2a_2 <- stat_dataset_ferritin %>% filter(chart_completion, eligibility == "Yes", !is.na(ferritin_pre)) %>%
                  mutate(ferritin_pre = factor(ferritin_pre, levels = c("low", "moderate", "high")),
                        ferritin_pre_max = factor(ferritin_pre_max, levels = c("low", "moderate", "high"))) 

auto_ct <- aim_2a_2 %>% filter(!is.na(survival_time_5yr_censored)) %>% filter(transplant_type_combined == "autologous/genetherapy") %>% nrow()
unknown_ct <- aim_2a_2 %>% filter(!is.na(survival_time_5yr_censored)) %>% filter(transplant_type_combined == "Unknown") %>% nrow()

pre_ferritin_cox <- aim_2a_2 %>% filter(!is.na(survival_time_5yr_censored)) %>%
                          select(record_id, ferritin_pre, age_at_ce, gender, graft_manip, transplant_type, scd_type, disease_relapse, graft_fail_date, donor_relation, survival_time_5yr_censored, survival_5yr, second_transplant, transplant_type_combined, graft_fail_5yr, graft_fail_days_since_ce_5yr_censored, has_busulfan)
label(pre_ferritin_cox$transplant_type_combined) = "match status"
table1(~gender + disease_relapse + age_at_ce + transplant_type_combined + ferritin_pre + has_busulfan + scd_type| survival_5yr, 
        data = pre_ferritin_cox %>% mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")), 
        overall = FALSE, 
        caption = "Overall survival at 5-yr following transplant")                          
```

- Multivariate Cox proportional hazard model adjusted for: 
  - pre-transplant most recent ferritin levels: low(<1000 ng/mL), moderate (1000-2500 ng/mL), or high (>=2500 ng/mL)
  - age at transplant: continuous 
  - gender: female or male
  - match status: 
    - Allogeneic: Cord Blood, Haploidentical, Matched related, Matched unrelated, Mismatched related
    - Removed: Mismatched related, Unknown, and autologous/genetherapy (see table above) due to missing obversation in outcomes. 
  - has busulfan: yes or no

- **Results**:
  - Increasing age is associated with a lower probability of survival. Although this association is statistically significant, the clinical impact appears to be minimal. It might be more informative to analyze this by breaking it down into age groups.
  - Bulsufan treatment increased the death probability by almost 80% and this is borderline significant. 

```{r overall_surv_pre_ferritin, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
# pre_ferritin_cox %>% group_by(survival_5yr, transplant_type_combined) %>% summarise(n = n_distinct(record_id)) %>% view()
pre_ferritin_cox_1 <- pre_ferritin_cox %>% filter(!(transplant_type_combined %in% c("autologous/genetherapy", "mismatched related")))
# 0 is censored/survived, 1 is death
# exclude match_status and graft_manip
cox_ferritin_os <- coxph(Surv(survival_time_5yr_censored/365.25, survival_5yr) ~ ferritin_pre + age_at_ce + gender + transplant_type_combined + has_busulfan, 
                data = pre_ferritin_cox_1)

# pre_ferritin_cox_1 %>% group_by(ferritin_pre, gender, transplant_type_combined, has_busulfan) %>%
#                       summarise(n_distinct(record_id))
# visualiza the impact on iron levels
ferritin_levels_df <- with(pre_ferritin_cox_1,
               data.frame(ferritin_pre = c("low", "moderate", "high"), 
                          age_at_ce = rep(mean(age_at_ce, na.rm = TRUE), 3),
                          gender = rep("Female", 3),
                          transplant_type_combined = rep("matched related",3),
                          has_busulfan = rep("No", 3)
                          # match_status = rep("Cord Blood", 3), 
                          # disease_relapse = rep("No", 3), 
                          # graft_manip = rep("No", 3), 
                          # scd_type = rep("beta thalassemia major", 3)
                          )
               )
fit <- survfit(cox_ferritin_os, newdata = ferritin_levels_df)
ggsurvplot(fit, data = ferritin_levels_df, 
                  conf.int = FALSE, ggtheme = theme_minimal(),
                  legend.labs = c("Low", "Moderate", "High"),
                  legend.title = "Pre-transplant ferritin levels", 
                  xlab = "Years", ylab = "Overall survival probability",
                  xlim = c(0, 5), ylim = c(0.5, 1.0), 
                  risk.table = TRUE)#, risk.table.title = "test")

# output overall survival curves
# survfit2(res.cox) %>% ggsurvfit() +
#   labs(
#     x = "Years",
#     y = "Overall survival probability",
#     title = "Overall survival analysis pre-transplant ferritin"
#     ) + 
#   add_confidence_interval() +
#   add_risktable()
```

```{r overall_surv_pre_ferritin_sum, echo= FALSE, message=FALSE}
# summary(res.cox)
cox_sum(cox_ferritin_os) %>% prettify_kable()
```

## Aim 2a-1: Overall survival stratified by <span style="color:red">pre-transplant LIC levels</span>
```{r overall_surv_pre_LIC_data, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
LIC_data <- results_tbl("cr_LIC_data") %>% 
                  select(record_id, transplant_date, LIC, LIC_type, LIC_date, LIC_days_since_transplant, LIC_level, 
                  LIC_3, LIC_days_since_transplant_3, LIC_3_level,
                  LIC_max, LIC_3_max, LIC_days_since_transplant_max, LIC_days_since_transplant_3_max,
                  LIC_max_level, LIC_3_max_level, LIC_level_3_binary, LIC_level_3_max_binary) %>%
                  collect()
                  
stat_dataset_LIC <- stat_dataset %>% left_join(LIC_data, by = c("record_id", "transplant_date"))

aim_2a_1 <- stat_dataset_LIC %>% filter(chart_completion, eligibility == "Yes", LIC_type == "pre-transplant", !is.na(LIC_level)) %>% 
                mutate(LIC_level_3_binary = factor(LIC_level_3_binary, levels = c("low", "high")))

pre_LIC_cox <- aim_2a_1 %>% filter(!is.na(survival_time_5yr_censored)) %>%
                          select(record_id, LIC_level_3_binary, age_at_ce, gender, graft_manip, transplant_type, scd_type, disease_relapse, graft_fail_date, donor_relation, survival_time_5yr_censored, survival_5yr, second_transplant, transplant_type_combined, graft_fail_5yr, graft_fail_days_since_ce_5yr_censored, has_busulfan, site)
# pre_LIC_cox %>% group_by(survival_5yr, transplant_type_combined) %>% summarise(n = n_distinct(record_id)) %>% view()

auto_ct <- aim_2a_1 %>% filter(!is.na(survival_time_5yr_censored)) %>% filter(transplant_type_combined == "autologous/genetherapy") %>% nrow()
unknown_ct <- aim_2a_1 %>% filter(!is.na(survival_time_5yr_censored)) %>% filter(transplant_type_combined == "Unknown") %>% nrow()

label(pre_LIC_cox$transplant_type_combined) = "match status"
label(pre_LIC_cox$LIC_level_3_binary) = "pre-transplant LIC-level"
table1(~gender + disease_relapse + age_at_ce + transplant_type_combined + LIC_level_3_binary + has_busulfan + scd_type| survival_5yr, 
        data = pre_LIC_cox %>% mutate(survival_5yr = ifelse(survival_5yr == 1, "Dead", "Survived")), 
        overall = FALSE, 
        caption = "Overall survival at 5-yr following transplant")
```

- Multivariate Cox proportional hazard model adjusted for: 
  - pre-transplant most recent LIC levels: low(<5 mg Fe/g dry wt), or high (>5 mg Fe/g dry wt)
  - age at transplant: continuous 
  - gender: female or male
  - match status: 
    - Allogeneic: Cord Blood, Haploidentical, matched related, matched unrelated, mismatched related, mismatched unrelated
    - Removed: Unknown and autologous/genetherapy (see table above) due to missing obversation in outcomes. 
  - has busulfan: yes or no

```{r overall_surv_pre_LIC, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
pre_LIC_cox_1 <- pre_LIC_cox %>% #filter(transplant_type_combined != "Unknown") %>% 
    filter(transplant_type_combined != "autologous/genetherapy")
# 0 is censored/survived, 1 is death
# exclude match_status and graft_manip
cox_LIC_os <- coxph(Surv(survival_time_5yr_censored/365.25, survival_5yr) ~ LIC_level_3_binary + age_at_ce + gender + transplant_type_combined + has_busulfan, 
                data = pre_LIC_cox_1)

# visualiza the impact on iron levels
LIC_levels_df <- with(pre_LIC_cox_1,
               data.frame(LIC_level_3_binary = c("low", "high"), 
                          age_at_ce = rep(mean(age_at_ce, na.rm = TRUE), 2),
                          gender = rep("Female", 2),
                          transplant_type_combined = rep("matched related",2),
                          has_busulfan = rep("No", 2)
                          )
               )
fit <- survfit(cox_LIC_os, newdata = LIC_levels_df)
ggsurvplot(fit, data = LIC_levels_df, 
                  conf.int = FALSE, ggtheme = theme_minimal(),
                  legend.labs = c("Low", "High"),
                  legend.title = "Pre-transplant LIC levels", 
                  xlab = "Years", ylab = "Overall survival probability",
                  xlim = c(0, 5), ylim = c(0.5, 1.0)) 
                  #risk.table = TRUE, risk.table.title = "test")

# output overall survival curves
# survfit2(res.cox) %>% ggsurvfit() +
#   labs(
#     x = "Years",
#     y = "Overall survival probability",
#     title = "Overall survival analysis pre-transplant ferritin"
#     ) + 
#   add_confidence_interval() +
#   add_risktable()
```

```{r overall_surv_pre_LIC_sum, echo= FALSE, message=FALSE}
# summary(res.cox)
cox_sum(cox_LIC_os) %>% prettify_kable()
```

# Graft failure
## Aim 2a-2: Graft failure stratified by <span style="color:red">pre-transplant ferritin levels</span>
```{r graft_fail_pre_ferritin_data, echo= FALSE, message=FALSE}
unknown_ct <- pre_ferritin_cox %>% filter(!is.na(graft_fail_days_since_ce_5yr_censored)) %>% filter(transplant_type_combined == "Unknown") %>% nrow()

label(pre_ferritin_cox$transplant_type_combined) = "match status"
label(pre_ferritin_cox$survival_5yr) = "Dies within 5 years of post-transplant"

table1(~gender + age_at_ce + transplant_type_combined + ferritin_pre + has_busulfan + survival_5yr| graft_fail_5yr, 
        data = pre_ferritin_cox %>% mutate(graft_fail_5yr = ifelse(graft_fail_5yr ==1, "Graft failure", "Graft failure free")) %>%
        mutate(survival_5yr = ifelse(survival_5yr == 1, "No", "Yes")), 
        overall = FALSE, 
        caption = "Graft failure at 5-yr following transplant")                          
```

- Competing risk analysis adjusted for: 
  - pre-transplant most recent ferritin levels: low(<1000 ng/mL), moderate (1000-2500 ng/mL), or high (>=2500 ng/mL)
  - age at transplant: continuous 
  - gender: female (ref) or male
  - match status: Allogeneic: Cord Blood, matched related (ref), matched unrelated/mismatched related/mismatched unrelated, autologous/genetherapy, and Unknown
  - has busulfan: yes or no
  - Outcomes: graft failure or death without graft failure 
- Composite outcome: event-free survival: death/graft failure (whichever is first) 
- n = `r unknown_ct` patients with unknown match status were removed from the analysis
- 1 patient with graft failure dates before transplant dates were removed from this analysis: pending Nora's response. 

```{r graft_fail_pre_ferritin, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
# plot distribution of time to event for graft failure
# cohort_covars %>%
#   mutate(graft_fail_5yr = recode(graft_fail_5yr, "0" = "censored", "1" = "graft failure", "2" = "death")) %>%
#   ggplot(aes(x = graft_fail_days_since_ce_5yr_censored/30, group = graft_fail_5yr, fill = factor(graft_fail_5yr))) +
#   geom_histogram()

pre_ferritin_cox_2 <- pre_ferritin_cox %>% 
                          filter(transplant_type_combined != "Unknown") %>% 
                          # filter(transplant_type_combined != "autologous/genetherapy") %>% 
                          filter(!is.na(graft_fail_5yr)) %>% 
                          mutate(graft_fail_5yr = recode(graft_fail_5yr, "0" = "censor", "1" = "graft failure", "2" = "death")) %>%
                          mutate(graft_fail_5yr = factor(graft_fail_5yr,
                                                  levels = c("censor", "graft failure", "death"))) %>% 
                          filter(graft_fail_days_since_ce_5yr_censored >= 0) %>%
                          mutate(transplant_type_combined = relevel(factor(transplant_type_combined), ref = "matched related"))

# 0 is censored/survived, 1 is death
# exclude match_status and graft_manip
# km <- survfit(Surv(graft_fail_days_since_ce_5yr_censored/30, graft_fail_5yr) ~ ferritin_pre# + age_at_ce + gender + transplant_type_combined + has_busulfan, 
#                 data = pre_ferritin_cox_2)

# km$transitions
# this plot function is especially for univariate analysis only
# cant do this for multivariate

# plot(km[, "graft failure"], 
#               mark.time = FALSE, lwd = 2, xscale = 12,
#               conf.int = FALSE, #lty = "dashed",
#               col = c("red", "blue", "black"), 
#               # linecolor = c("r", "b", "blue"),
#               main = "Graft failure probability --- Competing Risk Kaplan Meier univariate analysis", 
#               xlab = "Years post transplant", ylab = "Graft failure probability")
# legend(2*12, .02, c("low ferritin", "moderate ferritin", "high ferritin"), 
#       col = c("red", "blue", "black"), lty = "solid")

compRisk_ferritin_gf <- coxph(Surv(graft_fail_days_since_ce_5yr_censored/30, graft_fail_5yr) ~ ferritin_pre + age_at_ce + gender + has_busulfan  + transplant_type_combined, 
                data = pre_ferritin_cox_2, 
                id = record_id)
# print(res.cox, digits = 3)                
# ggsurvplot(res.cox)

# visualiza the impact on iron levels
# ferritin_levels_df <- with(pre_ferritin_cox_2,
#                data.frame(ferritin_pre = c("low", "moderate", "high"), 
#                           age_at_ce = rep(mean(age_at_ce, na.rm = TRUE), 3),
#                           gender = rep("Female", 3),
#                           transplant_type_combined = rep("matched related",3),
                          # has_busulfan = rep("No", 3)))

# fit <- survfit(res.cox, newdata = ferritin_levels_df)
# ggsurvplot(fit, data = ferritin_levels_df, 
#                   conf.int = FALSE, ggtheme = theme_minimal(),
#                   legend.labs = c("Low", "Moderate", "High"),
#                   legend.title = "Pre-transplant ferritin levels", 
#                   xlab = "Years", ylab = "Graft failure free probability",
#                   xlim = c(0, 5), ylim = c(0.5, 1.0), 
#                   risk.table = FALSE)#, risk.table.title = "test")
```

- Having a matched unrelated/mismatched unrelated/mismatched related transplant has a significant effect on graft failure. The risk of having graft failure increased by almost 3 folds comparing to having a matched related transplant. Gender is another significant factor, risk of graft failure in male patients is 2 and a haft times higher than females. 

- Patients with a moderate ferritin levels are 60% less likely to die without graft failure than patients with low ferritin level pre-transplant. Age is another signicant factor; however, the OR is 1.07 which indicates that older patients have higher risk of death without graft failure than younger patients (Nora, does it make sense?). Being treated with bulsufan increased the risk death without GF by almost 3 folds. No patients receiving augologous/gene therapy transplant dies without graft failure; thus, this coefficient could be ignored. 

- In the table below: 1:2 represents transition to Graft failure and 1:3 represents transition to death without graft failure
```{r graft_fail_pre_ferritin_sum, echo= FALSE, message=FALSE}
# summary(res.cox)
cox_sum(compRisk_ferritin_gf) %>% prettify_kable()
```

## Aim 2a-1: Graft failure stratified by <span style="color:red">pre-transplant LIC levels</span>
```{r graft_fail_pre_LIC_data, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
# pre_LIC_cox %>% group_by(survival_5yr, transplant_type_combined) %>% summarise(n = n_distinct(record_id)) %>% view()

cord_blood_ct <- aim_2a_1 %>% filter(!is.na(graft_fail_days_since_ce_5yr_censored)) %>% filter(transplant_type_combined == "Allogeneic: Cord Blood") %>% nrow()
unknown_ct <- aim_2a_1 %>% filter(!is.na(graft_fail_days_since_ce_5yr_censored)) %>% filter(transplant_type_combined == "Unknown") %>% nrow()

label(pre_LIC_cox$transplant_type_combined) = "match status"
label(pre_LIC_cox$LIC_level_3_binary) = "pre-transplant LIC-level"
label(pre_LIC_cox$survival_5yr) = "Dies within 5 years of post-transplant"

table1(~gender + age_at_ce + transplant_type_combined + LIC_level_3_binary + has_busulfan + survival_5yr| graft_fail_5yr, 
        data = pre_LIC_cox %>% mutate(graft_fail_5yr = recode(graft_fail_5yr, "0" = "censor", "1" = "graft failure", "2" = "death")) %>%
                      mutate(graft_fail_5yr = factor(graft_fail_5yr,
                                              levels = c("censor", "graft failure", "death"))) %>% 
                      filter(graft_fail_days_since_ce_5yr_censored >= 0) %>%
                      mutate(survival_5yr = ifelse(survival_5yr == 1, "No", "Yes")), 
        overall = FALSE, 
        caption = "Graft failure at 5-yr following transplant")
```

- Competing risk analysis adjusted for: 
  - pre-transplant most recent LIC levels: low(<5 mg Fe/g dry wt), or high (>5 mg Fe/g dry wt)
  - age at transplant: continuous 
  - gender: female (ref) or male
  - match status: Allogeneic: Cord Blood, matched related (ref), matched unrelated/mismatched related/mismatched unrelated, autologous/genetherapy, and Unknown
  - has busulfan: yes or no
  - Outcomes: graft failure or death without graft failure 
- All patients with cord blood transplant survived (n = `r cord_blood_ct`). 
- n = `r unknown_ct` patients with unknown match status

- Composite outcome: event-free survival: death/graft failure (whichever is first) 

```{r graft_fail_pre_LIC, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
pre_LIC_cox_2 <- pre_LIC_cox %>% 
                      filter(!(transplant_type_combined %in% c("Allogeneic: Cord Blood", "Unknown"))) %>%
                      # filter(transplant_type_combined != "autologous/genetherapy") %>% 
                      filter(!is.na(graft_fail_5yr)) %>% 
                      mutate(graft_fail_5yr = recode(graft_fail_5yr, "0" = "censor", "1" = "graft failure", "2" = "death")) %>%
                      mutate(graft_fail_5yr = factor(graft_fail_5yr,
                                              levels = c("censor", "graft failure", "death"))) %>% 
                      filter(graft_fail_days_since_ce_5yr_censored >= 0) %>%
                      mutate(transplant_type_combined = relevel(factor(transplant_type_combined), ref = "matched related"))

# 0 is censored/survived, 1 is death
# exclude match_status and graft_manip
compRisk_LIC_gf <- coxph(Surv(graft_fail_days_since_ce_5yr_censored/365.25, graft_fail_5yr) ~ LIC_level_3_binary + age_at_ce + gender + transplant_type_combined + has_busulfan, 
                data = pre_LIC_cox_2, 
                id = record_id)

# visualiza the impact on iron levels
# LIC_levels_df <- with(pre_ferritin_cox_2,
#                data.frame(LIC_level_3_binary = c("low", "high"), 
#                           age_at_ce = rep(mean(age_at_ce, na.rm = TRUE), 2),
#                           gender = rep("Female", 2),
#                           transplant_type_combined = rep("matched unrelated",2),
#                           has_busulfan  = rep("No", 2)))
# fit <- survfit(res.cox, newdata = LIC_levels_df)
# ggsurvplot(fit, data = LIC_levels_df, 
#                   conf.int = FALSE, ggtheme = theme_minimal(),
#                   legend.labs = c("Low", "High"),
#                   legend.title = "Pre-transplant LIC levels", 
#                   xlab = "Years", ylab = "Graft failure free probability",
#                   xlim = c(0, 5), ylim = c(0.5, 1.0)) 
#                   #risk.table = TRUE, risk.table.title = "test")

```

- The effects using pre-transplant LIC are similar to that of ferritin, having a matched unrelated/mismatched unrelated/mismatched related transplant increased the risk of graft failure by almost 4 folds comparing to having a matched related transplant. Gender is another significant factor, risk of graft failure in male patients is almost 3 times higher than females. 

- For death without graft failure, none of the effect is significant (no patients receiving augologous/gene therapy transplant dies without graft failure; thus, this coefficient could be ignored) 

- In the table below: 1:2 represents transition to Graft failure and 1:3 represents transition to death without graft failure
- Nora, patients with high LIC level are less likely to get graft failure, is it because these patients are more likely to receive bulsufan? 

```{r graft_fail_pre_LIC_sum, echo= FALSE, message=FALSE}
# summary(res.cox)
cox_sum(compRisk_LIC_gf) %>% prettify_kable()
```

# Bacteremia
## Bacteremia stratified by <span style="color:red">pre-transplant ferritin levels</span>
## Bacteremia stratified by <span style="color:red">pre-transplant LIC levels</span>

# Address missing data

# Sensitivity analysis
## Overall survival: Subgroup analysis for pre-transplant LIC binary classification
### sickle cell disease --- Overall survival: Subgroup analysis for pre-transplant LIC binary levels

```{r pre_LIC_binary_OS_disease_subgroup_scd}
table1(~age_at_ce + gender + transplant_type_combined + LIC_level_3_binary + has_busulfan| survival_5yr, 
        data = pre_LIC_cox %>% mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")) %>%
                filter(scd_type == "sickle cell disease"), 
        overall = FALSE, 
        caption = "Overall survival at 5-yr following transplant --- sickle cell disease")     

# pre_LIC_os_btm <- pre_LIC_cox %>%
#     filter(!(transplant_type_combined %in% c("mismatched related", "mismatched unrelated", "autologous/genetherapy"))) %>%
#     cox_fit_plot(time_var = "survival_time_5yr_censored", 
#                 status_var = "survival_5yr", 
#                 covariates = c("LIC_level_3_binary", "age_at_ce", "gender", "has_busulfan", "transplant_type_combined"),
#                 iron_level_var = "LIC_level_3_binary", 
#                 plot_labels = c("Pre-transplant LIC binary levels --- beta thalassemia major", "Overall survival probability"),
#                 sub_group = "beta thalassemia major")

# print(pre_LIC_os_btm$p)
# cox_sum(pre_LIC_os_btm$fit) %>% prettify_kable()
```

### sickle cell disease --- Overall survival: Subgroup analysis for pre-transplant LIC binary levels

```{r pre_LIC_binary_OS_disease_subgroup_aa}
table1(~age_at_ce + gender + transplant_type_combined + LIC_level_3_binary + has_busulfan| survival_5yr, 
        data = pre_LIC_cox %>% mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")) %>%
                filter(scd_type == "aplastic anemia"), 
        overall = FALSE, 
        caption = "Overall survival at 5-yr following transplant --- aplastic anemia")     

pre_LIC_os_aa <- pre_LIC_cox %>%
    filter(!(transplant_type_combined %in% c("Unknown", "Haploidentical", "autologous/genetherapy", "Allogeneic: Cord Blood"))) %>%
    cox_fit_plot(time_var = "survival_time_5yr_censored", 
                status_var = "survival_5yr", 
                covariates = c("LIC_level_3_binary", "age_at_ce", "gender", "has_busulfan", "transplant_type_combined"),
                iron_level_var = "LIC_level_3_binary", 
                plot_labels = c("Pre-transplant LIC binary levels --- aplastic anemia", "Overall survival probability"),
                sub_group = "aplastic anemia")

print(pre_LIC_os_aa$p)
cox_sum(pre_LIC_os_aa$fit) %>% prettify_kable()
```

### beta thalassemia major  --- Overall survival: Subgroup analysis for pre-transplant LIC binary levels

```{r pre_LIC_binary_OS_disease_subgroup_btm}
table1(~age_at_ce + gender + transplant_type_combined + LIC_level_3_binary + has_busulfan| survival_5yr, 
        data = pre_LIC_cox %>% mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")) %>%
                filter(scd_type == "beta thalassemia major"), 
        overall = FALSE, 
        caption = "Overall survival at 5-yr following transplant --- beta thalassemia major")     

pre_LIC_os_btm <- pre_LIC_cox %>%
    filter(!(transplant_type_combined %in% c("mismatched related", "mismatched unrelated", "autologous/genetherapy"))) %>%
    cox_fit_plot(time_var = "survival_time_5yr_censored", 
                status_var = "survival_5yr", 
                covariates = c("LIC_level_3_binary", "age_at_ce", "gender", "has_busulfan", "transplant_type_combined"),
                iron_level_var = "LIC_level_3_binary", 
                plot_labels = c("Pre-transplant LIC binary levels --- beta thalassemia major", "Overall survival probability"),
                sub_group = "beta thalassemia major")

print(pre_LIC_os_btm$p)
cox_sum(pre_LIC_os_btm$fit) %>% prettify_kable()
```

## Overall survival: Subgroup analysis for pre-transplant ferritin levels
### aplastic anemia  --- Overall survival: Subgroup analysis for pre-transplant ferritin levels
```{r pre_ferritin_OS_disease_subgroup_aa}
table1(~age_at_ce + gender + transplant_type_combined + ferritin_pre + has_busulfan| survival_5yr, 
        data = pre_ferritin_cox %>% mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")) %>%
                filter(scd_type == "aplastic anemia"), 
        overall = FALSE, 
        caption = "Overall survival at 5-yr following transplant --- aplastic anemia")                          

pre_ferritin_os_aa <- pre_ferritin_cox %>%
    filter(!(transplant_type_combined %in% c("Unknown", "autologous/genetherapy"))) %>%
    cox_fit_plot(time_var = "survival_time_5yr_censored", 
                status_var = "survival_5yr", 
                covariates = c("ferritin_pre", "age_at_ce", "gender", "transplant_type_combined", "has_busulfan"),
                iron_level_var = "ferritin_pre", 
                plot_labels = c("Pre-transplant ferritin levels --- aplastic anemia", "Overall survival probability"),
                sub_group = "aplastic anemia")
print(pre_ferritin_os_aa$p)
cox_sum(pre_ferritin_os_aa$fit) %>% prettify_kable()
```

### sickle cell disease  --- Overall survival: Subgroup analysis for pre-transplant ferritin levels
```{r pre_ferritin_OS_disease_subgroup_scd}
table1(~age_at_ce + gender + transplant_type_combined + ferritin_pre + has_busulfan| survival_5yr, 
        data = pre_ferritin_cox %>% mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")) %>%
                filter(scd_type == "sickle cell disease"), 
        overall = FALSE, 
        caption = "Overall survival at 5-yr following transplant --- sickle cell disease")                          

pre_ferritin_os_scd <- pre_ferritin_cox %>%
    filter((transplant_type_combined %in% c("matched related", "mismatched unrelated", "matched unrelated"))) %>%
    cox_fit_plot(time_var = "survival_time_5yr_censored", 
                status_var = "survival_5yr", 
                covariates = c("ferritin_pre", "age_at_ce", "gender", "has_busulfan","transplant_type_combined"),
                iron_level_var = "ferritin_pre", 
                plot_labels = c("Pre-transplant ferritin levels --- sickle cell disease", "Overall survival probability"),
                sub_group = "sickle cell disease")

print(pre_ferritin_os_scd$p)
cox_sum(pre_ferritin_os_scd$fit) %>% prettify_kable()
```

### beta thalassemia major  --- Overall survival: Subgroup analysis for pre-transplant ferritin levels
```{r pre_ferritin_OS_disease_subgroup_btm}
table1(~age_at_ce + gender + transplant_type_combined + ferritin_pre + has_busulfan| survival_5yr, 
        data = pre_ferritin_cox %>% mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")) %>%
                filter(scd_type == "beta thalassemia major"), 
        overall = FALSE, 
        caption = "Overall survival at 5-yr following transplant --- beta thalassemia major")     

pre_ferritin_os_btm <- pre_ferritin_cox %>%
    filter(!(transplant_type_combined %in% c("mismatched related", "mismatched unrelated", "autologous/genetherapy"))) %>%
    cox_fit_plot(time_var = "survival_time_5yr_censored", 
                status_var = "survival_5yr", 
                covariates = c("ferritin_pre", "age_at_ce", "gender", "has_busulfan", "transplant_type_combined"),
                iron_level_var = "ferritin_pre", 
                plot_labels = c("Pre-transplant ferritin levels --- beta thalassemia major", "Overall survival probability"),
                sub_group = "beta thalassemia major")

print(pre_ferritin_os_btm$p)
cox_sum(pre_ferritin_os_btm$fit) %>% prettify_kable()

# pre_ferritin_cox_1 %>% filter(scd_type == "sickle cell disease") %>%
#       group_by(transplant_type_combined, survival_5yr) %>% summarise(n = n_distinct(record_id))
```


- Aim 2a and 2b: Subgroup analyses by disease, specifically sickle cell disease 
- Aim 2a: VOD- stratify outcome by busulfan 
- Different cut-off thresholds for LIC and ferritin
- Average of the latest LIC, show the change in their LIC levels

