---
title: "Statistical analysis: The Effects of Iron Overload on Hemptotic Stemcell Transplant (HSCT) Outcomes in Sickle Cell Patients"
author: "Nhat Nguyen \n PEDSnet Coordinating Center"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    page-layout: full
    embed-resources: true
    smooth-scroll: true
    fig-cap: TRUE
    df-print: paged
    toc: TRUE
    toc-depth: 7
    toc-collapsed: TRUE
    toc-float: TRUE
    toc-location: left
    number-sections: TRUE
    number-depth: 5
    output-file: Gibson_sickle_cell_statisical_analysis
    output-ext: "html"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 6, fig.align = "center")
# Default taken from R notebook behavior: when knitting, wd will always be location of notebook
base_dir <- '..'
Sys.setenv(PEDSNET_SKIP_REQUEST_EXECUTION=1)
Sys.setenv(db_version = "5.1")
tryCatch(source('../site/run.R'),
         error = function (e) message(e)) # May not be able to make db connection

# Set to "local" to read data from ../results, or another value to read from db
data_source <- "not_local" #if_else(config('execution_mode') == 'distribution', 'local', 'not_local')

require(tidyr)
require(knitr)
require(kableExtra)
require(stringr)
require(tibble)
require(ggplot2)
require(table1)
require(tidyverse)
library(survival)
library(survminer)
library(ggsurvfit)
library(gtsummary)
library(cowplot)
library(mstate) # multi-state competing risk model
require(ggpubr) #linear fit 
library(WeightIt)
library(lme4)
require(broom.mixed) # plot plot_summs and export_summs
require(sjPlot)
require(jtools)
require(gridExtra)
```

```{r funtioncs, include=FALSE}
get_codesets <- function(codeset_name) {
  read_csv(paste0('../specs/', codeset_name, '.csv'), col_types = c('c', 'c', 'c', 'c'), show_col_types = FALSE) %>%
  mutate(concept_id = as.character((concept_id)), concept_code = as.character((concept_code))) %>% collect()
}

get_results <- function(tbl_name) {
  if (data_source == 'local') {
    rslt <- read.csv(paste0('../local/', tbl_name, '.csv'))
  }
  else {
    rslt <- results_tbl(tbl_name) %>% collect()
  }
  rslt
}

prettify_kable <- function(data) {
  rslt <- data %>% kable(digits = 4, format.args = list(big.mark = ','))
  if (knitr::is_html_output()) rslt <- rslt %>%
      kable_styling(bootstrap_options = c("striped", "hover")) %>%
      column_spec(1, bold = T, border_right = T)
  rslt
}

my.render.cont <- function(x) {
    with(stats.apply.rounding(stats.default(x), digits=2), c("",
        "Median [Min, Max]"=sprintf("%s [%s, %s])", MEDIAN, MIN, MAX)))
}

cox_sum <- function(cox_sum){
  cox_sum <- summary(cox_sum)
  res <- cbind.data.frame(cox_sum$conf.int, cox_sum$coefficients)

  res <- res[c("exp(coef)", "lower .95", "upper .95", "Pr(>|z|)")] %>% 
              rename(HR = "exp(coef)", lower = "lower .95", upper = "upper .95", p_val = "Pr(>|z|)") %>%
              mutate(`95 CI` = paste0("(",round(lower, 2), " - ", round(upper, 2), ")")) %>%
              mutate(p_val = case_when(p_val <= 0.0001~ paste0(round(p_val, 5), "***"), 
                                      p_val <= 0.001~ paste0(round(p_val, 5), "**"), 
                                      p_val <= 0.01~ paste0(round(p_val, 5), "*"), 
                                      p_val <= 0.05~ paste0(round(p_val, 5), "."), 
                                      TRUE ~ as.character(round(p_val, 5)))) %>%
              mutate(HR = ifelse(HR <= 0.005, "<0.005", as.character(round(HR, 2)))) %>%
              select("HR", "95 CI", "p_val")
  return(res)
}

cox_fit_plot <- function(data, time_var, status_var, covariates, iron_level_var, plot_labels, sub_group){
  data <- data %>% filter(scd_type == sub_group)

  # Create the formula 
  formula_str <- paste("Surv(", time_var, "/365.25,", status_var, ") ~", paste(covariates, collapse = " + "))
  formula <- as.formula(formula_str)

  # Fit the Cox proportional hazards model
  cox_model <- coxph(formula, data = data)

  iron_levels <- data %>% distinct(!!sym(iron_level_var)) 

  # visualiza the impact on iron levels
  iron_levels_df <- with(data,
               data.frame(iron_levels, 
                          age_at_ce = rep(mean(age_at_ce, na.rm = TRUE), length(iron_levels)),
                          gender = rep("Female", length(iron_levels)),
                          transplant_type_combined = rep("matched related",length(iron_levels)),
                          has_busulfan = rep("No", length(iron_levels))
                          )
               )

  # print(iron_levels_df)
  # iron_levels_df <- iron_levels_df %>% rename(iron_level_var = test)
  
  plot_fit <- survfit(cox_model, newdata = iron_levels_df)
  p <- ggsurvplot(plot_fit, data = iron_levels_df, 
                  conf.int = FALSE, #ggtheme = theme_minimal(),
                  legend.labs = iron_levels %>% pull(), #c("Low", "High"),
                  legend.title = plot_labels[1],  
                  xlab = "Years", ylab = plot_labels[2], 
                  title = paste(plot_labels[1], sub_group, sep = "-----"), 
                  xlim = c(0, 5), ylim = c(0.7, 1.0),
                  ggtheme = theme(plot.title = element_text(size = plot_title_text_size),    # Title text size
                              axis.title.x = element_text(size = axis_text_size), # X-axis title size
                              axis.title.y = element_text(size = axis_text_size), # Y-axis title size
                              axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
                              axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
                              legend.title = element_text(size = axis_text_size),    # Legend title size
                              legend.text = element_text(size = axis_text_size))) 

  return(list(fit = cox_model, p = p))
}
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
tribble( ~Item, ~Value,
         'Title', config('title'),
         'PI', config('requester_name'),
         'CDM Version', as.character(config('cdm_version'))) %>%
  prettify_kable()

```
```{r data, include= FALSE}
cohort <- results_tbl("study_cohorts") %>%
                    pivot_longer(cols = aim_2a_2:aim_3_2, names_to = "aim", values_to = "which_aim") %>% 
                    filter(which_aim == TRUE) %>% 
                    collect() %>%
                    mutate(age_at_ce = round(as.numeric(difftime(transplant_date, birth_date, units = "days"))/365.25)) %>%
                    # mutate(transplant_type = case_when(grepl("allogeneic", transplant_concept_name, ignore.case = TRUE) ~ "allogeneic",
                    #                           grepl("autologous", transplant_concept_name, ignore.case = TRUE) ~ "autologous",
                    #                           TRUE ~ "unknown")) %>% 
                    select(person_id, aim, site, age_at_ce , gender, scd_concept_name, scd_type, transplant_concept_name, transplant_type, transplant_date)
cohort_ct <- cohort %>% distinct_ct()
```

# Data analysis plan:
- This is a retrospective cohort study comparing outcomes between patients with pre/post-HSCT iron overload to patients without pre/post-HSCT iron overload. Analyze the impact of pre/post-HSCT elevated ferritin and LIC on short-term hematologic and transplant-related outcome: 
  - overall survival
  - graft-versus-host disease (GvHD)
  - graft failure
  - incidence of veno-occlusive disease (VOD)
  - bacteremia
  - immune reconstitution
  - time to neutrophil and platelet engraftment. 

## Aim_2a: 
- GRFS (Graft failure free, GVHD-free survival survival): Time-to-event analysis with Cox proportional hazards model, Kaplan-Meier curves with log-rank test 
- GVHD: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test.  
- Disease relapse: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test.  
- VOD: Multivariable logistic regression with stratification for busulfan conditioning, use interaction term for busulfan and iron score 
- Time to neutrophil and platelet engraftment: time-to-event analysis using a competing risks analysis. Competing risks include death or graft failure prior to engraftment. 
- Bacteremia within 1 year after HSCT: Time-to-event analysis using a competing risks analysis. Competing risk includes death. 
- Immune Reconstitution: time-to-event analysis using a competing risks analysis. Competing risks include death and graft failure requiring a second transplant. 

## Aim_2b: 
- GRFS: Time-to-event analysis with Cox proportional hazards model, Kaplan-Meier curves with log-rank test 
- Overall survival: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test.  
- Disease relapse: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test.  
- GVHD: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test. 

# Background:
- Armand et al 2007:
  - n = 590, follow-up times = 8yrs since transplant
  - Disease: acute leukemia, MDS, chronic myelogenous leukemia CML, non-Hodgkin lymphoma NHL. 
  - Ferritin_threshold = 2034 ng/mL: significant negative association between ferritin levels and OS and disease-free survival (DFS) for MDS and acute leukemia. For MDS, an independent significant association between higher ferritin levels and increased treatment-related mortality (TRM). No significant effects observed for CML or NHL
- Armand et al 2012: 
  - n = 45
  - Disease: acute leukemia (acute myelogenous leukemia (AML), acute lymphoblastic leukemia (ALL)) and myelodyplastic syndrome (MDS)
  - LIC_threshold = 5 (g/gdw): higher LIC level is associated with higher 1-yr OS and lower 2-yr OS but not significant. Using threshold of 3 or 7 did not change the results. 
  - Ferrintin_threshold = 2000 ng/ml: higher baseline ferritin level is significantly associated with lower OS at both 1-yr and 2-yr post-transplant. However, neither the association between baseline ferritin and cummulative incidence of relapse (CIR) nor non-relapse mortality (NMR) is significant. 
  - The authors suggested that ferritin rather than LIC is prognostic in this case not because it measures iron stores more accurately but because it's an acute phase reactants (such as albumin or C-reactie protein) which reflects inflammatory status from comorbid conditions, advance diseases, etc..  
- Armand et al 2014:
  - Meta-analysis from 4 published cohorts n = 274 with AML and MDS
  - They looked at the correlation between SF and LIC, and SF of 2500 ng/mL corresponds to an LIC around 6mg/gdw
  - 2 thresholds for each variables LIC 5 & 7 g/gdw, and SF 1000 and 2500 ng/mL. 
  - No siginificant association between SF or LIC and OS or TRM observed for thresholds of 7 and 1000, respectively in all 4 cohorts. 
  - In the pooled cohort, no siginificant association between LIC and OS or TRM observed for LIC threshold of 7. The association between SF and OS is only significant at 1000ng/mL but not at 2500ng/mL. The authors suggested that the drop in OS was not atributted to the elevated iron level but rather the underlying disease that causes the iron level elevation. If this were true, a threshold of 1000ng/mL for SF is a better choice since it allows us to better capture the individuals at risk. 
- Lee et al 2009:
  - Retrospective study, n = 101
  - Diseases: ALL, AML, MDS, aplastic anemia AA. 
  - SF_threshold = 1000 ng/mL: significant association between elevated SF and OS, DFS, and TMR. 
  - Had a third group: patients whose SF dropped to below 1000ng/mL after iron-chelating therapy (IC group)
- Wermke et al 2012:
  - n = 88
  - Diseases: AML and MDS
  - Significant correlation between SF and LIC (r = 0.782). 
  - SF_threshold = 2500 ng/mL and LIC_threshold = 125ng/mL: the results are very interesting: 
    - Elevated LIC but not ferritin is significantly associated with lower OS and NRM. 
- Important covariates:
  - History of transplant: we would not have this info if the patients changed their care sites. However, chart reviewers should have caught this? Is it in the cohort eligibility? 

# Overall survival

- match status: Haploidentical, matched related, matched unrelated, mismatched related and mismatched unrelated:
    - matched related: match_status = 8/8 or 10/10 AND donor_relation = "Related"
    - matched unrelated: match_status = 8/8 or 10/10 AND donor_relation = "Unrelated"
    - mismatched related: match_status = 8/10, 9/10, or 7/8 AND donor_relation = "Related"
    - mismatched unrelated: match_status = 8/10, 9/10, or 7/8 AND donor_relation = "Unrelated"
    - haploididentical (related)
    - autologous/genetherapy
    - Allogeneic: Cord Blood
    - Unknown

- We decided to group some of the categories. The new categories are:
  - matched related: match_status = 8/8 or 10/10 AND donor_relation = "Related"
  - autologous/genetherapy
  - matched unrelated/mismatched related/mismatched unrelated/haploididentical/cord blood
  - Unknown: either match_status or donor_relation is unknown

## Aim 2a-2: Overall survival stratified by <span style="color:red">pre-transplant ferritin levels</span>
```{r cohort_covars_survival_analysis, echo= FALSE, message=FALSE}
stat_dataset <- results_tbl("survival_analysis_dataset") %>% collect() %>%
                  mutate(transplant_type_combined = relevel(factor(transplant_type_combined), ref = "matched related"))
                  # , levels = c("matched related", "matched unrelated", "mismatched related", "mismatched unrelated", "Haploidentical", "Allogeneic: Cord Blood", "autologous/genetherapy", "Unknown")))

ferritin_data <- results_tbl("cr_ferritin_data") %>%
                    select(-no_ferritin_pre, - no_ferritin_post) %>% collect()

stat_dataset_ferritin <-  stat_dataset %>%
            left_join(ferritin_data, by = c("person_id", "transplant_date"))                    
          
aim_2a_2 <- stat_dataset_ferritin %>% filter(chart_completion, eligibility == "Yes", !is.na(ferritin_pre)) %>%
                  mutate(ferritin_pre = factor(ferritin_pre, levels = c("low", "moderate", "high")),
                        ferritin_pre_max = factor(ferritin_pre_max, levels = c("low", "moderate", "high"))) 

auto_ct <- aim_2a_2 %>% filter(!is.na(survival_time_5yr_censored)) %>% filter(transplant_type_combined == "autologous/genetherapy") %>% nrow()
unknown_ct <- aim_2a_2 %>% filter(!is.na(survival_time_5yr_censored)) %>% filter(transplant_type_combined == "Unknown") %>% nrow()

pre_ferritin_cox <- aim_2a_2 %>% filter(!is.na(survival_time_5yr_censored)) %>%
                          select(record_id, ferritin_pre, age_at_ce, gender, graft_manip, transplant_type, scd_type, disease_relapse, graft_fail_date, donor_relation, survival_time_5yr_censored, survival_5yr, second_transplant, transplant_type_combined, graft_fail_5yr, graft_fail_days_since_ce_5yr_censored, has_busulfan, vod, bacteremia, bacteremia_time_1yr_censored)

label(pre_ferritin_cox$transplant_type_combined) = "match status"
table1(~gender + disease_relapse + age_at_ce + transplant_type_combined + ferritin_pre + has_busulfan + scd_type| survival_5yr, 
        data = pre_ferritin_cox %>% mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")), 
        overall = TRUE, 
        caption = "Overall survival at 5-yr following transplant")                          
```

### KM curves:
```{r os_KM_pre_ferritin}
km_ferritin_os <- survfit(Surv(survival_time_5yr_censored/365.25, survival_5yr) ~ ferritin_pre,   
                            data = pre_ferritin_cox %>%
                            mutate(ferritin_pre = factor(ferritin_pre, levels = c("low", "moderate", "high"))))

print(km_ferritin_os)

ggsurvplot(km_ferritin_os,
          pval = TRUE, conf.int = FALSE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          xlab = "Time (Years)",   # customize X axis label.
          ylab = "Overall Survival", 
          title = "Kaplan-Meier curves for overall survival using pre-transplant ferritin levels",
          legend.labs = c("Low", "Moderate", "High"),    # change legend labels.
          palette = c("red", "green", "blue"))
```

### Multivariate Cox regression analysis: 
- Multivariate Cox proportional hazard model adjusted for: 
  - pre-transplant most recent ferritin levels: low(<1000 ng/mL) (ref), moderate (1000-2500 ng/mL), or high (>=2500 ng/mL)
  - age at transplant: continuous 
  - gender: female (ref) or male
  - match status: 
    - Allogeneic: Cord Blood, Matched related (ref), Matched unrelated/Mismatched related/Mismatched Unrelated, & Unknown
    - Removed: autologous/genetherapy (see table above) due to missing obversation in outcomes. 
  - has busulfan: yes or no(ref)

- **Results**:
  - Increasing age is associated with a lower survival probability (Nora, does this make sense???). Although this association is statistically significant, the clinical impact appears to be minimal. It might be more informative to analyze this by breaking it down into age groups.
  - Being treated with bulsufan almost double the risk of death compared to patients not treated with busulfan. 
  - Another significant factor is matched unrelated/mismatched unrelated/mismatched related transplant type increased the risk of death by almost 130% compared to match related patients. 
  - Even though not significant, having a high pre-transplant ferritin level almost double of risk of death compared to having a low ferritin level. 

```{r overall_surv_pre_ferritin, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
# pre_ferritin_cox %>% group_by(survival_5yr, transplant_type_combined) %>% summarise(n = n_distinct(record_id)) %>% view()
pre_ferritin_cox_1 <- pre_ferritin_cox %>% filter(!(transplant_type_combined %in% c("autologous/genetherapy"))) %>%
                          mutate(transplant_type_combined = relevel(transplant_type_combined, ref = "matched related"))

# 0 is censored/survived, 1 is death
# exclude match_status and graft_manip
cox_ferritin_os <- coxph(Surv(survival_time_5yr_censored/365.25, survival_5yr) ~ ferritin_pre + age_at_ce + gender + transplant_type_combined + has_busulfan, 
                data = pre_ferritin_cox_1)

# pre_ferritin_cox_1 %>% group_by(ferritin_pre, gender, transplant_type_combined, has_busulfan) %>%
#                       summarise(n_distinct(record_id))
# visualiza the impact on iron levels
ferritin_levels_df <- with(pre_ferritin_cox_1,
               data.frame(ferritin_pre = c("low", "moderate", "high"), 
                          age_at_ce = rep(mean(age_at_ce, na.rm = TRUE), 3),
                          gender = rep("Female", 3),
                          transplant_type_combined = rep("matched related",3),
                          has_busulfan = rep("No", 3)
                          # match_status = rep("Cord Blood", 3), 
                          # disease_relapse = rep("No", 3), 
                          # graft_manip = rep("No", 3), 
                          # scd_type = rep("beta thalassemia major", 3)
                          )
               )

plot_title_text_size <- 20
axis_text_size <- 14

fit <- survfit(cox_ferritin_os, newdata = ferritin_levels_df)
ggsurvplot(fit, data = ferritin_levels_df, 
                  conf.int = FALSE, 
                  legend.labs = c("Low", "Moderate", "High"),
                  palette = c("red", "green", "blue"), # custom color palette
                  legend.title = "Pre-transplant ferritin levels", 
                  xlab = "Years", ylab = "Overall survival probability",
                  xlim = c(0, 5), ylim = c(0.7, 1.0), 
                  risk.table = TRUE,
                  ggtheme = theme(
    plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size))) #, risk.table.title = "test")

# output overall survival curves
# survfit2(res.cox) %>% ggsurvfit() +
#   labs(
#     x = "Years",
#     y = "Overall survival probability",
#     title = "Overall survival analysis pre-transplant ferritin"
#     ) + 
#   add_confidence_interval() +
#   add_risktable()
```

```{r overall_surv_pre_ferritin_sum, echo= FALSE, message=FALSE}
# summary(res.cox)
cox_sum(cox_ferritin_os) %>% prettify_kable()
```

## Aim 2a-2: Overall survival stratified by <span style="color:red">pre-transplant ferritin levels with binary low-moderate and high levels</span>
### KM curves:
```{r os_KM_pre_ferritin_binary}
km_ferritin_os_binary <- survfit(Surv(survival_time_5yr_censored/365.25, survival_5yr) ~ ferritin_pre,   
                    data = pre_ferritin_cox %>% 
                    mutate(ferritin_pre = if_else(ferritin_pre %in% c("low", "moderate"), "low-moderate", "high")) %>%
                    mutate(ferritin_pre = factor(ferritin_pre, levels = c("low-moderate", "high"))))

ggsurvplot(km_ferritin_os_binary,
          pval = TRUE, conf.int = FALSE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          xlab = "Time (Years)",   # customize X axis label.
          ylab = "Overall Survival", 
          title = "Kaplan-Meier curves for overall survival using pre-transplant ferritin levels",
          legend.labs = c("Low-moderate", "High"),    # change legend labels.
          palette = c("orange", "blue"))
```

### Multivariate Cox regression analysis: 
```{r overall_survival_pre_ferritin_binary, echo= FALSE, message=FALSE}
label(pre_ferritin_cox$transplant_type_combined) = "match status"
table1(~gender + disease_relapse + age_at_ce + transplant_type_combined + ferritin_pre + has_busulfan + scd_type| survival_5yr, 
        data = pre_ferritin_cox %>% 
              mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")) %>%
              mutate(ferritin_pre = if_else(ferritin_pre %in% c("low", "moderate"), "low-moderate", "high")),  
        overall = TRUE, 
        caption = "Overall survival at 5-yr following transplant")    

# 0 is censored/survived, 1 is death
# exclude match_status and graft_manip
cox_ferritin_os_binary <- coxph(Surv(survival_time_5yr_censored/365.25, survival_5yr) ~ ferritin_pre + age_at_ce + gender + transplant_type_combined + has_busulfan, 
                data = pre_ferritin_cox_1 %>%
                        mutate(ferritin_pre = if_else(ferritin_pre %in% c("low", "moderate"), "low-moderate", "high")))

ferritin_levels_df_binary <- with(pre_ferritin_cox_1 %>% mutate(ferritin_pre = if_else(ferritin_pre %in% c("low", "moderate"), "low-moderate", "high")),
               data.frame(ferritin_pre = c("low-moderate", "high"), 
                          age_at_ce = rep(mean(age_at_ce, na.rm = TRUE), 2),
                          gender = rep("Female", 2),
                          transplant_type_combined = rep("matched related",2),
                          has_busulfan = rep("No", 2)))

fit <- survfit(cox_ferritin_os_binary, newdata = ferritin_levels_df_binary)
ggsurvplot(fit, data = ferritin_levels_df_binary, 
                  conf.int = FALSE, 
                  legend.labs = c("Low-moderate", "High"),
                  palette = c("orange", "blue"), # custom color palette
                  legend.title = "Pre-transplant ferritin levels", 
                  xlab = "Years", ylab = "Overall survival probability",
                  xlim = c(0, 5), ylim = c(0.7, 1.0), 
                  risk.table = TRUE,
                  ggtheme = theme(
    plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size))) #, risk.table.title = "test")

cox_sum(cox_ferritin_os_binary) %>% prettify_kable()
```

## Aim 2a-1: Overall survival stratified by <span style="color:red">pre-transplant LIC levels</span>
```{r overall_surv_pre_LIC_data, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
LIC_data <- results_tbl("cr_LIC_data") %>% 
                  select(record_id, transplant_date, LIC, LIC_type, LIC_date, LIC_days_since_transplant, LIC_level, 
                  LIC_3, LIC_days_since_transplant_3, LIC_3_level, LIC_level_1_binary,
                  LIC_max, LIC_3_max, LIC_days_since_transplant_max, LIC_days_since_transplant_3_max,
                  LIC_max_level, LIC_3_max_level, LIC_level_3_binary, LIC_level_3_max_binary) %>%
                  collect()
                  
stat_dataset_LIC <- stat_dataset %>% left_join(LIC_data, by = c("record_id", "transplant_date"))

aim_2a_1 <- stat_dataset_LIC %>% filter(chart_completion, eligibility == "Yes", LIC_type == "pre-transplant", !is.na(LIC_level)) %>% 
                mutate(LIC_level_1_binary = factor(LIC_level_1_binary, levels = c("low", "high"))) %>%
                mutate(LIC_level_3_binary = factor(LIC_level_3_binary, levels = c("low", "high"))) %>%
                mutate(LIC_level_3_max_binary = factor(LIC_level_3_max_binary, levels = c("low", "high")))

pre_LIC_cox <- aim_2a_1 %>% filter(!is.na(survival_time_5yr_censored)) %>% 
                          select(record_id, LIC_level_3_binary, LIC_level_1_binary, LIC_level_3_max_binary, age_at_ce, gender, graft_manip, transplant_type, scd_type, disease_relapse, graft_fail_date, donor_relation, survival_time_5yr_censored, survival_5yr, second_transplant, transplant_type_combined, graft_fail_5yr, graft_fail_days_since_ce_5yr_censored, has_busulfan, site, vod, bacteremia, bacteremia_time_1yr_censored, LIC)
# pre_LIC_cox %>% group_by(survival_5yr, transplant_type_combined) %>% summarise(n = n_distinct(record_id)) %>% view()

auto_ct <- aim_2a_1 %>% filter(!is.na(survival_time_5yr_censored)) %>% filter(transplant_type_combined == "autologous/genetherapy") %>% nrow()
unknown_ct <- aim_2a_1 %>% filter(!is.na(survival_time_5yr_censored)) %>% filter(transplant_type_combined == "Unknown") %>% nrow()

label(pre_LIC_cox$transplant_type_combined) = "match status"
label(pre_LIC_cox$LIC_level_1_binary) = "pre-transplant LIC-level"
table1(~gender + disease_relapse + age_at_ce + transplant_type_combined + LIC_level_1_binary + has_busulfan + scd_type| survival_5yr, 
        data = pre_LIC_cox %>% mutate(survival_5yr = ifelse(survival_5yr == 1, "Dead", "Survived")), 
        overall = TRUE, 
        caption = "Overall survival at 5-yr following transplant")
```

### KM curves:
```{r os_KM_pre_LIC}
pre_LIC_cox_1 <- pre_LIC_cox %>% 
    filter(transplant_type_combined != "autologous/genetherapy")
    
km_LIC_os <- survfit(Surv(survival_time_5yr_censored/365.25, survival_5yr) ~ LIC_level_1_binary,   
                            data = pre_LIC_cox_1 %>%
                            mutate(LIC_level_1_binary = factor(LIC_level_1_binary, levels = c("low", "high"))))

# print(km_ferritin_os)

ggsurvplot(km_LIC_os,
          pval = TRUE, conf.int = FALSE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          xlab = "Time (Years)",   # customize X axis label.
          ylab = "Overall Survival", 
          title = "Kaplan-Meier curves for overall survival using pre-transplant LIC levels",
          legend.labs = c("Low", "High"),    # change legend labels.
          palette = c("red", "blue"))
```

### Multivariate Cox regression analysis: 
- Multivariate Cox proportional hazard model adjusted for: 
  - pre-transplant most recent LIC levels: low(<5 mg Fe/g dry wt) (ref), or high (>5 mg Fe/g dry wt)
  - age at transplant: continuous 
  - gender: female (ref) or male
  - match status: 
    - Allogeneic: Cord Blood, matched related, matched unrelated/mismatched related/mismatched unrelated, and unknown
    - Removed: autologous/genetherapy (see table above) due to missing obversation in outcomes. 
  - has busulfan: yes or no

- The results are very different from what we saw in OS for pre-transplant ferritin values:
  - The pre-transplant LIC level and being treated with busulfan do not have any impact on OS
  - Even though having a matched unrelated/mismatched unrelated/mismatched related transplant increases the risk of death by ~60%, this effect is nolonger significant. 

```{r overall_surv_pre_LIC, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
# 0 is censored/survived, 1 is death
# exclude match_status and graft_manip
cox_LIC_os <- coxph(Surv(survival_time_5yr_censored/365.25, survival_5yr) ~ LIC_level_1_binary + age_at_ce + gender + transplant_type_combined + has_busulfan, 
                data = pre_LIC_cox_1)

# visualiza the impact on iron levels
LIC_levels_df <- with(pre_LIC_cox_1,
               data.frame(LIC_level_1_binary = c("low", "high"), 
                          age_at_ce = rep(mean(age_at_ce, na.rm = TRUE), 2),
                          gender = rep("Female", 2),
                          transplant_type_combined = rep("matched related",2),
                          has_busulfan = rep("No", 2)
                          ))

fit <- survfit(cox_LIC_os, newdata = LIC_levels_df)
ggsurvplot(fit, data = LIC_levels_df, 
                  conf.int = FALSE, 
                  legend.labs = c("Low", "High"),
                  legend.title = "Pre-transplant LIC levels", 
                  palette = c("red", "blue"), # custom color palette
                  xlab = "Years", ylab = "Overall survival probability",
                  xlim = c(0, 5), ylim = c(0.7, 1.0),
                  ggtheme = theme(plot.title = element_text(size = plot_title_text_size),    # Title text size
                              axis.title.x = element_text(size = axis_text_size), # X-axis title size
                              axis.title.y = element_text(size = axis_text_size), # Y-axis title size
                              axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
                              axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
                              legend.title = element_text(size = axis_text_size),    # Legend title size
                              legend.text = element_text(size = axis_text_size)), 
                  risk.table = FALSE) #, risk.table.title = "test")

```

```{r overall_surv_pre_LIC_sum, echo= FALSE, message=FALSE}
# summary(res.cox)
cox_sum(cox_LIC_os) %>% prettify_kable()
```

# Graft failure
## Aim 2a-2: Graft failure stratified by <span style="color:red">pre-transplant ferritin levels</span>
```{r graft_fail_pre_ferritin_data, echo= FALSE, message=FALSE}
unknown_ct <- pre_ferritin_cox %>% filter(!is.na(graft_fail_days_since_ce_5yr_censored)) %>% filter(transplant_type_combined == "Unknown") %>% nrow()

label(pre_ferritin_cox$transplant_type_combined) = "match status"
label(pre_ferritin_cox$survival_5yr) = "Dies within 5 years of post-transplant"

table1(~gender + age_at_ce + transplant_type_combined + ferritin_pre + has_busulfan + survival_5yr + scd_type| graft_fail_5yr, 
        data = pre_ferritin_cox %>% mutate(graft_fail_5yr = ifelse(graft_fail_5yr ==1, "Graft failure", "Graft failure free")) %>%
        mutate(survival_5yr = ifelse(survival_5yr == 1, "No", "Yes")), 
        overall = FALSE, 
        caption = "Graft failure at 5-yr following transplant")      

table1(~ferritin_pre| graft_fail_5yr + scd_type, 
        data = pre_ferritin_cox %>% mutate(graft_fail_5yr = ifelse(graft_fail_5yr ==1, "Graft failure", "Graft failure free")) %>%
        mutate(survival_5yr = ifelse(survival_5yr == 1, "No", "Yes")), 
        overall = TRUE, 
        caption = "Graft failure 5-yr following transplant for subgroups")                          
```

- Competing risk analysis adjusted for: 
  - pre-transplant most recent ferritin levels: low(<1000 ng/mL), moderate (1000-2500 ng/mL), or high (>=2500 ng/mL)
  - age at transplant: continuous 
  - gender: female (ref) or male
  - match status: Allogeneic: matched related (ref), matched unrelated/mismatched related/mismatched unrelated/Cord Blood, autologous/genetherapy (removed), and Unknown
  - has busulfan: yes or no
  - Outcomes: graft failure or death without graft failure
  - There are a few choices for competeting risk model based on diferrent assumptions:
    - The Fine-Gray model (more commonly used) turn a multi-state model into collections of 2-state models by grouping the outcomes. The Fine-Gray model assumes that each covariate's effect on log(1-F) is a constant and independent of the other variables (F(t) = p(t) probability in state function). Linearity of continuous variables, no interactions, time-independent covariates, and proportional hazards. 
    - In this analysis we use the Aalen-Johansen estimates (multi-state models)
- Composite outcome: event-free survival: death/graft failure (whichever is first) 
- n = `r unknown_ct` patients with unknown match status were removed from the analysis

```{r graft_fail_pre_ferritin, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
# 0 is censored/survived, 1 is death
# exclude match_status and graft_manip
# km <- survfit(Surv(graft_fail_days_since_ce_5yr_censored/30, graft_fail_5yr) ~ ferritin_pre# + age_at_ce + gender + transplant_type_combined + has_busulfan, 
#                 data = pre_ferritin_cox_2)
pre_ferritin_cox_2 <- pre_ferritin_cox %>% 
                          filter(transplant_type_combined != "Unknown") %>% 
                          # filter(transplant_type_combined != "autologous/genetherapy") %>% 
                          filter(!is.na(graft_fail_5yr)) %>% 
                          mutate(graft_fail_5yr = recode(graft_fail_5yr, "0" = "event-free", "1" = "graft failure", "2" = "death")) %>%
                          mutate(graft_fail_5yr = factor(graft_fail_5yr,
                                                  levels = c("event-free", "graft failure", "death"))) %>% 
                          filter(graft_fail_days_since_ce_5yr_censored >= 0) %>%
                          mutate(transplant_type_combined = relevel(factor(transplant_type_combined), ref = "matched related"))
# km$transitions
# this plot function is especially for univariate analysis only
# cant do this for multivariate

# plot(km[, "graft failure"], 
#               mark.time = FALSE, lwd = 2, xscale = 12,
#               conf.int = FALSE, #lty = "dashed",
#               col = c("red", "blue", "black"), 
#               # linecolor = c("r", "b", "blue"),
#               main = "Graft failure probability --- Competing Risk Kaplan Meier univariate analysis", 
#               xlab = "Years post transplant", ylab = "Graft failure probability")
# legend(2*12, .02, c("low ferritin", "moderate ferritin", "high ferritin"), 
#       col = c("red", "blue", "black"), lty = "solid")

compRisk_ferritin_gf <- coxph(Surv(graft_fail_days_since_ce_5yr_censored/30, graft_fail_5yr) ~ ferritin_pre + age_at_ce + gender + has_busulfan  + transplant_type_combined, 
                data = pre_ferritin_cox_2, 
                id = record_id)

# fine gray model
fg_ferritin_gf <- finegray(Surv(graft_fail_days_since_ce_5yr_censored/30, graft_fail_5yr) ~ ferritin_pre +        + age_at_ce + gender + has_busulfan  + transplant_type_combined, 
                data = pre_ferritin_cox_2, 
                etype = "death")

# multi-state Aalen-Johansen model
aj_ferritin_gf <- survfit(Surv(graft_fail_days_since_ce_5yr_censored/30, graft_fail_5yr) ~ ferritin_pre +        + age_at_ce + gender + has_busulfan  + transplant_type_combined, 
                data = pre_ferritin_cox_2)

# plot(aj_ferritin_gf[, "graft failure"])
# print(res.cox, digits = 3)                
# ggsurvplot(res.cox)

# visualiza the impact on iron levels
# ferritin_levels_df <- with(pre_ferritin_cox_2,
#                data.frame(ferritin_pre = c("low", "moderate", "high"), 
#                           age_at_ce = rep(mean(age_at_ce, na.rm = TRUE), 3),
#                           gender = rep("Female", 3),
#                           transplant_type_combined = rep("matched related",3),
                          # has_busulfan = rep("No", 3)))

# fit <- survfit(res.cox, newdata = ferritin_levels_df)
# ggsurvplot(fit, data = ferritin_levels_df, 
#                   conf.int = FALSE, ggtheme = theme_minimal(),
#                   legend.labs = c("Low", "Moderate", "High"),
#                   legend.title = "Pre-transplant ferritin levels", 
#                   xlab = "Years", ylab = "Graft failure free probability",
#                   xlim = c(0, 5), ylim = c(0.5, 1.0), 
#                   risk.table = FALSE)#, risk.table.title = "test")
```



- Having a matched unrelated/mismatched unrelated/mismatched related transplant has a significant effect on graft failure. The risk of having graft failure increased by almost 3 folds comparing to having a matched related transplant. Gender is another significant factor, risk of graft failure in male patients is 2 and a haft times higher than females. 

- Patients with a moderate ferritin levels are 60% less likely to die without graft failure than patients with low ferritin level pre-transplant. Age is another signicant factor; however, the OR is 1.07 which indicates that older patients have higher risk of death without graft failure than younger patients (Nora, does it make sense?). Being treated with bulsufan increased the risk death without GF by almost 3 folds. No patients receiving augologous/gene therapy transplant dies without graft failure; thus, this coefficient could be ignored. 

- In the table below: 1:2 represents transition to Graft failure and 1:3 represents transition to death without graft failure
```{r graft_fail_pre_ferritin_sum, echo= FALSE, message=FALSE}
cox_sum(compRisk_ferritin_gf) %>% prettify_kable()
```

<span style="background-color: #FFFF00"> Nora: looking at this summary, it doesn't look like there is any association between transplant type and busulfan. </span>

```{r graft_failure_pre_ferritin_transplant_type_sum, message = FALSE}
table1(~ transplant_type_combined| has_busulfan, data = pre_ferritin_cox_2)
```

### visualize graft failure probability
- I'm plotting a few different ways of presenting the data and let me know which one you prefer
```{r graft_fail_pre_ferritin_plot}

# there are many packages available cmprisk, we are using the mstate package

# Prepare transition matrix
tmat <- trans.comprisk(2, names = c("event-free", "graft failure", "death"))

# Run msprep
pre_ferritin_cox_2$graft_fail <- as.numeric(pre_ferritin_cox_2$graft_fail_5yr == "graft failure")
pre_ferritin_cox_2$death <- as.numeric(pre_ferritin_cox_2$graft_fail_5yr == "death")

# convert to long data format
# the status variable is just a numeric representation of graft_fail_5yr 
# If there are K competing events, each individual needs K rows in the new data file, one for each
# possible cause of failure. A column (cause in the example) is used to denote the event type or
# failure cause that the row refers to.
# The advantage of long format is that we can simultaneously model all the 1:2 and 1:3 transitions
# instead of separating them in wide format data

pre_ferritin_cox_2_long <- msprep(
  time = c(NA, "graft_fail_days_since_ce_5yr_censored", "graft_fail_days_since_ce_5yr_censored"), 
  status = c(NA, "graft_fail", "death"), 
  data = pre_ferritin_cox_2, 
  keep = c("scd_type", "has_busulfan", "transplant_type_combined", "age_at_ce", "gender", "ferritin_pre"), 
  trans = tmat)

# expand the dataset to include type-specific covariates (if any)
# extra dummy variables ferritin_premoderate.1 + ferritin_prehigh.2 .. were added 
# They have the value 0 except for the ferritin levels and the transitions that they correspond to (i.e. for a # patient with moderate ferritin ferritin_premoderate.1=1 for the first cause, ‘graft-failure’, 
# ferritin_premoderate.2=1 for the second cause, 'death without graft failure'. They
# are what Andersen et al call type-specific covariates.
pre_ferritin_cox_2_expanded <- expand.covs(pre_ferritin_cox_2_long, "ferritin_pre")

# colnames(pre_ferritin_cox_2_expanded)

# pre_ferritin_cox_2_expanded %>% view()

# there are many ways of doing this
# i. having ferritin_pre on the right handside
# ii. having the expanded terms: ferritin_premoderate.1 + ferritin_prehigh.2 etc on the right handside
# iii. including strata(trans) to allow for different baseline hazard functions across the levels of the trans variable. This is particularly useful in multi-state models where the hazard of transitioning from one state to another may differ significantly across transitions
# iv. adding a covariate by cause interactions ccr5 * factor(trans)
# check out all of these senarios: if the interactions is significant, maybe we dont want to remove it

ferritin_cp <- coxph(Surv(time/365.25, status) ~ ferritin_premoderate.1 + #ferritin_pre +
                                  ferritin_premoderate.2 +
                                  ferritin_prehigh.1 + ferritin_prehigh.2 + 
                                  gender + 
                                  age_at_ce + gender + has_busulfan  + transplant_type_combined + 
                                  strata(trans),
            data = pre_ferritin_cox_2_expanded)

# Visualising cumulative baseline hazards using plot.msfit()
# Data to predict
# in this expanded dataset ferritin_premoderate.1 means "transition 1": event-free to graft-failure 

f_low <- data.frame(
  # ferritin_pre = c("low", "low"),
  ferritin_premoderate.1 = c(0, 0),
  ferritin_premoderate.2 = c(0, 0),
  ferritin_prehigh.1 = c(0, 0),
  ferritin_prehigh.2 = c(0, 0),
  age_at_ce = c(10, 10), 
  gender = c("Female", "Female"),
  has_busulfan = c("No", "No"),  
  transplant_type_combined = c("matched related", "matched related"),
  trans = c(1, 2), 
  strata = c(1, 2)
)

f_moderate <- data.frame(
  # ferritin_pre = c("moderate", "moderate"),
  ferritin_premoderate.1 = c(1, 0),
  ferritin_premoderate.2 = c(0, 1),
  ferritin_prehigh.1 = c(0, 0),
  ferritin_prehigh.2 = c(0, 0),
  age_at_ce = c(10, 10), 
  gender = c("Female", "Female"),
  has_busulfan = c("No", "No"),  
  transplant_type_combined = c("matched related", "matched related"),
  trans = c(1, 2), 
  strata = c(1, 2)
)

f_high <- data.frame(
  # ferritin_pre = c("high", "high"),
  ferritin_premoderate.1 = c(0, 0),
  ferritin_premoderate.2 = c(0, 0),
  ferritin_prehigh.1 = c(1, 0),
  ferritin_prehigh.2 = c(0, 1),
  age_at_ce = c(10, 10), 
  gender = c("Female", "Female"),
  has_busulfan = c("No", "No"),  
  transplant_type_combined = c("matched related", "matched related"),
  trans = c(1, 2), 
  strata = c(1, 2)
)

# Make msfit object for each ferritin level
msf.f_low <- msfit(object = ferritin_cp, newdata = f_low, trans = tmat)
msf.f_moderate <- msfit(object = ferritin_cp, newdata = f_moderate, trans = tmat)
msf.f_high <- msfit(object = ferritin_cp, newdata = f_high, trans = tmat)

# Make probtrans objects for each ferritin level
pt.f_low <- probtrans(msf.f_low, predt = 0)
pt.f_moderate <- probtrans(msf.f_moderate, predt = 0)
pt.f_high <- probtrans(msf.f_high, predt = 0)
# if just plotting a single level, replace the c() dataset with a single one_of()
# or call plot() instead of vis.multiple.pt()

vis.multiple.pt(
  x = list(pt.f_low, pt.f_moderate, pt.f_high), 
  from = 1,
  to = 2, 
  conf.type = "none",
  # cols = c(1, 2, 3),
  cols = c("blue", "red", "green"), #  something is wrong with this argument
  labels = c("low", "moderate", "high"),
  legend.title = "Pre-transplant ferritin",
  ylab = "Probability of graft failure",
  xlim = c(0, 5), 
  ylim = c(0, .1)) +
  labs(title = "Graft failure") +
  theme(plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size)) 

vis.multiple.pt(
  x = list(pt.f_low, pt.f_moderate, pt.f_high), 
  from = 1,
  to = 3, 
  cols = c("blue", "red", "green"), #  something is wrong with this argument
  labels = c("low", "moderate", "high"),
  legend.title = "Pre-transplant ferritin",
  conf.type = "none",
  ylab = "Probability of death without graft failure",
  xlim = c(0, 5), 
  ylim = c(0, .1)) + 
  labs(title = "Death without graft failure") +
  theme(plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size))

plot(msf.f_low, 
    xlim = c(0, 5), 
    col = c("black", "grey"), 
    lty = c("dashed", "solid"), 
    legend.pos = "top",
    lwd = 2,
    use.ggplot = TRUE,
    type = "single", #for 2 boxes
    scale_type = "fixed",
    ylim = c(0, .1),
    xlab = "Years post transplant",
  ) + # Add title and center
ggtitle("Cumulative baseline hazards for competing risk model for graft-failure") +
theme(plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size))
```

### subgroup analysis:
#### aplastic anemia
- We can't do the subgroup analysis for aplastic anemia unless we ignore busulfan covariate
```{r pre_ferritin_gf_subgroup_aa}
table1(~gender + age_at_ce + transplant_type_combined + ferritin_pre + has_busulfan + survival_5yr + scd_type| graft_fail_5yr, 
        data = pre_ferritin_cox %>% 
              mutate(graft_fail_5yr = ifelse(graft_fail_5yr ==1, "Graft failure", "Graft failure free")) %>%
              mutate(survival_5yr = ifelse(survival_5yr == 1, "No", "Yes")) %>%
              filter(scd_type == "aplastic anemia"),
        overall = FALSE, 
        caption = "Graft failure at 5-yr following transplant")
```

#### beta thalassemia major
- We can't do the subgroup analysis unless we ignore busulfan covariate
```{r pre_ferritin_gf_subgroup_btm}

table1(~gender + age_at_ce + transplant_type_combined + ferritin_pre + has_busulfan + survival_5yr + scd_type| graft_fail_5yr, 
        data = pre_ferritin_cox %>% 
              mutate(graft_fail_5yr = ifelse(graft_fail_5yr ==1, "Graft failure", "Graft failure free")) %>%
              mutate(survival_5yr = ifelse(survival_5yr == 1, "No", "Yes")) %>%
              filter(scd_type == "beta thalassemia major"),
        overall = FALSE, 
        caption = "Graft failure at 5-yr following transplant")

pre_ferritin_df_btm <- pre_ferritin_cox %>% 
                          filter(scd_type == "beta thalassemia major") %>%
                          filter(transplant_type_combined != "Unknown") %>% 
                          filter(!is.na(graft_fail_5yr)) %>% 
                          mutate(graft_fail_5yr = recode(graft_fail_5yr, "0" = "event-free", "1" = "graft failure", "2" = "death")) %>%
                          mutate(graft_fail_5yr = factor(graft_fail_5yr,
                                                  levels = c("event-free", "graft failure", "death"))) %>% 
                          filter(graft_fail_days_since_ce_5yr_censored >= 0) %>%
                          mutate(transplant_type_combined = relevel(factor(transplant_type_combined), ref = "matched related"))

compRisk_ferritin_gf_btm <- coxph(Surv(graft_fail_days_since_ce_5yr_censored/30, graft_fail_5yr) ~ ferritin_pre + age_at_ce + gender + has_busulfan  + transplant_type_combined, 
                data = pre_ferritin_df_btm, 
                id = record_id)

cox_sum(compRisk_ferritin_gf_btm) %>% prettify_kable()
```

#### sickle cell disease
- Because the sample sizes for some categories are very small. I did this analysis just for completion. We should intepret the results carefully. 
```{r pre_ferritin_gf_subgroup_scd}
table1(~gender + age_at_ce + transplant_type_combined + ferritin_pre + has_busulfan + survival_5yr + scd_type| graft_fail_5yr, 
        data = pre_ferritin_cox %>% 
              mutate(graft_fail_5yr = ifelse(graft_fail_5yr ==1, "Graft failure", "Graft failure free")) %>%
              mutate(survival_5yr = ifelse(survival_5yr == 1, "No", "Yes")) %>%
              filter(scd_type == "sickle cell disease"), 
        overall = FALSE, 
        caption = "Graft failure at 5-yr following transplant")

pre_ferritin_df_scd <- pre_ferritin_cox %>% 
                          filter(scd_type == "sickle cell disease") %>%
                          filter(transplant_type_combined %in% c("matched related", "matched unrelated/mismatched unrelated/mismatched related/cord blood")) %>%
                          filter(!is.na(graft_fail_5yr)) %>% 
                          mutate(graft_fail_5yr = recode(graft_fail_5yr, "0" = "event-free", "1" = "graft failure", "2" = "death")) %>%
                          mutate(graft_fail_5yr = factor(graft_fail_5yr,
                                                  levels = c("event-free", "graft failure", "death"))) %>% 
                          filter(graft_fail_days_since_ce_5yr_censored >= 0) %>%
                          mutate(transplant_type_combined = relevel(factor(transplant_type_combined), ref = "matched related"))

compRisk_ferritin_gf_scd <- coxph(Surv(graft_fail_days_since_ce_5yr_censored/30, graft_fail_5yr) ~ ferritin_pre + age_at_ce + gender + has_busulfan  + transplant_type_combined, 
                data = pre_ferritin_df_scd, 
                id = record_id)

cox_sum(compRisk_ferritin_gf_scd) %>% prettify_kable()
```

#### Diamond-Blackfan anemia
- We can't do the subgroup analysis for this due to the small sample size
```{r pre_ferritin_gf_subgroup_dba}
table1(~gender + age_at_ce + transplant_type_combined + ferritin_pre + has_busulfan + survival_5yr + scd_type| graft_fail_5yr, 
        data = pre_ferritin_cox %>% 
              mutate(graft_fail_5yr = ifelse(graft_fail_5yr ==1, "Graft failure", "Graft failure free")) %>%
              mutate(survival_5yr = ifelse(survival_5yr == 1, "No", "Yes")) %>%
              filter(scd_type == "Diamond-Blackfan anemia"), 
        overall = FALSE, 
        caption = "Graft failure at 5-yr following transplant")
```

## Aim 2a-1: Graft failure stratified by <span style="color:red">pre-transplant LIC levels</span>
```{r graft_fail_pre_LIC_data, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
# pre_LIC_cox %>% group_by(survival_5yr, transplant_type_combined) %>% summarise(n = n_distinct(record_id)) %>% view()
cord_blood_ct <- aim_2a_1 %>% filter(!is.na(graft_fail_days_since_ce_5yr_censored)) %>% filter(transplant_type_combined == "Allogeneic: Cord Blood") %>% nrow()
unknown_ct <- aim_2a_1 %>% filter(!is.na(graft_fail_days_since_ce_5yr_censored)) %>% filter(transplant_type_combined == "Unknown") %>% nrow()

label(pre_LIC_cox$transplant_type_combined) = "match status"
label(pre_LIC_cox$LIC_level_1_binary) = "pre-transplant LIC-level"
label(pre_LIC_cox$survival_5yr) = "Dies within 5 years of post-transplant"

table1(~gender + age_at_ce + transplant_type_combined + LIC_level_1_binary + has_busulfan + survival_5yr + scd_type| graft_fail_5yr, 
        data = pre_LIC_cox %>% 
                  mutate(graft_fail_5yr = ifelse(graft_fail_5yr ==1, "Graft failure", "Graft failure free")) %>%
                  mutate(survival_5yr = ifelse(survival_5yr == 1, "No", "Yes")), 
        overall = TRUE, 
        caption = "Graft failure at 5-yr following transplant")

table1(~LIC_level_1_binary| graft_fail_5yr + scd_type,
      data = pre_LIC_cox %>% 
                  mutate(graft_fail_5yr = ifelse(graft_fail_5yr ==1, "Graft failure", "Graft failure free")) %>%
                  mutate(survival_5yr = ifelse(survival_5yr == 1, "No", "Yes")),
      overall = FALSE, 
      caption = "Graft failure 5-yr following transplant for subgroups")
```

- Competing risk analysis adjusted for: 
  - pre-transplant most recent LIC levels: low(<5 mg Fe/g dry wt), or high (>5 mg Fe/g dry wt)
  - age at transplant: continuous 
  - gender: female (ref) or male
  - match status: Allogeneic: Cord Blood, matched related (ref), matched unrelated/mismatched related/mismatched unrelated, autologous/genetherapy, and Unknown
  - has busulfan: yes or no
  - Outcomes: graft failure or death without graft failure 
- All patients with cord blood transplant survived (n = `r cord_blood_ct`). 
- n = `r unknown_ct` patients with unknown match status

- To do: Composite outcome: event-free survival: death/graft failure (whichever is first) 

```{r graft_fail_pre_LIC, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
pre_LIC_cox_2 <- pre_LIC_cox %>% 
                      filter(!(transplant_type_combined %in% c("Allogeneic: Cord Blood", "Unknown"))) %>%
                      # filter(transplant_type_combined != "autologous/genetherapy") %>% 
                      filter(!is.na(graft_fail_5yr)) %>% 
                      mutate(graft_fail_5yr = recode(graft_fail_5yr, "0" = "censor", "1" = "graft failure", "2" = "death")) %>%
                      mutate(graft_fail_5yr = factor(graft_fail_5yr,
                                              levels = c("censor", "graft failure", "death"))) %>% 
                      filter(graft_fail_days_since_ce_5yr_censored >= 0) %>%
                      mutate(transplant_type_combined = relevel(factor(transplant_type_combined), ref = "matched related"))

# 0 is censored/survived, 1 is death
# exclude match_status and graft_manip
compRisk_LIC_gf <- coxph(Surv(graft_fail_days_since_ce_5yr_censored/365.25, graft_fail_5yr) ~ LIC_level_1_binary + age_at_ce + gender + transplant_type_combined + has_busulfan, 
                data = pre_LIC_cox_2, 
                id = record_id)

```

- The effects using pre-transplant LIC are similar to that of ferritin, having a matched unrelated/mismatched unrelated/mismatched related transplant increased the risk of graft failure by almost 4 folds comparing to having a matched related transplant. Gender is another significant factor, risk of graft failure in male patients is almost 3 times higher than females. 

- For death without graft failure, none of the effect is significant (no patients receiving augologous/gene therapy transplant dies without graft failure; thus, this coefficient could be ignored) 

- In the table below: 1:2 represents transition to Graft failure and 1:3 represents transition to death without graft failure
- <span style="background-color: #FFFF00"> Nora, patients with high LIC level are less likely to get graft failure, is it because these patients are more likely to receive bulsufan? </span>

```{r graft_fail_pre_LIC_sum, echo= FALSE, message=FALSE}
# summary(res.cox)
cox_sum(compRisk_LIC_gf) %>% prettify_kable()
```

<span style="background-color: #FFFF00"> Nora, looking at this summary, it doesn't look like there is any association between transplant type and busulfan. </span>

```{r graft_failure_pre_LIC_transplant_type_sum, message = FALSE}
table1(~ transplant_type_combined| has_busulfan, data = pre_LIC_cox_2)
```

### visualize graft failure probability
- I'm plotting a few different ways of presenting the data and let me know which one you prefer
```{r graft_fail_pre_LIC_plot}

# there are many packages available cmprisk, we are using the mstate package

# Prepare transition matrix
tmat <- trans.comprisk(2, names = c("event-free", "graft failure", "death"))

# Run msprep
pre_LIC_cox_2$graft_fail <- as.numeric(pre_LIC_cox_2$graft_fail_5yr == "graft failure")
pre_LIC_cox_2$death <- as.numeric(pre_LIC_cox_2$graft_fail_5yr == "death")

# convert to long data format
# the status variable is just a numeric representation of graft_fail_5yr 
# If there are K competing events, each individual needs K rows in the new data file, one for each
# possible cause of failure. A column (cause in the example) is used to denote the event type or
# failure cause that the row refers to.
# The advantage of long format is that we can simultaneously model all the 1:2 and 1:3 transitions
# instead of separating them in wide format data

pre_LIC_cox_2_long <- msprep(
  time = c(NA, "graft_fail_days_since_ce_5yr_censored", "graft_fail_days_since_ce_5yr_censored"), 
  status = c(NA, "graft_fail", "death"), 
  data = pre_LIC_cox_2, 
  keep = c("scd_type", "has_busulfan", "transplant_type_combined", "age_at_ce", "gender",  "LIC_level_1_binary" ), 
  trans = tmat)

# expand the dataset to include type-specific covariates (if any)
# extra dummy variables ferritin_premoderate.1 + ferritin_prehigh.2 .. were added 
# They have the value 0 except for the ferritin levels and the transitions that they correspond to (i.e. for a # patient with moderate ferritin ferritin_premoderate.1=1 for the first cause, ‘graft-failure’, 
# ferritin_premoderate.2=1 for the second cause, 'death without graft failure'. They
# are what Andersen et al call type-specific covariates.
pre_LIC_cox_2_expanded <- expand.covs(pre_LIC_cox_2_long,  "LIC_level_1_binary")

# colnames(pre_LIC_cox_2_expanded)

# pre_LIC_cox_2_expanded %>% view()

# there are many ways of doing this
# i. having ferritin_pre on the right handside
# ii. having the expanded terms: ferritin_premoderate.1 + ferritin_prehigh.2 etc on the right handside
# iii. including strata(trans) to allow for different baseline hazard functions across the levels of the trans variable. This is particularly useful in multi-state models where the hazard of transitioning from one state to another may differ significantly across transitions
# iv. adding a covariate by cause interactions ccr5 * factor(trans)
# check out all of these senarios: if the interactions is significant, maybe we dont want to remove it

LIC_cp <- coxph(Surv(time/365.25, status) ~ LIC_level_1_binaryhigh.1 +
                                  LIC_level_1_binaryhigh.2 + 
                                  gender + 
                                  age_at_ce + gender + has_busulfan  + transplant_type_combined + 
                                  strata(trans),
            data = pre_LIC_cox_2_expanded)

# Visualising cumulative baseline hazards using plot.msfit()
# Data to predict
# in this expanded dataset ferritin_premoderate.1 means "transition 1": event-free to graft-failure 

LIC_low <- data.frame(
  LIC_level_1_binaryhigh.1 = c(0, 0),
  LIC_level_1_binaryhigh.2 = c(0, 0),
  age_at_ce = c(10, 10), 
  gender = c("Female", "Female"),
  has_busulfan = c("No", "No"),  
  transplant_type_combined = c("matched related", "matched related"),
  trans = c(1, 2), 
  strata = c(1, 2)
)

LIC_high <- data.frame(
  LIC_level_1_binaryhigh.1 = c(1, 0),
  LIC_level_1_binaryhigh.2 = c(0, 1),
  age_at_ce = c(10, 10), 
  gender = c("Female", "Female"),
  has_busulfan = c("No", "No"),  
  transplant_type_combined = c("matched related", "matched related"),
  trans = c(1, 2), 
  strata = c(1, 2)
)

# Make msfit object for each ferritin level
msf.LIC_low <- msfit(object = LIC_cp, newdata = LIC_low, trans = tmat)
msf.LIC_high <- msfit(object = LIC_cp, newdata = LIC_high, trans = tmat)

# Make probtrans objects for each ferritin level
pt.LIC_low <- probtrans(msf.LIC_low, predt = 0)
pt.LIC_high <- probtrans(msf.LIC_high, predt = 0)
# if just plotting a single level, replace the c() dataset with a single one_of()
# or call plot() instead of vis.multiple.pt()

vis.multiple.pt(
  x = list(pt.LIC_low, pt.LIC_high), 
  from = 1,
  to = 2, 
  conf.type = "none",
  cols = c("red", "blue"), 
  labels = c("low", "high"),
  legend.title = "Pre-transplant LIC levels",
  ylab = "Probability of graft failure",
  xlim = c(0, 5), 
  ylim = c(0, .1)) +
  labs(title = "Graft failure") +
  theme(plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size)) 

vis.multiple.pt(
  x = list(pt.LIC_low, pt.LIC_high), 
  from = 1,
  to = 3, 
  conf.type = "none",
  cols = c("red", "blue"), 
  labels = c("low", "high"),
  legend.title = "Pre-transplant LIC levels",
  ylab = "Probability of death without graft failure",
  xlim = c(0, 5), 
  ylim = c(0, .1)) +
  labs(title = "Death without graft failure") +
  theme(plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size)) 

plot(msf.LIC_low, 
    xlim = c(0, 5), 
    col = c("black", "grey"), 
    lty = c("dashed", "solid"), 
    legend.pos = "top",
    lwd = 2,
    use.ggplot = TRUE,
    type = "single", #for 2 boxes
    scale_type = "fixed",
    ylim = c(0, .1),
    xlab = "Years post transplant",
  ) + # Add title and center
ggtitle("Cumulative baseline hazards using pre-transplant LIC levels") +
theme(plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size))
```

### subgroup analysis:
#### aplastic anemia
- We can't do the subgroup analysis for aplastic anemia due to the missing patients in some categories
```{r pre_LIC_gf_subgroup_aa}
table1(~gender + age_at_ce + transplant_type_combined + LIC_level_1_binary + has_busulfan + survival_5yr + scd_type| graft_fail_5yr, 
        data = pre_LIC_cox %>% 
                  mutate(graft_fail_5yr = ifelse(graft_fail_5yr ==1, "Graft failure", "Graft failure free")) %>%
                  mutate(survival_5yr = ifelse(survival_5yr == 1, "No", "Yes")) %>%
                  filter(scd_type == "aplastic anemia"), 
        overall = TRUE, 
        caption = "Graft failure at 5-yr following transplant")
```

#### beta thalassemia major
- We can't do the subgroup analysis unless we ignore busulfan covariate
```{r pre_LIC_gf_subgroup_btm}
table1(~gender + age_at_ce + transplant_type_combined + LIC_level_1_binary + has_busulfan + survival_5yr + scd_type| graft_fail_5yr, 
        data = pre_LIC_cox %>% 
                  mutate(graft_fail_5yr = ifelse(graft_fail_5yr ==1, "Graft failure", "Graft failure free")) %>%
                  mutate(survival_5yr = ifelse(survival_5yr == 1, "No", "Yes")) %>%
                  filter(scd_type == "beta thalassemia major"), 
        overall = TRUE, 
        caption = "Graft failure at 5-yr following transplant")

pre_LIC_df_btm <- pre_LIC_cox %>% 
                          filter(scd_type == "beta thalassemia major") %>%
                          filter(transplant_type_combined != "Unknown") %>% 
                          filter(!is.na(graft_fail_5yr)) %>% 
                          mutate(graft_fail_5yr = recode(graft_fail_5yr, "0" = "event-free", "1" = "graft failure", "2" = "death")) %>%
                          mutate(graft_fail_5yr = factor(graft_fail_5yr,
                                                  levels = c("event-free", "graft failure", "death"))) %>% 
                          filter(graft_fail_days_since_ce_5yr_censored >= 0) %>%
                          mutate(transplant_type_combined = relevel(factor(transplant_type_combined), ref = "matched related"))

compRisk_LIC_gf_btm <- coxph(Surv(graft_fail_days_since_ce_5yr_censored/30, graft_fail_5yr) ~ LIC_level_1_binary + age_at_ce + gender + has_busulfan  + transplant_type_combined, 
                data = pre_LIC_df_btm, 
                id = record_id)

cox_sum(compRisk_LIC_gf_btm) %>% prettify_kable()
```

#### sickle cell disease
- We can't do the subgroup analysis due to the missing patients in some categories. 
```{r pre_LIC_gf_subgroup_scd}
table1(~gender + age_at_ce + transplant_type_combined + LIC_level_1_binary + has_busulfan + survival_5yr + scd_type| graft_fail_5yr, 
        data = pre_LIC_cox %>% 
                  mutate(graft_fail_5yr = ifelse(graft_fail_5yr ==1, "Graft failure", "Graft failure free")) %>%
                  mutate(survival_5yr = ifelse(survival_5yr == 1, "No", "Yes")) %>%
                  filter(scd_type == "sickle cell disease"), 
        overall = TRUE, 
        caption = "Graft failure at 5-yr following transplant")
```

#### Diamond-Blackfan anemia
- We can't do the subgroup analysis because there is no patients experienced graft failure in this subgroup. 
```{r pre_LIC_gf_subgroup_dba}
table1(~gender + age_at_ce + transplant_type_combined + LIC_level_1_binary + has_busulfan + survival_5yr + scd_type| graft_fail_5yr, 
        data = pre_LIC_cox %>% 
                  mutate(graft_fail_5yr = ifelse(graft_fail_5yr ==1, "Graft failure", "Graft failure free")) %>%
                  mutate(survival_5yr = ifelse(survival_5yr == 1, "No", "Yes")) %>%
              filter(scd_type == "Diamond-Blackfan anemia"), 
        overall = TRUE, 
        caption = "Graft failure at 5-yr following transplant")

```


# Bacteremia
## Bacteremia stratified by <span style="color:red">pre-transplant ferritin levels</span>
### KM curves:
```{r bacteremia_pre_ferritin_km}
pre_ferritin_bac <- pre_ferritin_cox %>% #filter(!(transplant_type_combined %in% c("autologous/genetherapy"))) %>%
                          # mutate(transplant_type_combined = relevel(transplant_type_combined, ref = "matched related"))
                          mutate(bacteremia = ifelse(is.na(bacteremia), FALSE, bacteremia)) %>%
                          mutate(ferritin_pre = factor(ferritin_pre, levels = c("low", "moderate", "high")))

table1(~ ferritin_pre + age_at_ce + gender + transplant_type_combined| bacteremia, 
      data = pre_ferritin_bac %>% mutate(bacteremia = ifelse(!bacteremia, "No bacteremia", "Bacteremia")))

table1(~ ferritin_pre | bacteremia + scd_type, 
      data = pre_ferritin_bac %>% mutate(bacteremia = ifelse(!bacteremia, "No bacteremia", "Bacteremia")),
      overall = FALSE)

km_ferritin_bac <- survfit(Surv(bacteremia_time_1yr_censored/365.25, bacteremia) ~ ferritin_pre,   
                    data = pre_ferritin_bac)

ggsurvplot(km_ferritin_bac,
          pval = TRUE, conf.int = FALSE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          xlab = "Time (Years)",   # customize X axis label.
          ylab = "Overall Survival", 
          title = "Kaplan-Meier curves for bacteremia probability using pre-transplant ferritin levels",
          legend.labs = c("Low", "Moderate", "High"),    # change legend labels.
          palette = c("red", "green", "blue"))
```

### Multivariate Cox regression analysis: 
- Cox proportional hazard model accounting for: 
  - Bacterenia vs no-bacteremia (ref) within 1 yr following transplant
  - Ferritin levels: low (ref), moderate, high
  - Age, sex, transplant match
- Subgroup analyses: disease type

```{r beacteremia_pre_ferritin, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
# 0 is censored/survived, 1 is death
# exclude match_status and graft_manip
ferritin_bac <- coxph(Surv(bacteremia_time_1yr_censored/365.25, bacteremia) ~ ferritin_pre + age_at_ce + gender + transplant_type_combined, 
                data = pre_ferritin_bac)

# pre_ferritin_cox_1 %>% group_by(ferritin_pre, gender, transplant_type_combined, has_busulfan) %>%
#                       summarise(n_distinct(record_id))
# visualiza the impact on iron levels
ferritin_levels_df <- with(pre_ferritin_bac,
               data.frame(ferritin_pre = c("low", "moderate", "high"), 
                          age_at_ce = rep(mean(age_at_ce, na.rm = TRUE), 3),
                          gender = rep("Female", 3),
                          transplant_type_combined = rep("matched related",3),
                          has_busulfan = rep("No", 3)
                          )
               )

plot_title_text_size <- 20
axis_text_size <- 14

fit <- survfit(ferritin_bac, newdata = ferritin_levels_df)
ggsurvplot(fit, data = ferritin_levels_df, 
                  conf.int = FALSE, 
                  legend.labs = c("Low", "Moderate", "High"),
                  legend.title = "Pre-transplant ferritin levels", 
                  xlab = "Years", ylab = "Probability of bacteremia",
                  xlim = c(0, 1), ylim = c(0, .5), 
                  risk.table = TRUE,
                  fun = "event",
                  palette = c("red", "green", "blue"), # custom color palette
                  ggtheme = theme(
    plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size))) #, risk.table.title = "test")

```

```{r bacteremia_pre_ferritin_sum, echo= FALSE, message=FALSE}
# summary(res.cox)
cox_sum(ferritin_bac) %>% prettify_kable()
```

### subgroup analysis

```{r bacteremia_pre_ferritin_subgroup, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
# table1(~ ferritin_pre + age_at_ce + gender + transplant_type_combined + scd_type| bacteremia, 
#       data = pre_ferritin_bac %>% mutate(bacteremia = ifelse(is.na(bacteremia), "No bacteremia", "Bacteremia")) %>%
#               filter(scd_type  == "Diamond-Blackfan anemia"))

# 0 is censored/survived, 1 is death
# exclude match_status and graft_manip
ferritin_bac_aa <- coxph(Surv(bacteremia_time_1yr_censored/365.25, bacteremia) ~ ferritin_pre + age_at_ce + gender + transplant_type_combined, 
                data = pre_ferritin_bac %>% 
                      filter(scd_type == "aplastic anemia", transplant_type_combined != "autologous/genetherapy"))

ferritin_bac_scd <- coxph(Surv(bacteremia_time_1yr_censored/365.25, bacteremia) ~ ferritin_pre + age_at_ce + gender + transplant_type_combined, 
                data = pre_ferritin_bac %>% 
                      filter(scd_type == "sickle cell disease") %>%
                      filter(transplant_type_combined != "Allogeneic: Cord Blood", , ignore.case = TRUE))

ferritin_bac_btm <- coxph(Surv(bacteremia_time_1yr_censored/365.25, bacteremia) ~ ferritin_pre + age_at_ce + gender + transplant_type_combined, 
                data = pre_ferritin_bac %>% 
                      filter(scd_type == "beta thalassemia major"))

ferritin_bac_dba <- coxph(Surv(bacteremia_time_1yr_censored/365.25, bacteremia) ~ ferritin_pre + age_at_ce + gender + transplant_type_combined, 
                data = pre_ferritin_bac %>% 
                      filter(scd_type == "Diamond-Blackfan anemia", !(transplant_type_combined %in% c("autologous/genetherapy", "Unknown"))))

# pre_ferritin_cox_1 %>% group_by(ferritin_pre, gender, transplant_type_combined, has_busulfan) %>%
#                       summarise(n_distinct(record_id))
# visualiza the impact on iron levels
ferritin_pre_levels_df <- with(pre_ferritin_bac,
               data.frame(ferritin_pre = c("low", "medium", "high"), 
                          age_at_ce = rep(mean(age_at_ce, na.rm = TRUE), 3),
                          gender = rep("Female", 3),
                          transplant_type_combined = rep("matched related",3),
                          has_busulfan = rep("No", 3)
                          )
               )

fit_aa <- survfit(ferritin_bac_aa, newdata = ferritin_levels_df)
fit_scd <- survfit(ferritin_bac_scd, newdata = ferritin_levels_df)
fit_btm <- survfit(ferritin_bac_btm, newdata = ferritin_levels_df)
fit_dba <- survfit(ferritin_bac_dba, newdata = ferritin_levels_df)

p_aa <- ggsurvplot(fit_aa, data = ferritin_levels_df, 
                  conf.int = FALSE, 
                  # legend = "none",
                  title = "Aplastic anemia", 
                  legend.labs = c("Low", "medium", "High"),
                  legend.title = "Pre-transplant ferritin levels", 
                  xlab = "Years", ylab = "Probability of bacteremia",
                  xlim = c(0, 1), ylim = c(0, .5), 
                  risk.table = TRUE,
                  fun = "event",
                  palette = c("red", "green", "blue"), # custom color palette
                  ggtheme = theme(
    plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size))) #, risk.table.title = "test")

p_scd <- ggsurvplot(fit_scd, data = ferritin_levels_df, 
                  conf.int = FALSE, 
                  legend = "none",
                  title = "Sickle cell anemia", 
                  # legend.labs = c("Low", "High"),
                  # legend.title = "Pre-transplant LIC levels", 
                  xlab = "Years", ylab = "Probability of bacteremia",
                  xlim = c(0, 1), ylim = c(0, .5), 
                  risk.table = TRUE,
                  fun = "event",
                  palette = c("red", "green", "blue"), # custom color palette
                  ggtheme = theme(
    plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size))) #, risk.table.title = "test")

p_btm <- ggsurvplot(fit_btm, data = ferritin_levels_df, 
                  conf.int = FALSE, 
                  legend = "none",
                  title = "beta thalassemia major", 
                  # legend.labs = c("Low", "High"),
                  # legend.title = "Pre-transplant LIC levels", 
                  xlab = "Years", ylab = "Probability of bacteremia",
                  xlim = c(0, 1), ylim = c(0, .5), 
                  risk.table = TRUE,
                  fun = "event",
                  palette = c("red", "green", "blue"), # custom color palette
                  ggtheme = theme(
    plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size))) #, risk.table.title = "test")

p_dba <- ggsurvplot(fit_dba, data = ferritin_levels_df, 
                  conf.int = FALSE, 
                  title = "Diamond-Blackfan anemia", 
                  legend = "none", 
                  xlab = "Years", ylab = "Probability of bacteremia",
                  xlim = c(0, 1), ylim = c(0, .5), 
                  risk.table = TRUE,
                  fun = "event",
                  palette = c("red", "green", "blue"), # custom color palette
                  ggtheme = theme(
    plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.position = "none")) #, risk.table.title = "test")
  
grid.arrange(p_aa$plot, p_scd$plot, p_btm$plot, p_dba$plot, ncol = 2, nrow = 2)
```

#### aplastic anemia
##### KM curves
```{r bacteremia_pre_ferritin_subgroup_aa_km}
km_LIC_bac <- survfit(Surv(bacteremia_time_1yr_censored/365.25, bacteremia) ~ ferritin_pre,   
                    data = pre_ferritin_bac %>% 
                      filter(scd_type == "aplastic anemia", transplant_type_combined != "autologous/genetherapy"))

ggsurvplot(km_LIC_bac,
          pval = TRUE, conf.int = FALSE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          xlab = "Time (Years)",   # customize X axis label.
          ylab = "Overall Survival", 
          title = "Kaplan-Meier curves for bacteremia probability using pre-transplant ferritin levels --- aplastic anemia",
          legend.labs = c("Low", "Moderate", "High"),    # change legend labels.
          palette = c("red", "green", "blue"))
```

##### Multivariate Cox regression analysis: 
```{r bacteremia_pre_ferritin_subgroup_aa}
cox_sum(ferritin_bac_aa) %>% prettify_kable()
```

#### sickle cell disease
```{r bacteremia_pre_ferritin_subgroup_scd}
cox_sum(ferritin_bac_scd) %>% prettify_kable()
```

#### beta thalassemia major
```{r bacteremia_pre_ferritin_subgroup_btm}
cox_sum(ferritin_bac_btm) %>% prettify_kable()
```

#### Diamond-Blackfan anemia
```{r bacteremia_pre_ferritin_subgroup_dba}
cox_sum(ferritin_bac_dba) %>% prettify_kable()
```


## Bacteremia stratified by <span style="color:red">pre-transplant LIC levels</span>
### KM curves:
```{r bacteremia_pre_LIC_km}
pre_LIC_bac <- pre_LIC_cox %>% 
                          #filter(!(transplant_type_combined %in% c("autologous/genetherapy"))) %>%
                          # mutate(transplant_type_combined = relevel(transplant_type_combined, ref = "matched related"))
                          mutate(bacteremia = ifelse(is.na(bacteremia), FALSE, bacteremia))

table1(~ LIC_level_1_binary + age_at_ce + gender + transplant_type_combined| bacteremia, 
      data = pre_LIC_bac %>% mutate(bacteremia = ifelse(!bacteremia, "No bacteremia", "Bacteremia")),
      caption = "Bacteremia 5-yr post transplant for LIC cut-off = 8 mg Fe/g dry wt-- entire cohort")

table1(~ LIC_level_1_binary | bacteremia + scd_type,
      data = pre_LIC_bac %>% mutate(bacteremia = ifelse(!bacteremia, "No bacteremia", "Bacteremia")),
      overall = FALSE,
      caption = "Bacteremia 5-yr post transplant for LIC cut-off = 8 mg Fe/g dry wt")

table1(~ LIC_level_1_binary | bacteremia + scd_type, 
      data = pre_LIC_bac %>% 
            mutate(bacteremia = ifelse(!bacteremia, "No bacteremia", "Bacteremia")) %>%
            mutate(LIC_level_1_binary = if_else(LIC <5, "low", "high")),
      overall = FALSE,
      caption = "Bacteremia 5-yr post transplant for LIC cut-off = 5 mg Fe/g dry wt")

km_LIC_bac <- survfit(Surv(bacteremia_time_1yr_censored/365.25, bacteremia) ~ LIC_level_1_binary,   
                    data = pre_LIC_bac)

ggsurvplot(km_LIC_bac,
          pval = TRUE, conf.int = FALSE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          xlab = "Time (Years)",   # customize X axis label.
          ylab = "Overall Survival", 
          title = "Kaplan-Meier curves for bacteremia probability using pre-transplant LIC levels",
          legend.labs = c("Low", "High"),    # change legend labels.
          palette = c("red", "blue"))
```

### Multivariate Cox regression analysis: 
- Cox proportional hazard model accounting for: 
  - Bacteremia vs no-bacteremia (ref) within 1 yr following transplant
  - LIC levels: high vs. low (ref), using the most recent LIC values with binary cut-off of 8 
  - Age, sex, transplant match
- Subgroup analyses: disease type

```{r bacteremia_pre_LIC, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
# pre_ferritin_cox %>% group_by(survival_5yr, transplant_type_combined) %>% summarise(n = n_distinct(record_id)) %>% view()


# 0 is censored/survived, 1 is death
# exclude match_status and graft_manip
LIC_bac <- coxph(Surv(bacteremia_time_1yr_censored/365.25, bacteremia) ~ LIC_level_1_binary+ age_at_ce + gender + transplant_type_combined, 
                data = pre_LIC_bac %>% mutate(no_bacteremia = ifelse(bacteremia, FALSE, TRUE)))

# pre_ferritin_cox_1 %>% group_by(ferritin_pre, gender, transplant_type_combined, has_busulfan) %>%
#                       summarise(n_distinct(record_id))
# visualiza the impact on iron levels
LIC_levels_df <- with(pre_LIC_bac,
               data.frame(LIC_level_1_binary = c("low", "high"), 
                          age_at_ce = rep(mean(age_at_ce, na.rm = TRUE), 2),
                          gender = rep("Female", 2),
                          transplant_type_combined = rep("matched related",2),
                          has_busulfan = rep("No", 2)
                          )
               )

fit <- survfit(LIC_bac, newdata = LIC_levels_df)
ggsurvplot(fit, data = LIC_levels_df, 
                  conf.int = FALSE, 
                  legend.labs = c("Low", "High"),
                  legend.title = "Pre-transplant LIC levels", 
                  xlab = "Years", ylab = "Probability of bacteremia",
                  xlim = c(0, 1), ylim = c(0, .5), 
                  risk.table = TRUE,
                  fun = "event",
                  palette = c("red", "blue"), # custom color palette
                  ggtheme = theme(
    plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size))) #, risk.table.title = "test")
  
```

```{r bacteremia_pre_LIC_sum, echo= FALSE, message=FALSE}
# summary(res.cox)
cox_sum(LIC_bac) %>% prettify_kable()
```

### subgroup analysis

```{r bacteremia_pre_LIC_subgroup, echo= FALSE, message=FALSE, fig.width = 9, fig.height = 6}
# table1(~ LIC_level_1_binary + age_at_ce + gender + transplant_type_combined + scd_type| bacteremia, 
#       data = pre_LIC_bac %>% mutate(bacteremia = ifelse(!bacteremia, "No bacteremia", "Bacteremia")) %>%
#               filter(scd_type  == "Diamond-Blackfan anemia"))

# 0 is censored/survived, 1 is death
# exclude match_status and graft_manip
LIC_bac_aa <- coxph(Surv(bacteremia_time_1yr_censored/365.25, bacteremia) ~ LIC_level_1_binary+ age_at_ce + gender + transplant_type_combined, 
                data = pre_LIC_bac %>% 
                      filter(scd_type == "aplastic anemia", transplant_type_combined != "autologous/genetherapy"))

LIC_bac_scd <- coxph(Surv(bacteremia_time_1yr_censored/365.25, bacteremia) ~ LIC_level_1_binary+ age_at_ce + gender + transplant_type_combined, 
                data = pre_LIC_bac %>% 
                      filter(scd_type == "sickle cell disease") %>%
                      filter(transplant_type_combined != "Allogeneic: Cord Blood", , ignore.case = TRUE))

LIC_bac_btm <- coxph(Surv(bacteremia_time_1yr_censored/365.25, bacteremia) ~ LIC_level_1_binary+ age_at_ce + gender + transplant_type_combined, 
                data = pre_LIC_bac %>% 
                      filter(scd_type == "beta thalassemia major"))

LIC_bac_dba <- coxph(Surv(bacteremia_time_1yr_censored/365.25, bacteremia) ~ LIC_level_1_binary+ age_at_ce + gender + transplant_type_combined, 
                data = pre_LIC_bac %>% 
                      filter(scd_type == "Diamond-Blackfan anemia", !(transplant_type_combined %in% c("autologous/genetherapy", "Unknown"))))

# pre_ferritin_cox_1 %>% group_by(ferritin_pre, gender, transplant_type_combined, has_busulfan) %>%
#                       summarise(n_distinct(record_id))
# visualiza the impact on iron levels
LIC_levels_df <- with(pre_LIC_bac,
               data.frame(LIC_level_1_binary = c("low", "high"), 
                          age_at_ce = rep(mean(age_at_ce, na.rm = TRUE), 2),
                          gender = rep("Female", 2),
                          transplant_type_combined = rep("matched related",2),
                          has_busulfan = rep("No", 2)
                          )
               )

fit_aa <- survfit(LIC_bac_aa, newdata = LIC_levels_df)
fit_scd <- survfit(LIC_bac_scd, newdata = LIC_levels_df)
fit_btm <- survfit(LIC_bac_btm, newdata = LIC_levels_df)
fit_dba <- survfit(LIC_bac_dba, newdata = LIC_levels_df)

p_aa <- ggsurvplot(fit_aa, data = LIC_levels_df, 
                  conf.int = FALSE, 
                  # legend = "none",
                  title = "Aplastic anemia", 
                  legend.labs = c("Low", "High"),
                  legend.title = "Pre-transplant LIC levels", 
                  xlab = "Years", ylab = "Probability of bacteremia",
                  xlim = c(0, 1), ylim = c(0, .5), 
                  risk.table = TRUE,
                  fun = "event",
                  palette = c("red", "blue"), # custom color palette
                  ggtheme = theme(
    plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size))) #, risk.table.title = "test")

p_scd <- ggsurvplot(fit_scd, data = LIC_levels_df, 
                  conf.int = FALSE, 
                  legend = "none",
                  title = "sickle cell anemia", 
                  # legend.labs = c("Low", "High"),
                  # legend.title = "Pre-transplant LIC levels", 
                  xlab = "Years", ylab = "Probability of bacteremia",
                  xlim = c(0, 1), ylim = c(0, .5), 
                  risk.table = TRUE,
                  fun = "event",
                  palette = c("red", "blue"), # custom color palette
                  ggtheme = theme(
    plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size))) #, risk.table.title = "test")

p_btm <- ggsurvplot(fit_btm, data = LIC_levels_df, 
                  conf.int = FALSE, 
                  legend = "none",
                  title = "beta thalassemia major", 
                  # legend.labs = c("Low", "High"),
                  # legend.title = "Pre-transplant LIC levels", 
                  xlab = "Years", ylab = "Probability of bacteremia",
                  xlim = c(0, 1), ylim = c(0, .5), 
                  risk.table = TRUE,
                  fun = "event",
                  palette = c("red", "blue"), # custom color palette
                  ggtheme = theme(
    plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.title = element_text(size = axis_text_size),    # Legend title size
    legend.text = element_text(size = axis_text_size))) #, risk.table.title = "test")

p_dba <- ggsurvplot(fit_dba, data = LIC_levels_df, 
                  conf.int = FALSE, 
                  title = "Diamond-Blackfan anemia", 
                  legend = "none", 
                  xlab = "Years", ylab = "Probability of bacteremia",
                  xlim = c(0, 1), ylim = c(0, .5), 
                  risk.table = TRUE,
                  fun = "event",
                  palette = c("red", "blue"), # custom color palette
                  ggtheme = theme(
    plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size),   # Y-axis labels size
    legend.position = "none")) #, risk.table.title = "test")
  
grid.arrange(p_aa$plot, p_scd$plot, p_btm$plot, p_dba$plot, ncol = 2, nrow = 2)
```

#### aplastic anemia
##### KM curves
```{r bacteremia_pre_LIC_subgroup_aa_km}
km_LIC_bac <- survfit(Surv(bacteremia_time_1yr_censored/365.25, bacteremia) ~ LIC_level_1_binary,   
                    data = pre_LIC_bac %>% 
                    filter(scd_type == "aplastic anemia", transplant_type_combined != "autologous/genetherapy"))

ggsurvplot(km_LIC_bac,
          pval = TRUE, conf.int = FALSE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          xlab = "Time (Years)",   # customize X axis label.
          ylab = "Overall Survival", 
          title = "Kaplan-Meier curves for bacteremia probability using pre-transplant LIC levels --- aplastic anemia",
          legend.labs = c("Low", "High"),    # change legend labels.
          palette = c("red", "blue"))
```

##### Multivariate Cox regression analysis: 
```{r bacteremia_pre_LIC_subgroup_aa}
cox_sum(ferritin_bac_aa) %>% prettify_kable()
```

#### sickle cell disease
```{r bacteremia_pre_LIC_subgroup_scd}
cox_sum(ferritin_bac_scd) %>% prettify_kable()
```

#### beta thalassemia major
```{r bacteremia_pre_LIC_subgroup_btm}
cox_sum(ferritin_bac_btm) %>% prettify_kable()
```

#### Diamond-Blackfan anemia
```{r bacteremia_pre_LIC_subgroup_dba}
cox_sum(ferritin_bac_dba) %>% prettify_kable()
```

# VOD
## Odds of having VOD using <span style="color:red">pre-transplant ferritin levels</span>
- Multivariate logistic regression ONLY patients who received busulfan accounting for: 
  - VOD vs no-VOD (ref) within 90 days following transplant
  - Ferritin levels: low (ref), moderate, high (Can also consider < 3000/>3000)
  - Age, sex
- Subgroup analyses: disease type

```{r vod_preferritin, fig.width = 6, fig.height = 4}
pre_ferritin_vod_lr <- pre_ferritin_cox %>% 
                          filter(has_busulfan == "Yes") %>%
                          mutate(vod = ifelse(vod, "Yes", "No")) %>%
                          mutate(vod= relevel(factor(vod), ref = "No")) %>%
                          mutate(transplant_type_combined = relevel(factor(transplant_type_combined), ref = "matched related"))

vod_fit <- glm_weightit(formula = vod ~ gender + age_at_ce +  ferritin_pre,
            data = pre_ferritin_vod_lr, 
            family = "binomial", link = "logit", estimand = "ATT")

sjPlot::tab_model(vod_fit, title = "Odds of getting VOD for patients treated with bulsufan")

# sjPlot::tab_model(unweighted, title = "Odds of having masked hypertension for all age groups UNWEIGHTED")
p1 <- plot_summs(vod_fit, exp = TRUE)
require(sjPlot)
require(jtools)
require(broom.mixed)
p1 <- p1 + scale_y_discrete(labels = c("High ferritin", "Moderate ferritin", "Age at transplant", "Male")) + 
      labs(x = "Odds ratio", y = "")
print(p1)

table1(~scd_type|vod, data = pre_ferritin_vod_lr)
```

### subgroup analysis
```{r vod_preferritin_subgroups, fig.width = 6, fig.height = 4}

vod_fit_scd <- glm_weightit(formula = vod ~ gender + age_at_ce +  ferritin_pre,
            data = pre_ferritin_vod_lr %>% filter(scd_type == "sickle cell disease"), 
            family = "binomial", link = "logit", estimand = "ATT")
vod_fit_btm <- glm_weightit(formula = vod ~ gender + age_at_ce +  ferritin_pre,
            data = pre_ferritin_vod_lr %>% filter(scd_type == "beta thalassemia major"), 
            family = "binomial", link = "logit", estimand = "ATT")
vod_fit_dba <- glm_weightit(formula = vod ~ gender + age_at_ce +  ferritin_pre,
            data = pre_ferritin_vod_lr %>% filter(scd_type == "Diamond-Blackfan anemia"), 
            family = "binomial", link = "logit", estimand = "ATT")
vod_fit_aa <- glm_weightit(formula = vod ~ gender + age_at_ce +  ferritin_pre,
            data = pre_ferritin_vod_lr %>% filter(scd_type == "aplastic anemia"), 
            family = "binomial", link = "logit", estimand = "ATT")

export_summs(vod_fit_aa, vod_fit_scd, vod_fit_btm, vod_fit_dba, scale = TRUE,
             error_format = "[{conf.low}, {conf.high}]", #p = {p.value}, 
             exp = TRUE,
             model.names = c("aplastic anemia", "sickle cell disease", "beta thalassemia major", "Diamond-Blackfan anemia"))

# sjPlot::tab_model(unweighted, title = "Odds of having masked hypertension for all age groups UNWEIGHTED")
p2 <- plot_summs(vod_fit_aa, vod_fit_scd, vod_fit_btm, vod_fit_dba, 
        exp = TRUE,
        model.names = c("aplastic anemia", "sickle cell disease", "beta thalassemia major", "Diamond-Blackfan anemia"))
p2 <- p2 + scale_y_discrete(labels = c("High ferritin", "Moderate ferritin", "Age at transplant", "Male")) + 
      labs(x = "Odds ratio", y = "")
print(p2)

```

## Odds of having VOD using <span style="color:red">pre-transplant LIC levels</span>
- Multivariate logistic regression ONLY patients who received busulfan accounting for: 
  - VOD vs no-VOD (ref) within 90 days following transplant
  - LIC levels: low (ref), high
  - Groups: Low, moderate, high, if possible. 
  - Age, sex
- Subgroup analyses: disease type

```{r vod_preLIC, fig.width = 6, fig.height = 4}
pre_LIC_vod_lr <- pre_LIC_cox %>% 
                          filter(has_busulfan == "Yes") %>%
                          mutate(vod = ifelse(vod, "Yes", "No")) %>%
                          mutate(vod= relevel(factor(vod), ref = "No")) %>%
                          mutate(transplant_type_combined = relevel(factor(transplant_type_combined), ref = "matched related"))

vod_fit <- glm_weightit(formula = vod ~ gender + age_at_ce +  LIC_level_1_binary,
            data = pre_LIC_vod_lr, 
            family = "binomial", link = "logit", estimand = "ATT")

sjPlot::tab_model(vod_fit, title = "Odds of getting VOD for patients treated with bulsufan")

# sjPlot::tab_model(unweighted, title = "Odds of having masked hypertension for all age groups UNWEIGHTED")
p3 <- plot_summs(vod_fit, exp = TRUE)
require(sjPlot)
require(jtools)
require(broom.mixed)
p3 <- p3 + scale_y_discrete(labels = c("High LIC", "Age at transplant", "Male")) + 
      labs(x = "Odds ratio", y = "")
print(p3)

table1(~scd_type|vod, data = pre_LIC_vod_lr)
```

### subgroup analysis
- Omit aplastic anemia due to small sample size
```{r vod_preLIC_subgroups, fig.width = 6, fig.height = 4}

vod_fit_scd <- glm_weightit(formula = vod ~ gender + age_at_ce +  LIC_level_1_binary,
            data = pre_LIC_vod_lr %>% filter(scd_type == "sickle cell disease"), 
            family = "binomial", link = "logit", estimand = "ATT")
vod_fit_btm <- glm_weightit(formula = vod ~ gender + age_at_ce +  LIC_level_1_binary,
            data = pre_LIC_vod_lr %>% filter(scd_type == "beta thalassemia major"), 
            family = "binomial", link = "logit", estimand = "ATT")
vod_fit_dba <- glm_weightit(formula = vod ~ gender + age_at_ce +  LIC_level_1_binary,
            data = pre_LIC_vod_lr %>% filter(scd_type == "Diamond-Blackfan anemia"), 
            family = "binomial", link = "logit", estimand = "ATT")
# vod_fit_aa <- glm_weightit(formula = vod ~ gender + age_at_ce +  ferritin_pre,
#             data = pre_ferritin_vod_lr %>% filter(scd_type == "aplastic anemia"), 
#             family = "binomial", link = "logit", estimand = "ATT")

export_summs(vod_fit_scd, vod_fit_btm, vod_fit_dba, scale = TRUE,
             error_format = "[{conf.low}, {conf.high}]", #p = {p.value}, 
             exp = TRUE, 
             model.names = c("sickle cell disease", "beta thalassemia major", "Diamond-Blackfan anemia"))

# sjPlot::tab_model(unweighted, title = "Odds of having masked hypertension for all age groups UNWEIGHTED")
p4 <- plot_summs(vod_fit_scd, vod_fit_btm, vod_fit_dba, 
        exp = TRUE,
        model.names = c("sickle cell disease", "beta thalassemia major", "Diamond-Blackfan anemia"))
p4 <- p4 + scale_y_discrete(labels = c("High LIC", "Age at transplant", "Male")) + 
      labs(x = "Odds ratio", y = "")
print(p4)

```


# Ferritin vs. LIC correlation
- The ferritin value used should be the closest one to the LIC, within 6 months. If there is no ferritin within 6 months of the LIC, omit the values. 
- When there are more than 1 ferritin values within 6 months, take the closet one. 

```{r ferritin_LIC_corr}
LIC <- results_tbl("cr_data") %>% select(record_id, site, eligibility, LIC_date, LIC) %>%
            left_join(results_tbl("study_cohorts") %>% select(person_id, record_id), by = "record_id") %>%
            filter(!is.na(LIC))
ferritin <- results_tbl("covar_ferritin_mx") %>% select(person_id, ferritin, ferritin_date)

# get the closet measurements within 14 days
LIC_ferritin <- LIC %>% inner_join(ferritin, by = "person_id") %>%
    collect() %>%
    mutate(LIC_ferritin_period = abs(as.numeric(difftime(LIC_date, ferritin_date, units = "days")))) %>%
    filter(LIC_ferritin_period <= 180) %>%
    group_by(record_id, LIC_date) %>% #if a single value matches to multiple values, take the closet one
    slice_min(LIC_ferritin_period, with_ties = FALSE, na_rm = TRUE)

LIC_ferritin %>% ggplot(aes(x = ferritin, y = LIC)) +
  geom_point(color = "black", size = 3.5, alpha = 0.6) + # Scatter points
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") + # Regression line
  stat_regline_equation(label.x = 7500, label.y = 50, aes(label = ..eq.label..), 
                        size = 8) + # Equation
  stat_regline_equation(label.x = 7500, # R-squared
                        label.y = 40, 
                        aes(label = paste(..rr.label..)), 
                        parse = TRUE, # Enables proper superscript
                        size = 8) + # Increase text size
  labs(title = "LIC vs. Ferritin Correlation",
       x = "Ferritin (ng/mL)",
       y = "LIC (mg Fe/g dry wt)") +
  theme(plot.title = element_text(size = plot_title_text_size),    # Title text size
    axis.title.x = element_text(size = axis_text_size), # X-axis title size
    axis.title.y = element_text(size = axis_text_size), # Y-axis title size
    axis.text.x = element_text(size = axis_text_size),  # X-axis labels size
    axis.text.y = element_text(size = axis_text_size)   # Y-axis labels size
  )  
  # theme_minimal() 
```


# Sensitivity analysis
## Overall survival: Subgroup analysis for pre-transplant LIC binary classification
### sickle cell disease --- Overall survival: Subgroup analysis for pre-transplant LIC binary levels

- In this subgroup analysis: I have to remove the transplant type from the list of covariates. 
- The LIC cut-off for high vs. low is 8

```{r pre_LIC_binary_OS_disease_subgroup_scd}
table1(~age_at_ce + gender + transplant_type_combined + LIC_level_1_binary + has_busulfan| survival_5yr, 
        data = pre_LIC_cox %>% mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")) %>%
                filter(scd_type == "sickle cell disease"), 
        overall = FALSE, 
        caption = "Overall survival at 5-yr following transplant --- sickle cell disease")     

pre_LIC_os_scd <- pre_LIC_cox %>%
    # filter(!(transplant_type_combined %in% c("mismatched related", "mismatched unrelated", "autologous/genetherapy"))) %>%
    cox_fit_plot(time_var = "survival_time_5yr_censored", 
                status_var = "survival_5yr", 
                covariates = c("LIC_level_1_binary", "age_at_ce", "gender", "has_busulfan"),
                iron_level_var = "LIC_level_1_binary", 
                plot_labels = c("Pre-transplant LIC binary levels", "Overall survival probability"),
                sub_group = "sickle cell disease")

print(pre_LIC_os_scd$p)
cox_sum(pre_LIC_os_scd$fit) %>% prettify_kable()
```

### aplastic anemia --- Overall survival: Subgroup analysis for pre-transplant LIC binary levels

- In this subgroup analysis: I have to remove the transplant type from the list of covariates. 

```{r pre_LIC_binary_OS_disease_subgroup_aa}
table1(~age_at_ce + gender + transplant_type_combined + LIC_level_1_binary + has_busulfan| survival_5yr, 
        data = pre_LIC_cox %>% mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")) %>%
                filter(scd_type == "aplastic anemia"), 
        overall = FALSE, 
        caption = "Overall survival at 5-yr following transplant --- aplastic anemia")     

pre_LIC_os_aa <- pre_LIC_cox %>%
    # filter((transplant_type_combined == "matched unrelated/mismatched unrelated/mismatched related")) %>% distinct(transplant_type_combined)
    cox_fit_plot(time_var = "survival_time_5yr_censored", 
                status_var = "survival_5yr", 
                covariates = c("LIC_level_1_binary", "age_at_ce", "gender", "has_busulfan"),
                iron_level_var = "LIC_level_1_binary", 
                plot_labels = c("Pre-transplant LIC binary levels", "Overall survival probability"),
                sub_group = "aplastic anemia")

print(pre_LIC_os_aa$p)
cox_sum(pre_LIC_os_aa$fit) %>% prettify_kable()
```

### beta thalassemia major  --- Overall survival: Subgroup analysis for pre-transplant LIC binary levels

```{r pre_LIC_binary_OS_disease_subgroup_btm}
table1(~age_at_ce + gender + transplant_type_combined + LIC_level_1_binary + has_busulfan| survival_5yr, 
        data = pre_LIC_cox %>% mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")) %>%
                filter(scd_type == "beta thalassemia major"), 
        overall = FALSE, 
        caption = "Overall survival at 5-yr following transplant --- beta thalassemia major")     

pre_LIC_os_btm <- pre_LIC_cox %>%
    filter(!(transplant_type_combined %in% c("autologous/genetherapy"))) %>%
    cox_fit_plot(time_var = "survival_time_5yr_censored", 
                status_var = "survival_5yr", 
                covariates = c("LIC_level_1_binary", "age_at_ce", "gender", "has_busulfan", "transplant_type_combined"),
                iron_level_var = "LIC_level_1_binary", 
                plot_labels = c("Pre-transplant LIC binary levels", "Overall survival probability"),
                sub_group = "beta thalassemia major")

print(pre_LIC_os_btm$p)
cox_sum(pre_LIC_os_btm$fit) %>% prettify_kable()
```

## Overall survival: Subgroup analysis for pre-transplant ferritin levels
### aplastic anemia  --- Overall survival: Subgroup analysis for pre-transplant ferritin levels
```{r pre_ferritin_OS_disease_subgroup_aa}
table1(~age_at_ce + gender + transplant_type_combined + ferritin_pre + has_busulfan| survival_5yr, 
        data = pre_ferritin_cox %>% mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")) %>%
                filter(scd_type == "aplastic anemia"), 
        overall = FALSE, 
        caption = "Overall survival at 5-yr following transplant --- aplastic anemia")                          

pre_ferritin_os_aa <- pre_ferritin_cox %>%
    filter(!(transplant_type_combined %in% c("Unknown", "autologous/genetherapy"))) %>%
    cox_fit_plot(time_var = "survival_time_5yr_censored", 
                status_var = "survival_5yr", 
                covariates = c("ferritin_pre", "age_at_ce", "gender", "transplant_type_combined", "has_busulfan"),
                iron_level_var = "ferritin_pre", 
                plot_labels = c("Pre-transplant ferritin levels", "Overall survival probability"),
                sub_group = "aplastic anemia")
print(pre_ferritin_os_aa$p)
cox_sum(pre_ferritin_os_aa$fit) %>% prettify_kable()
```

### sickle cell disease  --- Overall survival: Subgroup analysis for pre-transplant ferritin levels
```{r pre_ferritin_OS_disease_subgroup_scd}
table1(~age_at_ce + gender + transplant_type_combined + ferritin_pre + has_busulfan| survival_5yr, 
        data = pre_ferritin_cox %>% mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")) %>%
                filter(scd_type == "sickle cell disease"), 
        overall = FALSE, 
        caption = "Overall survival at 5-yr following transplant --- sickle cell disease")                          

pre_ferritin_os_scd <- pre_ferritin_cox %>%
    filter((transplant_type_combined %in% c("matched related", "matched unrelated/mismatched unrelated/mismatched related"))) %>%
    cox_fit_plot(time_var = "survival_time_5yr_censored", 
                status_var = "survival_5yr", 
                covariates = c("ferritin_pre", "age_at_ce", "gender", "has_busulfan","transplant_type_combined"),
                iron_level_var = "ferritin_pre", 
                plot_labels = c("Pre-transplant ferritin levels", "Overall survival probability"),
                sub_group = "sickle cell disease")

print(pre_ferritin_os_scd$p)
cox_sum(pre_ferritin_os_scd$fit) %>% prettify_kable()
```

### beta thalassemia major  --- Overall survival: Subgroup analysis for pre-transplant ferritin levels
```{r pre_ferritin_OS_disease_subgroup_btm}
table1(~age_at_ce + gender + transplant_type_combined + ferritin_pre + has_busulfan| survival_5yr, 
        data = pre_ferritin_cox %>% mutate(survival_5yr = ifelse(survival_5yr ==1, "Dead", "Survived")) %>%
                filter(scd_type == "beta thalassemia major"), 
        overall = FALSE, 
        caption = "Overall survival at 5-yr following transplant --- beta thalassemia major")     

pre_ferritin_os_btm <- pre_ferritin_cox %>%
    filter(!(transplant_type_combined %in% c("autologous/genetherapy"))) %>%
    cox_fit_plot(time_var = "survival_time_5yr_censored", 
                status_var = "survival_5yr", 
                covariates = c("ferritin_pre", "age_at_ce", "gender", "has_busulfan", "transplant_type_combined"),
                iron_level_var = "ferritin_pre", 
                plot_labels = c("Pre-transplant ferritin levels", "Overall survival probability"),
                sub_group = "beta thalassemia major")

print(pre_ferritin_os_btm$p)
cox_sum(pre_ferritin_os_btm$fit) %>% prettify_kable()

# pre_ferritin_cox_1 %>% filter(scd_type == "sickle cell disease") %>%
#       group_by(transplant_type_combined, survival_5yr) %>% summarise(n = n_distinct(record_id))
```
