---
title: "The Effects of Iron Overload on Hemptotic Stemcell Transplant (HSCT) Outcomes in Sickle Cell Patients"
author: "Nhat Nguyen \n PEDSnet Coordinating Center"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    page-layout: full
    embed-resources: true
    smooth-scroll: true
    fig-cap: TRUE
    df-print: paged
    toc: TRUE
    toc-depth: 5
    toc-collapsed: TRUE
    toc-float: TRUE
    toc-location: left
    number-sections: TRUE
    number-depth: 3
    output-file: analysis
    output-ext: "html"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 6, fig.align = "center")
# Default taken from R notebook behavior: when knitting, wd will always be location of notebook
base_dir <- '..'
Sys.setenv(PEDSNET_SKIP_REQUEST_EXECUTION=1)
tryCatch(source('../site/run.R'),
         error = function (e) message(e)) # May not be able to make db connection

# Set to "local" to read data from ../results, or another value to read from db
data_source <- "not_local" #if_else(config('execution_mode') == 'distribution', 'local', 'not_local')

```
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
tribble( ~Item, ~Value,
         'Title', config('title'),
         'Requester', config('requester_name'),
         'CDM Version', as.character(config('cdm_version')),
         'Request Version', as.character(config('req_version'))) %>%
  prettify_kable()

```
```{r data, include= FALSE}
cohort <- results_tbl("study_cohorts") %>%
                    pivot_longer(cols = aim_2b_2:aim_3_1, names_to = "aim", values_to = "which_aim") %>% 
                    select(person_id, aim, transplant_site= site, age_at_ce , gender, scd_concept_id, scd_concept_name, scd_type, transplant_concept_id, transplant_concept_name, transplant_type)
```

## Confidentiality

Please treat the contents of this report as confidential, and do not redistrubute beyond PEDSnet staff and the requester.

# Cohort Definition

-   **Time Anchors:** The specific details of cohort formation are described below. Note that for dates, we use notation to indicate time, where 0 refers to the day that a patient enters the cohort, a positive integer refers to the number of days after the cohort entrance that the observation period lasts, a negative interger refers to the number of days before the cohort entrance that the observation period lasts, oo refers to the end of the study period, and -oo refers to the beginning of the study period.
-   **Study period:** 1/1/2010 – 6/30/2023
-   **Cohort entrance date:** Date of first transplant
-   **Follow-up window:** 
    - [-365, 00]: interval during which outcome(s) of interest may be included in the analysis.
    - Exposures of interest (LIC, ferritin) for Aims 2a can be observed from -365 days to 0.
    - Exposures of interest (LIC, ferritin) for Aims 2b and 3 can be observed from day +90 to +365.
    - Exposures of interest (deferasirox or deferoxamine prescription, phlebotomy procedure) for Aim 3 can be observed from day +90 to 00. 
    - Outcomes of interest for Aim 2a can be observed from day 0 to 00 
    - Outcomes of interest for Aim 2b can be observed from date of post-transplant LIC or ferritin to 00. 
    - Outcomes of interest for Aim 3 can be observed from date of prescription of deferasirox or deferoxamine to 00, or date of first phlebotomy to 00
-   **Inclusion/exclusion criteria:** 
    - Aim 1: Pre-HSCT iron overload: Exclude patients who do not have either a pre-HSCT ferritin or LIC. Post-HSCT iron overload: exclude patients who do not have an LIC or ferritin 6-12 months after HSCT. 
    - Aim 2a: Include patients whose first transplant between 1/1/2010 – 6/30/2023. Exclude patients who do not have either a pre-HSCT ferritin or LIC. 
    - Aim 2b and 3: Include patients whose first transplant between 1/1/2010 – 12/1/2022. Exclude patients who do not have an LIC or ferritin 6-12 months after HSCT. 
    - Aim 3: Exclude patients who did not require post-transplant chelation or phlebotomy.

# Conditions and Procedure Codesets:

-   A compehensive code search was performed for all the conditions, procedures, and measurements specified in the DAP. The list of codesets could be found [here](https://chop365.sharepoint.com/teams/RSCH-ACRC/Shared%20Documents/Forms/AllItems.aspx?id=%2Fteams%2FRSCH%2DACRC%2FShared%20Documents%2FPEDSnet%2FPEDSnet%20Studies%2FActive%20Studies%2FGibson%5FIron%20Overload%2FQueries%2FCodesets%5F2024Feb27&viewid=fbc819dd%2Dfbe7%2D4805%2Da6fc%2D03eda92856f5).


# Result Summary

The cohort comprises `r summarise(cohort_all, ct = n_distinct(person_id)) %>% pull(ct)` patients from `r summarise(cohort_all, ct = n_distinct(transplant_site)) %>% pull(ct)` PEDSnet sites.

## Site stratification
```{r site_strat}
  ct <- cohort_all %>% ungroup() %>% distinct(person_id) %>% count() %>% pull(n)
  cohort_all %>% group_by(transplant_site) %>%
  mutate(persons = n_distinct(person_id)) %>% group_by(transplant_site, cohort) %>% 
    summarize(cohort_size = n_distinct(person_id), total = unique(persons)) %>% 
    mutate(cohort_size = if_else(cohort_size < 11L, "<11", as.character(cohort_size))) %>%
  pivot_wider(names_from = cohort, values_from = cohort_size) %>% 
  select(1, 3:8, 2) %>% mutate(perc = round(total/ct*100, 2)) %>% 
  rename(site = transplant_site) %>%
  prettify_kable()
```

## Attrition

```{r attrition, echo=FALSE}
  cohort_2a_1 %>% mutate(cohort = "Aim 2a.1") %>% select(person_id, cohort) %>%
                bind_rows(cohort_2a_2 %>% mutate(cohort = "Aim 2a.2") %>% select(person_id, cohort)) %>% 
                bind_rows(cohort_2b_1 %>% mutate(cohort = "Aim 2b.1") %>% select(person_id, cohort)) %>% 
                bind_rows(cohort_2b_2 %>% mutate(cohort = "Aim 2b.2") %>% select(person_id, cohort)) %>% 
                bind_rows(cohort_3_1 %>% mutate(cohort = "Aim 3.1") %>% select(person_id, cohort)) %>% 
                bind_rows(cohort_3_2 %>% mutate(cohort = "Aim 3.2") %>% select(person_id, cohort)) %>% 
                group_by(cohort) %>% 
                summarise(persons = n_distinct(person_id)) %>% 
                mutate(persons = if_else(persons<11, "<11", as.character(persons))) %>%
                prettify_kable()




# DBI doesn't have a good way to interrogate tables in a specified schema,
# so you have to list the step logs you want to summarise (e.g. DCC + remote
# sites).
# step_log <- list(config('qry_site')) %>%
#   map(.f = function(x) {
#     get_results(paste0(x, '_steps'))
#   }) %>%
#   reduce(.f = dplyr::union)
# 
# step_log %>%
#   arrange(stamp) %>%
#   select(cohort, persons) %>%
#   mutate(persons = if_else(persons > 0L & persons < 11L,
#                            as.integer(NA), as.integer(persons))) %>%
#   mutate(percent_prior = round(100 * persons /
#                                  coalesce(lag(persons), persons), 2)) %>%
#   prettify_kable()
```

## Demographics

```{r demog, echo=FALSE}

cohort_all %>% group_by(gender) %>% summarise(persons = n_distinct(person_id)) %>% 
            mutate(perc = paste0(round(persons/ct*100, 2), " %")) %>% prettify_kable()

cohort_all %>% mutate(age_group = case_when(age_at_transplant < 5 ~ "[0, 5)",
                                            age_at_transplant < 10 ~ "[5, 10)",
                                            age_at_transplant < 15 ~ "[10, 15)",
                                            age_at_transplant < 20 ~ "[15, 20)",
                                            age_at_transplant < 40 ~ "[20, 40)",
                                            TRUE ~ "40+")) %>% 
               group_by(age_group, gender) %>% 
               summarise(persons = n_distinct(person_id)) %>%
               pivot_wider(names_from = gender, values_from = persons) %>% 
               mutate(F_perc = paste0(round(F/ct*100, 2), "%"),
                      M_perc = paste0(round(M/ct*100, 2), "%"),
                      F = if_else(F<11, "<11", as.character(F)),
                      M = if_else(M<11, "<11", as.character(M))) %>%
               slice(1, 4, 5, 2, 3) %>% select(age_group, Female = F, F_perc, Male = M, M_perc) %>% 
               prettify_kable()

# c_names <- vocabulary_tbl('concept') %>%
#   filter(domain_id %in% c('Gender', 'Race', 'Ethnicity')) %>%
#   select(concept_id, concept_name) %>% collect()
# 
# demog <- cdm_tbl('person') %>%
#   semi_join(distinct(cohort, person_id),
#             by = 'person_id', copy = TRUE) %>% collect() %>%
#   inner_join(select(c_names, concept_id, sex = concept_name),
#              by = c('gender_concept_id' = 'concept_id')) %>%
#   inner_join(select(c_names, concept_id, race = concept_name),
#              by = c('race_concept_id' = 'concept_id')) %>%
#   inner_join(select(c_names, concept_id, hispanic = concept_name),
#              by = c('ethnicity_concept_id' = 'concept_id')) %>%
#   select(person_id, sex, race, hispanic)
# 
# h_cts <- demog %>% group_by(sex, hispanic) %>% summarise(persons = n()) %>%
#   mutate(persons = if_else(persons > 0L & persons < 11L,
#                            as.integer(NA), as.integer(persons))) %>%
#   pivot_wider(names_from = sex, values_from = persons) %>% rename(group = hispanic)
# r_cts <- demog %>% group_by(sex, race) %>% summarise(persons = n()) %>%
#   mutate(persons = if_else(persons > 0L & persons < 11L,
#                            as.integer(NA), as.integer(persons))) %>%
#   pivot_wider(names_from = sex, values_from = persons) %>% rename(group = race)
# 
# prettify_kable(r_cts)
# prettify_kable(h_cts)

```

## Sample Counts by Codesets:

The below tables only showed the top most frequent codesets. 

### Non-malignant sickle cell dieases
```{r codeset_ct_scd, echo = FALSE}
    cohort_all %>% group_by(scd_concept_id, scd_concept_name, scd_type) %>% summarise(persons = n_distinct(person_id)) %>% 
    ungroup() %>%
    arrange(desc(persons)) %>% 
    mutate(scd_concept_id = as.character(scd_concept_id), 
            scd_concept_name = as.character(scd_concept_name),
            persons = if_else(persons < 11, "<11", as.character(persons))) %>% 
    slice_head(n = 15) %>% prettify_kable()
```
### Steam cell transplants
```{r codeset_ct_transplant, echo = FALSE}
    cohort_all %>% group_by(transplant_concept_id, transplant_concept_name, transplant_type) %>% summarise(persons = n_distinct(person_id)) %>% 
    ungroup() %>%
    arrange(desc(persons)) %>% 
    mutate(transplant_concept_id = as.character(transplant_concept_id), 
            transplant_concept_name = as.character(transplant_concept_name),
            persons = if_else(persons < 11, "<11", as.character(persons))) %>% 
    slice_head(n = 15) %>% prettify_kable()
```
### MRI
```{r codeset_ct_mri, echo = FALSE}
    cohort_2a_1 %>% union(cohort_2b_1) %>%
      group_by(mri_concept_id, mri_concept_name) %>%
      summarise(persons = n_distinct(person_id)) %>%
      arrange(desc(persons)) %>%
      mutate(persons = if_else(persons < 11, "<11", as.character(persons)),
            mri_concept_id = as.character(mri_concept_id)) %>%
      prettify_kable()
```
### Therapeutic phlebotomy
```{r codeset_ct_phleb, echo = FALSE}
    cohort_3_1 %>% 
      group_by(phleb_concept_id, phleb_concept_name) %>%
      summarise(persons = n_distinct(person_id)) %>%
      arrange(desc(persons)) %>%
      mutate(persons = if_else(persons < 11, "<11", as.character(persons)),
            phleb_concept_id = as.character(phleb_concept_id)) %>%
      prettify_kable()
```
### Drugs
```{r codeset_ct_drugs, echo = FALSE}
    cohort_3_2 %>% 
      group_by(drug_type) %>%
      summarise(persons = n_distinct(person_id)) %>%
      arrange(desc(persons)) %>%
      mutate(persons = if_else(persons < 11, "<11", as.character(persons))) %>%
      prettify_kable()

    cohort_3_2 %>% 
      # filter( == "deferoxamine") %>% 
      group_by(drug_type, drug_concept_id, drug_concept_name) %>%
      summarise(persons = n_distinct(person_id)) %>%
      # arrange(desc(persons)) %>%
      slice_max(persons, n = 10, with_ties = TRUE) %>%
      mutate(persons = if_else(persons < 11, "<11", as.character(persons)),
            drug_concept_id = as.character(drug_concept_id)) %>%
      prettify_kable()
```

# Aim 2_1:
- This is a retrospective cohort study comparing outcomes between patients with pre-HSCT iron overload to patients without pre-HSCT iron overload. Analyze the impact of pre-HSCT elevated ferritin and LIC on short-term hematologic and transplant-related outcomes. These outcomes include overall survival, graft-versus-host disease (GvHD), graft failure, incidence of veno-occlusive disease (VOD), bacteremia, immune reconstitution, and time to neutrophil and platelet engraftment. 

## Descriptive analysis: 
    - Percentage of overall survival in iron-overloaded patients vs non-iron overloaded patients 
    - Percentage of patients who experienced disease relapse in iron-overloaded patients vs non-iron overloaded patients 
    - Percentage of patients with GVHD in iron-overloaded patients vs non-iron overloaded patients 
    - Percentage of patients with VOD in iron-overloaded patients vs non-iron overloaded patients 
    - Neutrophil engraftment times in iron-overloaded patients vs non-iron overloaded patients 
    - Platelet engraftment times in iron-overloaded patients vs non-iron overloaded patients 

```{r aim_2_1}
cohort %>% select(person_id, scd_concept_id, scd_concept_name, type = scd_type, site = transplant_site) %>%
    concept_counts(scd_concept_name, scd_concept_id) %>% 
    ungroup() %>% collect_new() %>%
    mask_table(col_groups = c("persons"), percentage = TRUE, overwrite_columns = FALSE)
```

## Statistical Analysis
- GRFS: Time-to-event analysis with Cox proportional hazards model, Kaplan-Meier curves with log-rank test 
- Overall survival: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test.  
- Disease relapse: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test.  
- GVHD: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test.  
- VOD: Multivariable logistic regression with stratification for busulfan conditioning, use interaction term for busulfan and iron score 
- Time to neutrophil and platelet engraftment: time-to-event analysis using a competing risks analysis. Competing risks include death or graft failure prior to engraftment. 
- Bacteremia within 1 year after HSCT: Time-to-event analysis using a competing risks analysis. Competing risk includes death. 
- Immune Reconstitution: Definition of immune reconstitution: CD3 count >500, CD4 count >200, CD8 count >200, IgM >25. 
- Statistical analysis: time-to-event analysis using a competing risks analysis. Competing risks include death and graft failure requiring a second transplant. 
