---
title: "The Effects of Iron Overload on Hemptotic Stemcell Transplant (HSCT) Outcomes in Sickle Cell Patients"
author: "Nhat Nguyen \n PEDSnet Coordinating Center"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    page-layout: full
    embed-resources: true
    smooth-scroll: true
    fig-cap: TRUE
    df-print: paged
    toc: TRUE
    toc-depth: 5
    toc-collapsed: TRUE
    toc-float: TRUE
    toc-location: left
    number-sections: TRUE
    number-depth: 3
    output-file: analysis
    output-ext: "html"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 6, fig.align = "center")
# Default taken from R notebook behavior: when knitting, wd will always be location of notebook
base_dir <- '..'
Sys.setenv(PEDSNET_SKIP_REQUEST_EXECUTION=1)
tryCatch(source('../site/run.R'),
         error = function (e) message(e)) # May not be able to make db connection

# Set to "local" to read data from ../results, or another value to read from db
data_source <- "not_local" #if_else(config('execution_mode') == 'distribution', 'local', 'not_local')

require(tidyr)
require(knitr)
require(kableExtra)
require(stringr)
require(tibble)
require(ggplot2)
require(table1)
```

```{r funtioncs, include=FALSE}
get_codesets <- function(codeset_name) {
  read_csv(paste0('../specs/', codeset_name, '.csv'), col_types = c('c', 'c', 'c', 'c'), show_col_types = FALSE) %>%
  mutate(concept_id = as.character((concept_id)), concept_code = as.character((concept_code))) %>% collect()
}

get_results <- function(tbl_name) {
  if (data_source == 'local') {
    rslt <- read.csv(paste0('../local/', tbl_name, '.csv'))
  }
  else {
    rslt <- results_tbl(tbl_name) %>% collect()
  }
  rslt
}

prettify_kable <- function(data) {
  rslt <- data %>% kable(digits = 4, format.args = list(big.mark = ','))
  if (knitr::is_html_output()) rslt <- rslt %>%
      kable_styling(bootstrap_options = c("striped", "hover")) %>%
      column_spec(1, bold = T, border_right = T)
  rslt
}

```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
tribble( ~Item, ~Value,
         'Title', config('title'),
         'PI', config('requester_name'),
         'CDM Version', as.character(config('cdm_version'))) %>%
  prettify_kable()

```
```{r data, include= FALSE}
cohort <- results_tbl("study_cohorts") %>%
                    pivot_longer(cols = aim_2a_2:aim_3_2, names_to = "aim", values_to = "which_aim") %>% 
                    filter(which_aim == TRUE) %>% 
                    mutate(transplant_type = case_when(grepl("allogeneic", transplant_concept_name, ignore.case = TRUE) ~ "allogeneic",
                                              grepl("autologous", transplant_concept_name, ignore.case = TRUE) ~ "autologous",
                                              TRUE ~ "unknown")) %>% 
                    select(person_id, aim, site, age_at_ce , gender, scd_concept_id, scd_concept_name, scd_type, transplant_concept_id, transplant_concept_name, transplant_type, transplant_date)
cohort_ct <- cohort %>% distinct_ct()
```

**Confidentiality:** Please treat the contents of this report as confidential, and do not redistrubute beyond PEDSnet staff and the requester.

# Cohort Definition

-   **Time Anchors:** The specific details of cohort formation are described below. Note that for dates, we use notation to indicate time, where 0 refers to the day that a patient enters the cohort, a positive integer refers to the number of days after the cohort entrance that the observation period lasts, a negative interger refers to the number of days before the cohort entrance that the observation period lasts, oo refers to the end of the study period, and -oo refers to the beginning of the study period.

-   **Study period:** 1/1/2010 – 6/30/2023, specifically:
    - Aim 2a: 1/1/2010 – 6/30/2023
    - Aim 2b & 3: 1/1/2010 – 12/31/2022

-   **Cohort entrance date:** Date of first transplant

-   **Follow-up window:** 
    - Interval during which outcome(s) of interest may be included in the analysis: [-365, 00].
    - Exposures of interest (LIC, ferritin) for Aims 2a can be observed from -365 days to 0.
    - Exposures of interest (LIC, ferritin) for Aims 2b and 3 can be observed from day +90 to +365.
    - Exposures of interest (deferasirox or deferoxamine prescription, phlebotomy procedure) for Aim 3 can be observed from day +90 to 00. 
    - Outcomes of interest for Aim 2a can be observed from day 0 to 00 
    - Outcomes of interest for Aim 2b can be observed from date of post-transplant LIC or ferritin to 00. 
    - Outcomes of interest for Aim 3 can be observed from date of prescription of deferasirox or deferoxamine to 00, or date of first phlebotomy to 00

-   **Inclusion/exclusion criteria:** 
    - Aim 1: Pre-HSCT iron overload: Exclude patients who do not have either a pre-HSCT ferritin or LIC. Post-HSCT iron overload: exclude patients who do not have an LIC or ferritin 6-12 months after HSCT. 
    - Aim 2a: Include patients whose first transplant between 1/1/2010 – 6/30/2023. Exclude patients who do not have either a pre-HSCT ferritin or LIC. 
    - Aim 2b and 3: Include patients whose first transplant between 1/1/2010 – 12/1/2022. Exclude patients who do not have an LIC or ferritin 6-12 months after HSCT. 
    - Aim 3: Exclude patients who did not require post-transplant chelation or phlebotomy.

# Conditions and Procedure Codesets:
-   A compehensive code search was performed for all the conditions, procedures, and measurements specified in the DAP. The list of codesets could be found [here](https://chop365.sharepoint.com/teams/RSCH-ACRC/Shared%20Documents/Forms/AllItems.aspx?id=%2Fteams%2FRSCH%2DACRC%2FShared%20Documents%2FPEDSnet%2FPEDSnet%20Studies%2FActive%20Studies%2FGibson%5FIron%20Overload%2FQueries%2FCodesets%5F2024Feb27&viewid=fbc819dd%2Dfbe7%2D4805%2Da6fc%2D03eda92856f5).


# Result Summary

- <span style="background-color: #FFFF00"> This version of the report only include patients with verifiable transplant dates</span>. Additional patients could be added as chart review data come in. 

- The cohort comprises `r format(cohort_ct, big.mark = ",")` patients from `r summarise(cohort, ct = n_distinct(site)) %>% pull(ct)` PEDSnet sites for all 3 Aims.
    - Chart review data will include `r cohort %>% filter(aim != "aim_3_1", aim != "aim_3_2") %>% distinct_ct()` patients from aim 2a and 2b of this cohort and an additional n = 16 patients whose transplant dates and number of transplants were not verifiable from EHR data. 

## PEDSnet site attrition
<!-- need to filter out invalid measurements to aim size matches below --> 
```{r site_strat}
cohort %>% group_by(site, aim) %>% 
  summarize(no_persons = n_distinct(person_id)) %>% ungroup() %>% collect() %>%
    # mutate(cohort_size = if_else(cohort_size < 11L, "<11", as.character(cohort_size))) %>%
  pivot_wider(names_from = aim, values_from = no_persons) %>% 
  mutate(site_total = rowSums(select(., contains("aim")), na.rm = TRUE)) %>% 
  prettify_kable()
```

```{r attrition, echo=FALSE}
  # cohort_2a_1 %>% mutate(cohort = "Aim 2a.1") %>% select(person_id, cohort) %>%
  #               bind_rows(cohort_2a_2 %>% mutate(cohort = "Aim 2a.2") %>% select(person_id, cohort)) %>% 
  #               bind_rows(cohort_2b_1 %>% mutate(cohort = "Aim 2b.1") %>% select(person_id, cohort)) %>% 
  #               bind_rows(cohort_2b_2 %>% mutate(cohort = "Aim 2b.2") %>% select(person_id, cohort)) %>% 
  #               bind_rows(cohort_3_1 %>% mutate(cohort = "Aim 3.1") %>% select(person_id, cohort)) %>% 
  #               bind_rows(cohort_3_2 %>% mutate(cohort = "Aim 3.2") %>% select(person_id, cohort)) %>% 
  #               group_by(cohort) %>% 
  #               summarise(persons = n_distinct(person_id)) %>% 
  #               mutate(persons = if_else(persons<11, "<11", as.character(persons))) %>%
  #               prettify_kable()




# DBI doesn't have a good way to interrogate tables in a specified schema,
# so you have to list the step logs you want to summarise (e.g. DCC + remote
# sites).
# step_log <- list(config('qry_site')) %>%
#   map(.f = function(x) {
#     get_results(paste0(x, '_steps'))
#   }) %>%
#   reduce(.f = dplyr::union)
# 
# step_log %>%
#   arrange(stamp) %>%
#   select(cohort, persons) %>%
#   mutate(persons = if_else(persons > 0L & persons < 11L,
#                            as.integer(NA), as.integer(persons))) %>%
#   mutate(percent_prior = round(100 * persons /
#                                  coalesce(lag(persons), persons), 2)) %>%
#   prettify_kable()
```

## Demographics

```{r demog, echo=FALSE}

# cohort %>% group_by(gender) %>% summarise(persons = n_distinct(person_id)) %>% 
#             mutate(perc = paste0(round(persons/ct*100, 2), " %")) %>% prettify_kable()

cohort <- cohort %>% collect() %>% mutate(age_group = case_when(age_at_ce < 5 ~ "[0, 5)",
                                            age_at_ce < 10 ~ "[5, 10)",
                                            age_at_ce < 15 ~ "[10, 15)",
                                            age_at_ce < 20 ~ "[15, 20)",
                                            age_at_ce < 40 ~ "[20, 40)",
                                            TRUE ~ "40+"),
                            age_group = factor(age_group, levels = c("[0, 5)","[5, 10)","[10, 15)",
                                                                  "[15, 20)", "[20, 40)", "40+")),
                            scd_type = case_when(scd_type == "aa" ~ "aplastic anemia", 
                                                scd_type == "bta" ~ "beta thalassemia major", 
                                                scd_type == "dba" ~ "Diamond-Blackfan anemia",
                                                scd_type == "scd" ~ "sickle cell disease",
                                                TRUE ~ NA)) 
label(cohort$age_group) = "Age at first transplant"
label(cohort$gender) = "Gender"
label(cohort$transplant_type) = "Type of transplant"
label(cohort$scd_type) = "Type of scd"

table1(~gender + age_group + transplant_type + scd_type, data = cohort %>% distinct(person_id, transplant_date, .keep_all = TRUE))

# c_names <- vocabulary_tbl('concept') %>%
#   filter(domain_id %in% c('Gender', 'Race', 'Ethnicity')) %>%
#   select(concept_id, concept_name) %>% collect()
# 
# demog <- cdm_tbl('person') %>%
#   semi_join(distinct(cohort, person_id),
#             by = 'person_id', copy = TRUE) %>% collect() %>%
#   inner_join(select(c_names, concept_id, sex = concept_name),
#              by = c('gender_concept_id' = 'concept_id')) %>%
#   inner_join(select(c_names, concept_id, race = concept_name),
#              by = c('race_concept_id' = 'concept_id')) %>%
#   inner_join(select(c_names, concept_id, hispanic = concept_name),
#              by = c('ethnicity_concept_id' = 'concept_id')) %>%
#   select(person_id, sex, race, hispanic)
# 
# h_cts <- demog %>% group_by(sex, hispanic) %>% summarise(persons = n()) %>%
#   mutate(persons = if_else(persons > 0L & persons < 11L,
#                            as.integer(NA), as.integer(persons))) %>%
#   pivot_wider(names_from = sex, values_from = persons) %>% rename(group = hispanic)
# r_cts <- demog %>% group_by(sex, race) %>% summarise(persons = n()) %>%
#   mutate(persons = if_else(persons > 0L & persons < 11L,
#                            as.integer(NA), as.integer(persons))) %>%
#   pivot_wider(names_from = sex, values_from = persons) %>% rename(group = race)
# 
# prettify_kable(r_cts)
# prettify_kable(h_cts)

```

# Data analysis plan:
- This is a retrospective cohort study comparing outcomes between patients with pre/post-HSCT iron overload to patients without pre/post-HSCT iron overload. Analyze the impact of pre/post-HSCT elevated ferritin and LIC on short-term hematologic and transplant-related outcome: 
  - overall survival
  - graft-versus-host disease (GvHD)
  - graft failure
  - incidence of veno-occlusive disease (VOD)
  - bacteremia
  - immune reconstitution
  - time to neutrophil and platelet engraftment. 

## Descriptive analysis: 
### Aim_2a: 
- Percentage of overall survival in iron-overloaded patients vs non-iron overloaded patients 
- Percentage of patients who experienced disease relapse in iron-overloaded patients vs non-iron overloaded patients 
- Percentage of patients with GVHD in iron-overloaded patients vs non-iron overloaded patients 
- Percentage of patients with VOD in iron-overloaded patients vs non-iron overloaded patients 
- Neutrophil engraftment times in iron-overloaded patients vs non-iron overloaded patients 
- Platelet engraftment times in iron-overloaded patients vs non-iron overloaded patients 

### Aim_2b: 
- Percentage of patients who had an episode of bacteremia in iron-overloaded patients vs non-iron overloaded patients 
- Time to immune reconstitution in iron-overloaded patients vs non-iron overloaded patients 

### Aim 3:  
- Percentage of patients requiring post-transplant IRT 
- Percentage of patients receiving IRT with chelation compared to phlebotomy 
- Rate of ferritin reduction for patients receiving IRT with chelation vs phlebotomy (ng/ml/month) 

## Statistical Analysis
- GRFS: Time-to-event analysis with Cox proportional hazards model, Kaplan-Meier curves with log-rank test 
- Overall survival: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test.  
- Disease relapse: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test.  
- GVHD: Time-to-event analysis using Cox proportional hazard model and Kaplan-Meier curve with log-rank test.  
- VOD: Multivariable logistic regression with stratification for busulfan conditioning, use interaction term for busulfan and iron score 
- Time to neutrophil and platelet engraftment: time-to-event analysis using a competing risks analysis. Competing risks include death or graft failure prior to engraftment. 
- Bacteremia within 1 year after HSCT: Time-to-event analysis using a competing risks analysis. Competing risk includes death. 
- Immune Reconstitution: Definition of immune reconstitution: CD3 count >500, CD4 count >200, CD8 count >200, IgM >25. 
- Statistical analysis: time-to-event analysis using a competing risks analysis. Competing risks include death and graft failure requiring a second transplant. 

# Aim 2_1:
## Aim 2a_2: Post-transplant ferritin status
```{r aim_2a_2}
source("../code/EDA.R")
aim_2a_2 <- cohort_covars %>% filter(aim == "aim_2a_2") %>%
  mutate(death = if_else(death == TRUE, "Death", "Alive"), 
        disease_relapse_days_since_ce = as.numeric(difftime(disease_relapse_date, transplant_date, units = "days")))
label(aim_2a_2$disease_relapse) = "Disease relapse"
label(aim_2a_2$disease_relapse_days_since_ce) = "No days since ce until disease relapse*"
label(aim_2a_2$death) = "Overall survival"  
label(aim_2a_2$death_days_since_ce) = "No days since ce until death"
label(aim_2a_2$death_days_since_last_transplant) = "No days since disease relapse until death*"
label(aim_2a_2$GVHD) = "Graft-versus-host disease (GvHD)"
label(aim_2a_2$VOD) = "Incidence of veno-occlusive disease (VOD)"
label(aim_2a_2$bacteremia) = "Bacteremia**"

table1(~death + death_days_since_ce + 
      disease_relapse + disease_relapse_days_since_ce + 
      death_days_since_last_transplant +
      GVHD + VOD + bacteremia| ferritin_level, 
      data = aim_2a_2,
      footnote = c("* Missing indicates patients without disease relapse/death", 
                  "** Missing indicates no culture test was found, No means a negative result was reported"))
```

## Aim 2b_2: Pre-transplant ferritin status
```{r aim_2b_2}
aim_2b_2 <- cohort_covars %>% filter(aim == "aim_2b_2") %>%
  mutate(death = if_else(death == TRUE, "Death", "Alive"), 
        disease_relapse_days_since_ce = as.numeric(difftime(disease_relapse_date, transplant_date, units = "days")))
label(aim_2b_2$disease_relapse) = "Disease relapse"
label(aim_2b_2$disease_relapse_days_since_ce) = "No days since ce until disease relapse*"
label(aim_2b_2$death) = "Overall survival"  
label(aim_2b_2$death_days_since_ce) = "No days since ce until death"
label(aim_2b_2$death_days_since_last_transplant) = "No days since disease relapse until death*"
label(aim_2b_2$GVHD) = "Graft-versus-host disease (GvHD)"
label(aim_2b_2$VOD) = "Incidence of veno-occlusive disease (VOD)"
label(aim_2b_2$bacteremia) = "Bacteremia**"

table1(~death + death_days_since_ce + 
      disease_relapse + disease_relapse_days_since_ce + 
      death_days_since_last_transplant +
      GVHD + VOD + bacteremia| ferritin_level, 
      data = aim_2b_2,
      footnote = c("* Missing indicates patients without disease relapse/death", 
                  "** Missing indicates no culture test was found, No means a negative result was reported"))
```
## Aim 3: Post-transplant iron reduction therapy (IRT)
```{r aim_3}
# extract unique patients
table1(~IRT| ferritin_level, 
      data = cohort_covars %>% distinct(person_id, transplant_date, .keep_all = TRUE))
```

# Overall statistics: 
## Counts stratified by transplant concept names: 
```{r transplant_id_sum}
results_tbl("study_cohorts") %>% group_by(transplant_concept_id, transplant_concept_name) %>%
        summarise(n = n_distinct(person_id)) %>% 
        mutate(transplant_concept_id = as.character(transplant_concept_id)) %>%
        prettify_kable()
```