---
title: "EDA: The Effects of Iron Overload on Hemptotic Stemcell Transplant (HSCT) Outcomes in Sickle Cell Patients"
author: "Nhat Nguyen \n PEDSnet Coordinating Center"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    page-layout: full
    embed-resources: true
    smooth-scroll: true
    fig-cap: TRUE
    df-print: paged
    toc: TRUE
    toc-depth: 5
    toc-collapsed: TRUE
    toc-float: TRUE
    toc-location: left
    number-sections: TRUE
    number-depth: 3
    output-file: Gibson_sickle_cell_EDA
    output-ext: "html"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 6, fig.align = "center")
# Default taken from R notebook behavior: when knitting, wd will always be location of notebook
base_dir <- '..'
Sys.setenv(PEDSNET_SKIP_REQUEST_EXECUTION=1)
tryCatch(source('../site/run.R'),
         error = function (e) message(e)) # May not be able to make db connection

# Set to "local" to read data from ../results, or another value to read from db
data_source <- "not_local" #if_else(config('execution_mode') == 'distribution', 'local', 'not_local')

require(tidyr)
require(knitr)
require(kableExtra)
require(stringr)
require(tibble)
require(ggplot2)
require(table1)
require(tidyverse)
```

```{r functions, include=FALSE}
get_codesets <- function(codeset_name) {
  read_csv(paste0('../specs/', codeset_name, '.csv'), col_types = c('c', 'c', 'c', 'c'), show_col_types = FALSE) %>%
  mutate(concept_id = as.character((concept_id)), concept_code = as.character((concept_code))) %>% collect()
}

get_results <- function(tbl_name) {
  if (data_source == 'local') {
    rslt <- read.csv(paste0('../local/', tbl_name, '.csv'))
  }
  else {
    rslt <- results_tbl(tbl_name) %>% collect()
  }
  rslt
}

prettify_kable <- function(data) {
  rslt <- data %>% kable(digits = 4, format.args = list(big.mark = ','))
  if (knitr::is_html_output()) rslt <- rslt %>%
      kable_styling(bootstrap_options = c("striped", "hover")) %>%
      column_spec(1, bold = T, border_right = T)
  rslt
}

my.render.cont <- function(x) {
    with(stats.apply.rounding(stats.default(x), digits=2), c("",
        "Median [Min, Max]"=sprintf("%s [%s, %s])", MEDIAN, MIN, MAX)))
}
```

```{r report_recap, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
tribble( ~Item, ~Value,
         'Title', config('title'),
         'PI', config('requester_name'),
         'CDM Version', as.character(config('cdm_version'))) %>%
  prettify_kable()

```
```{r data, include= FALSE}
cohort <- results_tbl("study_cohorts") %>%
                    pivot_longer(cols = aim_2a_2:aim_3_2, names_to = "aim", values_to = "which_aim") %>% 
                    filter(which_aim == TRUE) %>% 
                    collect() %>%
                    mutate(age_at_ce = round(as.numeric(difftime(transplant_date, birth_date, units = "days"))/365.25)) %>%
                    # mutate(transplant_type = case_when(grepl("allogeneic", transplant_concept_name, ignore.case = TRUE) ~ "allogeneic",
                    #                           grepl("autologous", transplant_concept_name, ignore.case = TRUE) ~ "autologous",
                    #                           TRUE ~ "unknown")) %>% 
                    select(person_id, aim, site, age_at_ce , gender, scd_concept_name, scd_type, transplant_concept_name, transplant_type, transplant_date)
cohort_ct <- cohort %>% distinct_ct()
```

**Confidentiality:** Please treat the contents of this report as confidential, and do not redistrubute beyond PEDSnet staff and the requester.

# Cohort Definition

-   **Time Anchors:** The specific details of cohort formation are described below. Note that for dates, we use notation to indicate time, where 0 refers to the day that a patient enters the cohort, a positive integer refers to the number of days after the cohort entrance that the observation period lasts, a negative interger refers to the number of days before the cohort entrance that the observation period lasts, oo refers to the end of the study period, and -oo refers to the beginning of the study period.

-   **Study period:** 1/1/2010 – 6/30/2023, specifically:
    - Aim 2a: 1/1/2010 – 6/30/2023
    - Aim 2b & 3: 1/1/2010 – 12/31/2022

-   **Cohort entrance date:** Date of first transplant

-   **Follow-up window:** 
    - Interval during which outcome(s) of interest may be included in the analysis: [-365, 00].
    - Exposures of interest (LIC, ferritin) for Aims 2a can be observed from -365 days to 0.
    - Exposures of interest (LIC, ferritin) for Aims 2b and 3 can be observed from day +90 to +365.
    - Exposures of interest (deferasirox or deferoxamine prescription, phlebotomy procedure) for Aim 3 can be observed from day +90 to 00. 
    - Outcomes of interest for Aim 2a can be observed from day 0 to 00 
    - Outcomes of interest for Aim 2b can be observed from date of post-transplant LIC or ferritin to 00. 
    - Outcomes of interest for Aim 3 can be observed from date of prescription of deferasirox or deferoxamine to 00, or date of first phlebotomy to 00

-   **Inclusion/exclusion criteria:** 
    - Aim 1: Pre-HSCT iron overload: Exclude patients who do not have either a pre-HSCT ferritin or LIC. Post-HSCT iron overload: exclude patients who do not have an LIC or ferritin 6-12 months after HSCT. 
    - Aim 2a: Include patients whose first transplant between 1/1/2010 – 6/30/2023. Exclude patients who do not have either a pre-HSCT ferritin or LIC. 
    - Aim 2b and 3: Include patients whose first transplant between 1/1/2010 – 12/1/2022. Exclude patients who do not have an LIC or ferritin 6-12 months after HSCT. 
    - Aim 3: Exclude patients who did not require post-transplant chelation or phlebotomy.

# Variable definition:
- **LIC level**:
  - Low: <5 mg Fe/g dry wt
  - Moderate: [5, 8] mg Fe/g dry wt
  - High: >8 mg Fe/g dry wt
  - Binary classification: Low: <8 mg Fe/g dry wt and High otherwise
- **Ferritin level**:
  - Low: <1000 ng/mL 
  - Moderate: 1000-2500 ng/mL 
  - High: >=2500 ng/mL 
- **Immune reconstitution**: 
  - CD3 count >500
  - CD4 count >200 
  - CD8 count >200
  - IgM >25. 
- **VOD**: 
  - **(1)** (Prescription of defibrotide after transplant date AND Abdominal ultrasound within 2 days) within 90 days of transplant. 
  - ~~**(2)** (Abdominal ultrasound AND total bilirubin >2 mg/dL within 3 days) within 90 days of transplant~~ --- old definition, no longer used 
  OR 
  - **(3)** (Abdominal ultrasound AND total bilirubin >2 mg/dL within 2 days AND significant weight gain defined as: weight gain on 3 consecutive measurements within 10 days of ultrasound OR weight gain >5% of admission weight within 10 days of ultrasound) within 90 days of transplant
  OR
  - **(4)** Peritoneal drain placement procedure during transplant admission
(this should show up as a procedure code during admission) --- new criterion 
  OR 
  - **(5)** Veno-occlusive disease (VOD) or sinusoidal obstruction syndrome (SOS) on patient problem list --- new criterion
- **Neutrophil Engraftment**: The first of 3 successive days with an absolute neutrophil count (ANC) of 500/uL after post-transplantation nadi (Nadir usually occurs within 7-14 days after date of transplant) 
- **Platelet engraftment**: The first of 3 consecutive days with a platelet count of 20,000/uL or higher in the absence of platelet transfusion for 7 consecutive days. 
- **Primary graft failure**: 
  - Must have not met criteria for neutrophil engraftment or platelet engraftment previously
  - ANC <500/uL by day +28 after transplant AND hemoglobin < 8g/dL AND platelets < 20,000/uL 
- **Secondary graft failure**: 
  - Must have NOT met the requirement for primary graft failure 
  - Requirement for red blood cell or platelet transfusions, or need for G-CSF medication (growth colony stimulating factor --- normally given by injections) AFTER meeting criteria for neutrophil or platelet engraftment previously 

# Conditions and Procedure Codesets:
-   A compehensive code search was performed for all the conditions, procedures, and measurements specified in the DAP. The list of codesets could be found [here](https://chop365.sharepoint.com/teams/RSCH-ACRC/Shared%20Documents/Forms/AllItems.aspx?id=%2Fteams%2FRSCH%2DACRC%2FShared%20Documents%2FPEDSnet%2FPEDSnet%20Studies%2FActive%20Studies%2FGibson%5FIron%20Overload%2FQueries%2FCodesets%5F2024Feb27&viewid=fbc819dd%2Dfbe7%2D4805%2Da6fc%2D03eda92856f5).

# Transplant types and donor relationship
- Transplant types and donor relationship are combined into a single variable
  - Category 1: Matched related donor (related, 10/10 or 8/8)
  - Category 2: Matched unrelated donor (unrelated, 10/10 or 8/8)
  - Category 3: Mismatched unrelated donor (includes 8/10, 9/10, 7/8) 
  - Category 4: Haploidentical 
  - Category 5: Mismatched related donor (includes 8/10, 9/10, 7/8) 
  - Category 6: Cord (related or unrelated)
  - Category 7: Autologous (no donor)

# Result Summary

- <span style="background-color: #FFFF00"> For patients without chart review data, we use date of first transplant from EHR as cohort entry (CE)</span>. 

- The provisional cohort comprises `r format(cohort_ct, big.mark = ",")` patients from `r summarise(cohort, ct = n_distinct(site)) %>% pull(ct)` PEDSnet sites for all 3 Aims.
    - Chart review data include `r cohort %>% filter(aim != "aim_3_1", aim != "aim_3_2") %>% distinct_ct()` patients from aim 2a and 2b of this cohort and an additional n = 16 patients whose transplant dates and number of transplants were not verifiable from EHR data. 
    - The columns are not mutally exclusive as patients could be eligible for multiple aims. 

# Chart review 
## Chart review progress
```{r cohort_covars_run, echo= FALSE, message=FALSE, results='hide'}
cohort_covars <- results_tbl("analytics_dataset") %>% collect() 
```

```{r cr_progress, echo= FALSE, message=FALSE, results='hide'}
cr_data_raw <- read.csv("../redcap/IronOverloadChartRev_DATA_2024-11-08_1206.csv") %>% 
            # filter out old patient ids
           filter(!grepl("^[A-Za-z]", record_id)) %>% rename(site = redcap_data_access_group)
total_cr_ct <- cr_data_raw %>% distinct_ct("record_id")

# cleaned data
# cr_data <- get_redcap_data(redcap_filename = "../redcap/IronOverloadChartRev_DATA_2024-11-08_1206.csv") %>%
#             mutate(second_transplant = if_else(!is.na(second_transplant_date), TRUE, FALSE))
cr_data <- results_tbl("cr_data") %>% collect() %>%
            mutate(second_transplant = if_else(!is.na(second_transplant_date), TRUE, FALSE))
completed_cr_ct <- cr_data %>% distinct_ct("record_id")
true_positives_cr_ct <- cr_data %>% filter(eligibility == "Yes") %>% distinct_ct("record_id")
false_positives_cr_ct <- cr_data %>% filter(eligibility == "No") %>% distinct_ct("record_id")

# among chart reviews patients, how many had leukemia
leukemia_dx_px <- results_tbl(name = "leukemia_dx_px") %>% collect()
cr_data_w_leukemia <- cr_data %>% distinct(record_id, eligibility) %>%
                        left_join(results_tbl("study_cohorts")%>% select(person_id, record_id) %>% collect(), by = c("record_id")) %>%
                        inner_join(leukemia_dx_px, by = "person_id") %>%
                        distinct(record_id, eligibility, person_id) 

false_positives_cr_w_leukemia_ct <- cr_data_w_leukemia %>% filter(eligibility == "No") %>% 
                        distinct_ct("record_id")
true_positives_cr_w_leukemia_ct <- cr_data_w_leukemia %>% filter(eligibility == "Yes") %>% 
                        distinct_ct("record_id")
true_positives_cr_w_leukemia_prior_ct <- cohort_covars %>% filter(eligibility == "Yes") %>%
                        filter(leukemia_prior) %>%
                        distinct_ct("record_id")
# completed_cr_w_leukemia_ct <- r_data_w_leukemia %>% distinct_ct("record_id")
cr_w_leukemia_ct <- cr_data_raw %>% distinct(record_id) %>% left_join(results_tbl("study_cohorts")%>% select(person_id, record_id) %>% collect(), by = c("record_id")) %>%
                        inner_join(leukemia_dx_px, by = "person_id") %>%
                        distinct_ct("record_id")

# number of true positives with matched transplant dates
true_positves_matched_dates_ct <- cohort_covars %>% filter(transplant_date_consistency ==1) %>% filter(eligibility == "Yes") %>% distinct_ct()                       
```

- Total number of chart reviews: `r total_cr_ct`
  - Number of patients with leukemia in this cohort: `r cr_w_leukemia_ct`
- Number of completed chart reviews: `r completed_cr_ct` (`r round(completed_cr_ct/total_cr_ct*100)`%)
- Number of false positives: `r false_positives_cr_ct` (`r round(false_positives_cr_ct/total_cr_ct*100)`%)
  - <span style="background-color: #FFFF00">Number of false positives with leukemia: `r false_positives_cr_w_leukemia_ct` (`r round(false_positives_cr_w_leukemia_ct/false_positives_cr_ct*100)`% of total false positives) </span>
- Number of true positives: `r true_positives_cr_ct` (`r round(true_positives_cr_ct/total_cr_ct*100)`%)
  - <span style="background-color: #FFFF00"> Number of true positives with leukemia: `r true_positives_cr_w_leukemia_ct` (`r round(true_positives_cr_w_leukemia_ct/true_positives_cr_ct*100)`% of total true positives). Among these patients, n = `r true_positives_cr_w_leukemia_prior_ct` had leukemia diagnosed before transplants.</span>
  - Number of true positives with transplant dates correctly idenftified by PEDSnet EHR: n = `r true_positves_matched_dates_ct` (`r round(true_positves_matched_dates_ct/true_positives_cr_ct*100)`% of total true positives)

- <span style="background-color: #FFFF00">For cases when no. true positives and total valid charts dont match, due to the pending LIC/Ferritin values awaiting confirmation from Nora. In the final analytic dataset, the 2 columns should match </span>. 
```{r cr_progress_by_site}
# count false positives, true positives, num complted and num total for each site
cr_completion_by_site <- cr_data %>% group_by(site, eligibility) %>% 
      summarise(n = n_distinct(record_id)) %>% ungroup() %>%
      mutate(eligibility = if_else(eligibility == "No", "n_false_positives", "n_true_positives")) %>%
      # total counts per site
      union(cr_data_raw %>% group_by(site) %>% summarise(n = n_distinct(record_id)) %>% ungroup() %>% mutate(eligibility = "n_total")) %>%
      # total completed charts per site
      union(cr_data %>% filter(chart_completion) %>% group_by(site) %>% summarise(n = n_distinct(record_id)) %>% ungroup() %>% mutate(eligibility = "n_completed")) %>%
      pivot_wider(id_cols = site, names_from = eligibility, values_from = n) %>% 
      rename("no. false positives" = n_false_positives, 
             "no. true positives" = n_true_positives, 
             "total no.charts" =  n_total,
             "no. chart completed" =  n_completed)

# count number of patients for each aim per site
aim_ct_by_site <- cohort_covars %>% select(person_id, site, starts_with("aim")) %>%
                    pivot_longer(cols = aim_2a_2:aim_3_2, values_drop_na = TRUE, names_to = "aim", values_to = "val") %>% filter(val) %>%
                    group_by(site) %>% mutate(site_total = n_distinct(person_id)) %>% ungroup() %>% 
                    group_by(site, aim, site_total) %>% 
                    summarize(no_persons = n_distinct(person_id)) %>% ungroup() %>% collect() %>%
                    rename("no. valid per site" = site_total) %>%
                    pivot_wider(names_from = aim, values_from = no_persons) 
                    
cr_completion_by_site %>% left_join(aim_ct_by_site, by = "site") %>%                    
                  prettify_kable()
```

```{r cr_summary}
cr_t1 <- cr_data %>% select(record_id, chart_completion, eligibility, donor_relation,
                            match_status, graft_fail, second_transplant) %>% distinct()
label(cr_t1$chart_completion) = "Chart review completed"
label(cr_t1$eligibility) = "Cohort Eligibility"
label(cr_t1$donor_relation) = "Donor Relationship"
label(cr_t1$match_status) = "Match Status"
label(cr_t1$graft_fail) = "Graft failure*"
label(cr_t1$second_transplant) = "Had a second transplant"

table1(~eligibility + donor_relation + match_status + graft_fail +
      second_transplant, 
      data = cr_t1,
      caption = "Chart Review Summary (Only Completed Charts)", 
      footnote = "*Missing indicates false positive patients")
```

## Table 1: cohort characteristics
```{r cohort_covars}
cohort_covars <- cohort_covars %>% collect() %>% mutate(age_at_ce = as.numeric(difftime(transplant_date, birth_date, units = "days"))/365.25) %>% 
                            mutate(age_group = case_when(age_at_ce < 5 ~ "[0, 5)",
                                            age_at_ce < 10 ~ "[5, 10)",
                                            age_at_ce < 15 ~ "[10, 15)",
                                            age_at_ce < 20 ~ "[15, 20)",
                                            age_at_ce < 40 ~ "[20, 40)",
                                            TRUE ~ "40+"),
                            age_group = factor(age_group, levels = c("[0, 5)","[5, 10)","[10, 15)",
                                                                  "[15, 20)", "[20, 40)", "40+"))) %>%
                            mutate(ethnicity = if_else(ethnicity == "No information", "Unknown", ethnicity),
                            ethnicity = factor(ethnicity, levels = c("Hispanic or Latino", "Not Hispanic or Latino", "Other", "Unknown"))) %>%
          mutate(survival = if_else(survival, "Alive", "Death"), 
                survival_1yr = if_else(death_days_since_ce >= 365.25 | is.na(death_days_since_ce), "Alive", "Death"),
                survival_3yr = if_else(death_days_since_ce >= (365.25*3) | is.na(death_days_since_ce), "Alive", "Death"),
                survival_5yr = if_else(death_days_since_ce >= (365.25*5) | is.na(death_days_since_ce), "Alive", "Death"),
                disease_relapse_days_since_ce = as.numeric(difftime(graft_fail_date, transplant_date, units = "days")),
                bacteremia = if_else(is.na(bacteremia), "No", "Yes")) %>%
          mutate(last_in_person_visit_since_ce = case_when(last_in_person_visit_since_ce <= 180 ~ "0-6 months",
                                                  last_in_person_visit_since_ce <= 365.25 ~ "6-12 months",
                                                  last_in_person_visit_since_ce <= 365.25*3 ~ "1-3 yrs",
                                                  last_in_person_visit_since_ce <= 365.25*5 ~ "3-5 yrs",
                                                  last_in_person_visit_since_ce > 365.25*5 ~ "more than 5-yrs",
                                                  TRUE ~ NA),
         age_at_ce = round(as.numeric(difftime(transplant_date, birth_date, unit = "days"))/365.25),
         donor_relation = if_else(is.na(donor_relation), "Unknown", donor_relation),
         match_status = if_else(is.na(match_status), "Unknown", match_status),
         second_transplant = if_else(is.na(second_transplant_date), "No", "Yes"),
         gender = if_else(gender == "FEMALE", "Female", "Male"),
         scd_type = recode(scd_type, "aa" = "aplastic anemia", "dba" = "Diamond-Blackfan anemia", "bta" = "beta thalassemia major", "scd" = "sickle cell disease")) %>%
        mutate(CD3_reconstitution = case_when(CD3_reconstitute_3mon ~ "0-3 months",
                                        CD3_reconstitute_6mon ~ "3-6 months",
                                        CD3_reconstitute_9mon ~ "6-9 months",
                                        CD3_reconstitute_12mon ~ "9-12 months", TRUE ~ "Never/after 12 months"),
         CD4_reconstitution = case_when(CD4_reconstitute_3mon ~ "0-3 months",
                                        CD4_reconstitute_6mon ~ "3-6 months",
                                        CD4_reconstitute_9mon ~ "6-9 months",
                                        CD4_reconstitute_12mon ~ "9-12 months", TRUE ~ "Never/after 12 months"),
         CD8_reconstitution = case_when(CD8_reconstitute_3mon ~ "0-3 months",
                                        CD8_reconstitute_6mon ~ "3-6 months",
                                        CD8_reconstitute_9mon ~ "6-9 months",
                                        CD8_reconstitute_12mon ~ "9-12 months", TRUE ~ "Never/after 12 months"),
         IgM_reconstitution = case_when(IgM_reconstitute_3mon~ "0-3 months",
                                        IgM_reconstitute_6mon ~ "3-6 months",
                                        IgM_reconstitute_3mon ~ "6-9 months",
                                        IgM_reconstitute_12mon ~ "9-12 months", TRUE ~ "Never/after 12 months")) %>%
         mutate(across(ends_with("reconstitution"), ~factor(.x, levels = c("0-3 months", "3-6 months", "6-9 months", "9-12 months", "Never/after 12 months")))) %>%
         mutate(vod_subgroup = if_else(!vod | is.na(vod), "No VOD", scd_type)) %>% 
         mutate(vod_subgroup = factor(vod_subgroup, levels = c("aplastic anemia", "diamond-blackfan anemia", "beta thalassemia major", "sickle cell disease", "No VOD"))) %>%
         mutate(busulfan_and_vod = case_when(has_busulfan & vod ~ "busulfan & VOD",
                                            has_busulfan & !vod ~ "busulfan & No VOD", 
                                            TRUE ~ "Other categories")) %>%
         mutate(across(c("has_busulfan"), ~if_else(is.na(.x) | !(.x), "No", "Yes"))) %>%
         mutate(across(c("no_ferritin_pre", "no_ferritin_post", "no_LIC_pre", "no_LIC_post"), ~ifelse(.x >3, ">3", as.character(.x)))) %>%
         mutate(across(c("no_ferritin_pre", "no_ferritin_post", "no_LIC_pre", "no_LIC_post"), ~factor(.x, levels = c("1", "2", "3", ">3")))) %>%
        #  mutate(across(c("LIC_level", "ferritin_pre", "ferritin_post"), ~factor(.x, levels = c("low", "moderate", "high")))) %>%
        #  mutate(LIC_level_3_binary = factor(LIC_level_3_binary, levels = c("low", "high"))) %>%
         mutate(no_LIC_over_10_pre = ifelse(is.na(no_LIC_over_10_pre) | no_LIC_over_10_pre <1, "No", "Yes")) %>%
         mutate(busulfan_and_LIC_over_10 = ifelse(has_busulfan == "Yes" & no_LIC_over_10_pre == "Yes", "Yes", "No")) %>%
         mutate(busulfan_and_VOD_and_LIC_over_10 = ifelse(has_busulfan == "Yes" & no_LIC_over_10_pre == "Yes" & vod, "Yes", "No"))

label(cohort_covars$disease_relapse) = "Disease relapse/Graft failure"
label(cohort_covars$age_at_ce) = "Age at transplant"
label(cohort_covars$transplant_type) = "Transplant type"
label(cohort_covars$gender) = "Gender"
label(cohort_covars$disease_relapse_days_since_ce) = "No days since CE until disease relapse*"
# label(cohort_covars$ferritin_days_pre) = "No days between CE and ferritin measurements"
label(cohort_covars$survival) = "Overall survival"
label(cohort_covars$survival_1yr) = "Overall survival at 1 yr after CE" 
label(cohort_covars$survival_3yr) = "Overall survival at 3 yrs after CE"  
label(cohort_covars$survival_5yr) = "Overall survival at 5 yrs after CE"  
label(cohort_covars$death_days_since_ce) = "No days since CE until death*"
# label(aim_2a_2$death_days_since_last_transplant) = "No days since disease relapse until death*"
label(cohort_covars$GVHD) = "Graft-versus-host disease (GvHD)"
label(cohort_covars$vod_subgroup) = "Incidence of veno-occlusive disease (VOD)"
label(cohort_covars$bacteremia) = "Bacteremia"
# label(cohort_covars$ferritin_pre) = "Pre-transplant Ferritin Level"
label(cohort_covars$second_transplant) = "Had a second transplant"
label(cohort_covars$donor_relation) = "Donor relationship to the patient"
label(cohort_covars$match_status) = "Match status"
label(cohort_covars$scd_type) = "Disease Type"
label(cohort_covars$last_visit_since_ce) = "Follow-up duration (yrs)"
# label(cohort_covars$LIC_level) = "Pre-transplant Ferritin Level"
# label(cohort_covars$LIC_days_since_transplant) = "No days between CE and LIC measurements"
label(cohort_covars$CD3_reconstitution) = "CD3 reconstition post-transplant"
label(cohort_covars$CD4_reconstitution) = "CD4 reconstition post-transplant"
label(cohort_covars$CD8_reconstitution) = "CD8 reconstition post-transplant"
label(cohort_covars$IgM_reconstitution) = "IgM reconstition post-transplant"
label(cohort_covars$conditioning_type) = "Type of confitioning agents"
label(cohort_covars$busulfan_and_vod) = "Bulsufan & VOD"
label(cohort_covars$has_busulfan) = "Received Bulsufan"
label(cohort_covars$no_LIC_over_10_pre) = "Has at least 1 LIC > 10"
label(cohort_covars$busulfan_and_LIC_over_10) = "Has at least 1 LIC > 10 & received busulfan"
label(cohort_covars$busulfan_and_VOD_and_LIC_over_10) = "Has at least 1 LIC > 10 & VOD & received busulfan"
label(cohort_covars$no_LIC_pre) = "No. LIC pre-transplant measuremnts"
label(cohort_covars$no_ferritin_pre) = "No. ferritin pre-transplant measuremnts"
label(cohort_covars$transplant_type_combined) = "Transplant & donor status"
```

```{r table1, echo=FALSE}
cohort_t1 <- cohort_covars %>% filter(aim_2a_2 | aim_2a_1) %>% mutate(aim = "aim_2a") %>%
                union(cohort_covars %>% filter(aim_2b_1 | aim_2b_2) %>% mutate(aim = "aim_2b")) %>%
                union(cohort_covars %>% filter(aim_3_1 | aim_3_2) %>% mutate(aim = "aim_3") )%>%
                union(cohort_covars %>% distinct(person_id, .keep_all = TRUE) %>% mutate(aim = "Total"))
label(cohort_t1$age_group) = "Age at first transplant"
label(cohort_t1$gender) = "Gender"
label(cohort_t1$transplant_type) = "Type of transplant"
label(cohort_t1$scd_type) = "Type of scd"

table1(~gender + age_group + ethnicity + transplant_type + scd_type + transplant_type_combined| aim, 
      data = cohort_t1,
      overall = FALSE, 
      caption = "Table 1: cohort characteristics",
      footnote = "*3 aims are not multually exclusive")

```

# Descriptive analysis: 
## Aim_2a: Pre-transplant ferritin/LIC status
- Percentage of overall survival in iron-overloaded patients vs non-iron overloaded patients 
- Percentage of patients who experienced disease relapse in iron-overloaded patients vs non-iron overloaded patients 
- Percentage of patients with GVHD in iron-overloaded patients vs non-iron overloaded patients 
- Percentage of patients with VOD in iron-overloaded patients vs non-iron overloaded patients 
- Neutrophil engraftment times in iron-overloaded patients vs non-iron overloaded patients 
- Platelet engraftment times in iron-overloaded patients vs non-iron overloaded patients 

### Aim_2a_1: Pre-transplant LIC status
- <span style="background-color: #FFFF00">This table only has patients with complete chart reviews AND verifiable LIC unit conversions </span>
- We can also look at the max LIC, the average of the 3 max values, or the average of the 3 most recent values. However, since many patients only had 1 or 2 LIC values before transplant. This makes little difference
```{r aim_2a_1_LIC_pre}
LIC_data <- results_tbl("cr_LIC_data") %>% 
                  select(record_id, transplant_date, LIC, LIC_type, LIC_date, LIC_days_since_transplant, LIC_level, 
                  LIC_3, LIC_days_since_transplant_3, LIC_3_level,
                  LIC_max, LIC_3_max, LIC_days_since_transplant_max, LIC_days_since_transplant_3_max,
                  LIC_max_level, LIC_3_max_level, LIC_level_3_binary, LIC_level_3_max_binary) %>%
                  collect()
cohort_covars_LIC <- cohort_covars %>% left_join(LIC_data, by = c("record_id", "transplant_date"))
aim_2a_1 <- cohort_covars_LIC %>% filter(!is.na(LIC_level),  #aim_2a_1, 
                                    LIC_type == "pre-transplant", chart_completion) %>% collect()

table1(~no_LIC_pre + scd_type + age_at_ce + gender + 
       disease_relapse + disease_relapse_days_since_ce + 
       survival + death_days_since_ce + 
       LIC_days_since_transplant + 
       survival_1yr + survival_3yr + survival_5yr + 
      # death_days_since_last_transplant +
      GVHD + vod_subgroup + bacteremia +
      conditioning_type + 
      has_busulfan + 
      busulfan_and_vod + no_LIC_over_10_pre +
      busulfan_and_LIC_over_10 + 
      busulfan_and_VOD_and_LIC_over_10 + 
      transplant_type_combined| LIC_level, 
      data = aim_2a_1,
      caption = "Pre-transplant LIC status (Only patients with completed chart reviews AND verifiable LIC unit conversions)",
      footnote = c("* Missing indicates patients without disease relapse/death"))
```

### Aim_2a_1: Pre-transplant LIC status with binary classification
- <span style="background-color: #FFFF00">This table only has patients with complete chart reviews AND verifiable LIC unit conversions </span>
- This analysis uses the average of the 3 most recent LIC before transplant. 
  - We can also look at the max LIC of the 3 most recent values, or just the most recent LIC values. However, since many patients only had 1 or 2 LIC values before transplant (see in table below). This makes little difference. 

```{r aim_2a_1_binary_LIC_pre}
aim_2a_1_binary <- cohort_covars_LIC %>% filter(!is.na(LIC_level_3_binary), 
                                    LIC_type == "pre-transplant", 
                                    chart_completion) %>% collect()

table1(~no_LIC_pre + scd_type + age_at_ce + gender + 
       disease_relapse + disease_relapse_days_since_ce + 
       survival + death_days_since_ce + 
       LIC_days_since_transplant + 
       survival_1yr + survival_3yr + survival_5yr + 
      # death_days_since_last_transplant +
      GVHD + vod_subgroup + bacteremia +
      conditioning_type + 
      has_busulfan + 
      busulfan_and_vod + no_LIC_over_10_pre +
      busulfan_and_LIC_over_10 + 
      busulfan_and_VOD_and_LIC_over_10 + 
      transplant_type_combined| LIC_level_3_binary, 
      data = aim_2a_1,
      caption = "Pre-transplant LIC status (Only patients with completed chart reviews AND verifiable LIC unit conversions)",
      footnote = c("* Missing indicates patients without disease relapse/death"))
```

### Aim_2a_2: Pre-transplant ferritin status

- <span style="background-color: #FFFF00">To-do: </span>
  - The fraction of GVHD looks low compared to Armand 2007 
  - Patients with very short follow up, did they die? 

```{r aim_2a_2_ferritin_pre}
ferritin_data <- results_tbl("cr_ferritin_data") %>%
                    select(-no_ferritin_pre, - no_ferritin_post) %>% collect()
cohort_covars_ferritin <- cohort_covars %>% left_join(ferritin_data, by = c("person_id", "transplant_date"))

aim_2a_2 <- cohort_covars_ferritin %>% 
                filter(chart_completion, !is.na(ferritin_pre)) 

table1(~no_ferritin_pre + scd_type + age_at_ce + gender + disease_relapse + 
       disease_relapse_days_since_ce + 
       ferritin_days_pre + 
       survival + survival_1yr + survival_3yr + survival_5yr + 
       death_days_since_ce + #death_days_since_last_transplant +
       transplant_type + 
       donor_relation + match_status + 
       second_transplant + scd_type + 
       round(last_visit_since_ce/365.25, 1) + 
       GVHD + vod_subgroup + bacteremia +
       conditioning_type + has_busulfan + 
       busulfan_and_vod + transplant_type_combined| ferritin_pre, 
       data = aim_2a_2, 
       caption = "Pre-transplant ferritin status",
       footnote = c("* Missing indicates patients without disease relapse/death"), 
       render.continuous=my.render.cont)
```

## Aim_2b: 
- Percentage of patients who had an episode of bacteremia in iron-overloaded patients vs non-iron overloaded patients 
- Time to immune reconstitution in iron-overloaded patients vs non-iron overloaded patients

### Aim_2b_1: Post-transplant LIC status 
- <span style="background-color: #FFFF00">This table only has patients with complete chart reviews AND verifiable LIC unit conversions </span
```{r aim_2b_1_LIC_post}
aim_2b_1 <- cohort_covars_LIC %>% filter(!is.na(LIC_level), LIC_type == "post-transplant", chart_completion) %>% collect() 

table1(~no_LIC_post + scd_type + age_at_ce + gender + 
       disease_relapse + disease_relapse_days_since_ce + death_days_since_ce + 
       LIC_days_since_transplant +
       survival + 
       survival_1yr + survival_3yr + survival_5yr + 
       CD3_reconstitution + CD4_reconstitution + CD8_reconstitution + IgM_reconstitution + 
       vod + GVHD + 
      # death_days_since_last_transplant +
      bacteremia + transplant_type_combined| LIC_level, 
      data = aim_2b_1,
      caption = "Post-transplant LIC status (Only patients with completed chart reviews AND verifiable LIC unit conversions)",
      footnote = c("* Missing indicates patients without disease relapse/death"))
```

### Aim_2b_2: Post-transplant ferritin status
- <span style="background-color: #FFFF00">This table only has patients with complete chart reviews </span>
```{r aim_2b_2_ferritin_post}
aim_2b_2 <- cohort_covars_ferritin %>% 
              filter(chart_completion, !is.na(ferritin_post)) %>% collect() 

table1(~no_ferritin_post + scd_type + age_at_ce + gender + 
       disease_relapse + 
       disease_relapse_days_since_ce + 
       ferritin_days_post + 
       survival + 
       death_days_since_ce + 
       survival_1yr + survival_3yr + survival_5yr + 
       CD3_reconstitution + CD4_reconstitution + CD8_reconstitution + IgM_reconstitution + 
       GVHD + vod_subgroup + 
      # death_days_since_last_transplant +
      bacteremia + transplant_type_combined| ferritin_post, 
      data = aim_2b_2,
      caption = "Post-transplant ferritin status",
      footnote = c("* Missing indicates patients without disease relapse/death"))
```

## Aim 3: Post-transplant iron reduction therapy (IRT) 
- Percentage of patients requiring post-transplant IRT 
- Percentage of patients receiving IRT with chelation compared to phlebotomy 
- Rate of ferritin reduction for patients receiving IRT with chelation vs phlebotomy (ng/ml/month) 

```{r aim_3}
# potential aim 3, n = 158
aim_3 <- results_tbl("study_cohorts") %>% filter(aim_3_2 | aim_3_1)
# exclude false positive patients from chart reviews, n = 139
aim_3 <- aim_3 %>% anti_join(results_tbl("cr_data") %>%
                    filter(eligibility == "No") %>% select(record_id), by = "record_id")
no_ferritin_aim3_ct <- aim_3 %>% distinct_ct()
# if the patient had chart reviews, use the transplant_date from chart reviews
aim_3 <- aim_3 %>% left_join(results_tbl("analytics_dataset") %>% filter(chart_completion, eligibility == "Yes") %>% 
                                  distinct(record_id, transplant_date) %>% mutate(is_cr = TRUE), by = "record_id") %>%
                    mutate(transplant_date.x = ifelse(is_cr, transplant_date.y, transplant_date.x)) %>%
                    rename(transplant_date = transplant_date.x) %>%
                    select(-transplant_date.y)

# get ferritin data, n = 139 patients with valid ferritin 
aim3 <- aim_3 %>%
      select(record_id, person_id, transplant_date) %>%
      left_join(results_tbl("covar_ferritin_mx"), by = "person_id") %>% 
      filter(transplant_date <= ferritin_date) %>% 
      filter(ferritin_date - transplant_date <= 600) %>%
      filter(!is.na(ferritin)) %>%
      group_by(person_id) %>% 
      mutate(n = n()) %>% filter(n >=1) %>% ungroup()
      
aim3_ct <- aim_3 %>% distinct_ct()

# find therapeutic phlebotomy, n = 43
phleb_px <- aim_3 %>%
      # filter(chart_completion, eligibility == "Yes") %>%
      select(record_id, person_id, transplant_date) %>%
      find_procedures(procedure_codeset_name = "therapeutic_phleb_px") %>% 
      filter(procedure_date >= transplant_date) %>%
      filter(procedure_date - transplant_date <= 365) %>% 
      group_by(person_id) %>%
      collect_new() %>%
      mutate(start_date = min(procedure_date, na.rm = TRUE), 
             end_date = max(procedure_date, na.rm = TRUE),
             duration = as.numeric(end_date - start_date)) %>% ungroup() %>%
      distinct(person_id, record_id, start_date, end_date, transplant_date, duration) %>% 
      mutate(route = NA) %>%
      mutate(IRT_type = "phlebotomy") 

# patients who got chelation, n = 20
dx_codeset<- load_codeset("deferoxamine_rx") %>% mutate(type = "deferoxamine") %>% #deferoxamine
                union(load_codeset("deferasirox_rx") %>% mutate(type = "deferasirox")) %>% #deferasirox
                compute_new(temp = TRUE, name = "drug_id")

# patients who got chelation, n = 20
chelation_rx <- aim_3 %>%
      # filter(chart_completion, eligibility == "Yes") %>%
      select(record_id, person_id, transplant_date) %>%
      find_drugs(dx_codeset) %>% 
      filter(drug_exposure_start_date >= transplant_date) %>% 
      filter(drug_exposure_start_date - transplant_date <= 365) %>% 
      distinct(person_id, drug_exposure_start_date, IRT_type = type, .keep_all = TRUE) 

# for intravenous drugs, assume a single one per day, n = 3
chelation_iv <- chelation_rx %>% filter(route_source_value %in% c("Intravenous", "IV", "Injection", "Subcutaneous")) %>%
      group_by(person_id, IRT_type) %>%
      collect_new() %>% 
      mutate(start_date = min(drug_exposure_start_date, na.rm = TRUE), 
             end_date = max(drug_exposure_start_date, na.rm = TRUE),
             duration = as.numeric(end_date - start_date)) %>% ungroup() %>%
      distinct(person_id, record_id, start_date, end_date, transplant_date, duration, .keep_all = TRUE) %>% 
      mutate(route = "IV") %>%
      select(person_id, record_id, start_date, end_date, transplant_date, duration, IRT_type, route)

# oral drugs, n = 6
chelation_oral <- chelation_rx %>% filter(!(route_source_value %in% c("Intravenous", "IV", "Injection", "Subcutaneous"))) %>%
      filter(!is.na(quantity)) %>%
      collect_new() %>%
      mutate(refills = if_else(is.na(refills), 0, refills)) %>%
      group_by(person_id, IRT_type) %>% 
      mutate(start_date = drug_exposure_start_date, # min(drug_exposure_start_date, na.rm = TRUE), 
             end_date = drug_exposure_start_date + days(quantity*(refills+1)),
             duration = quantity) %>% ungroup() %>%
      distinct(person_id, record_id, start_date, end_date, transplant_date, duration, quantity, refills, .keep_all = TRUE) %>% 
      select(person_id, record_id, start_date, end_date, transplant_date, duration, quantity, refills, IRT_type, route = route_source_value) %>% 
      arrange(person_id) 

# Only took the duration of the first prescription, some prescriptions have up to 11 refills 

# append the chelation and IRT patients
irt <- phleb_px %>% 
          union(chelation_iv) %>% mutate(quantity = NA, refills = NA) %>% 
          union(chelation_oral) 

```

- When doing the feasibility counts, the number of patients with either therapeutic phlebotomy or deferasirox or deferoxamine after transplant is `r no_ferritin_aim3_ct`
- Among these, `r aim3_ct` has valid ferritin measurements within a year
- A lot of patients with missing quanity, days_supply, refills in the dataset; thus, we could not calculate the treatment durations

### Patients with more than 1 IRT
- 6 patients had more than 1 treatment separated by more than 6 months. What should we do with these? 
```{r aim_3_more_than_1treatment, echo = FALSE, warning = FALSE}
# who got more than 1 treatment
irt_exclude <- irt %>% group_by(record_id) %>% 
      summarise(n = n()) %>% 
      filter(n> 1) %>% ungroup()
      
irt_exclude %>% 
      inner_join(irt, by = "record_id") %>% 
      arrange(record_id) %>%
      select(-person_id) %>% prettify_kable()

# there are patients who got more than 1 IRT treaments, we do have to double check
irt <- irt %>% group_by(record_id, IRT_type) %>%
          mutate(start_date = min(start_date, na.rm = TRUE)) %>%
          mutate(end_date = max(end_date, na.rm = TRUE))

```

```{r IRT_ferritin_TS_data, include = FALSE, echo = FALSE, warning = FALSE}
# exclude patients with more than 1 treatment for now: 
irt <- irt %>% anti_join(irt_exclude, by = "record_id")

ferritin_start <- irt %>% copy_to_new(df = ., name = "irt") %>%
      inner_join(results_tbl("covar_ferritin_mx") %>%
                  select(person_id, ferritin_date, ferritin), by = "person_id") %>% 
      filter(transplant_date <= ferritin_date) %>%
      filter(!is.na(ferritin)) %>%
      # filter(transplant_date - ferritin_date <= 365) %>%
      group_by(person_id) %>% 
      filter(start_date <= ferritin_date) %>%
      slice_max(start_date, n =1, with_ties = FALSE) %>% ungroup() %>%
      select(person_id, record_id, transplant_date, start_date, ferritin0_date = ferritin_date, ferritin0 = ferritin) 
      

ferritin_end <- irt %>% copy_to_new(df = ., name = "irt") %>%
      inner_join(results_tbl("covar_ferritin_mx") %>%
                  select(person_id, ferritin_date, ferritin), by = "person_id") %>% 
      filter(transplant_date <= ferritin_date) %>%
      filter(!is.na(ferritin)) %>%
      filter(end_date <= ferritin_date) %>%
      group_by(person_id) %>% 
      slice_min(end_date, n =1, with_ties = FALSE) %>% ungroup() %>%
      select(person_id, record_id, transplant_date, end_date, ferritinf_date = ferritin_date, ferritinf = ferritin)

ferritin_all <- irt %>% copy_to_new(df = ., name = "irt") %>% 
      inner_join(results_tbl("covar_ferritin_mx") %>%
                  select(person_id, ferritin_date, ferritin), by = "person_id") %>%
      filter(transplant_date <= ferritin_date) %>%
      filter(!is.na(ferritin)) %>%
      filter(transplant_date - ferritin_date <= 365) %>% collect_new()
```

```{r aim_3_table1, include = FALSE}
ferritin_0 <- ferritin_all %>% group_by(person_id, IRT_type) %>%
          slice_min(abs(start_date - ferritin_date), n = 1, with_ties = FALSE) %>% ungroup() %>%
          select(record_id, person_id, IRT_type, start_date, ferritin_date_0 = ferritin_date, ferritin_0 = ferritin)

ferritin_f <- ferritin_all %>% group_by(person_id, IRT_type) %>%
          slice_min(abs(end_date - ferritin_date), n = 1, with_ties = FALSE) %>% ungroup() %>%
          select(record_id, person_id, IRT_type, start_date, ferritin_date_f = ferritin_date, ferritin_f = ferritin)
ferritin_change <- ferritin_0 %>% left_join(ferritin_f, by = c("record_id", "person_id", "IRT_type", "start_date")) %>%
          mutate(ferritin_change_rate = (ferritin_f - ferritin_0)/as.numeric((ferritin_date_f - ferritin_date_0))) 

table1(~ferritin_change_rate| IRT_type, 
      data = ferritin_change,
      caption = "Post-transplant iron reduction therapy",
      overall = FALSE)
```

### Ferritin trajectories
```{r IRT_ferritin_TS, fig.width = 10, fig.height = 20, echo = FALSE, warning = FALSE}
print(ferritin_all %>% 
      mutate(ferritin_date = as.numeric(ferritin_date - transplant_date)) %>%
      mutate(start_date = as.numeric(start_date - transplant_date)) %>%
      mutate(end_date = as.numeric(end_date - transplant_date)) %>% 
      filter(duration > 0) %>%
      ggplot(aes(x = ferritin_date, y = ferritin, group = record_id)) + 
      geom_point() + 
      geom_line() +
      geom_vline(aes(xintercept = start_date), linetype = "dashed", color = "red") +
      geom_vline(aes(xintercept = end_date), linetype = "dashed", color = "red") +
      xlim(c(0, 500)) + 
      facet_wrap(~record_id, ncol = 3)) 
```

# Overall statistics: 
- 01_0064_cr, 01_0113 (cchmc) met all criteria but was excluded for some other reasons
- 08_0017 (stanford): Quantitative liver iron assessment is nondiagnostic due to respiratory motion.
- Most cases were excluded because patients had leukemia and received transplant not for the disaese of interest.

## False positive patients by site:
```{r false_positves_by_site, echo = FALSE, warning = FALSE}
false_pos <- results_tbl("false_positive_ids")

false_pos %>% distinct(person_id) %>% inner_join(results_tbl("scd_dx") %>% 
                        distinct(person_id, site), by = c("person_id")) %>% 
               group_by(site) %>% summarise(num = n()) %>% 
               ungroup() 

false_pos %>% copy_to_new(df=., name = "false_pos") %>%
      inner_join(results_tbl("leukemia_dx_px"), by = "person_id") %>% distinct_ct()
```

### Disease counts for false positive and true positive patients:
- Even though a few codes only generated false positive cases, it was just a single instance. Patients might have been ineligible for the cohort for other reasons. 

```{r false_positives_by_scd_name}
# false_positve vs true positive counts
results_tbl("scd_dx") %>% distinct(person_id, scd_concept_name) %>%
        left_join(false_pos %>% distinct(person_id) %>%
                        mutate(false_pos_ct = 1), by = "person_id") %>%  
        # group_by(scd_concept_name) %>%
        # summarise(false_pos_ct = sum(false_pos_ct, na.rm = TRUE)) %>% ungroup() %>% 
        collect() %>%
        left_join(cohort_covars %>% filter(chart_completion) %>% 
                  distinct(person_id) %>% 
                  mutate(true_pos_ct = 1), by = "person_id") %>%
        group_by(scd_concept_name) %>%
        summarise(true_pos_ct = sum(true_pos_ct, na.rm = TRUE),
                false_pos_ct = sum(false_pos_ct, na.rm = TRUE)) %>% ungroup() %>%
        filter(true_pos_ct > 0 | false_pos_ct > 0) %>%
        arrange(desc(false_pos_ct), scd_concept_name) %>%
        prettify_kable()
```

### False positive patients by transplant procedures
- The procedure codes were reviewed by Nora before data curation. 
```{r false_positives_by_transplant_code}
# false_pos %>% inner_join(results_tbl("no_multi_transplant_px") %>%
#                             distinct(transplant_concept_name, person_id), by = "person_id") %>%
#                             group_by(transplant_concept_name) %>%
#                             summarise(num = n_distinct(person_id)) %>% arrange(desc(num)) %>% prettify_kable()
                            
results_tbl("study_cohorts") %>% distinct(person_id, transplant_concept_name) %>%
        left_join(false_pos %>% distinct(person_id) %>%
                        mutate(false_pos_ct = 1), by = "person_id") %>%  
        # group_by(scd_concept_name) %>%
        # summarise(false_pos_ct = sum(false_pos_ct, na.rm = TRUE)) %>% ungroup() %>% 
        collect() %>%
        left_join(cohort_covars %>% filter(chart_completion) %>% 
                  distinct(person_id) %>% 
                  mutate(true_pos_ct = 1), by = "person_id") %>%
        group_by(transplant_concept_name) %>%
        summarise(true_pos_ct = sum(true_pos_ct, na.rm = TRUE),
                false_pos_ct = sum(false_pos_ct, na.rm = TRUE)) %>% ungroup() %>%
        filter(true_pos_ct > 0 | false_pos_ct > 0, !is.na(transplant_concept_name) ) %>%
        arrange(desc(false_pos_ct), transplant_concept_name) %>%
        prettify_kable()
```

## False negative patients at CHOP
- There are 14 false negative patients from CHOP for the following reason:
  - 1 patient is in our transplant cohort but not the disease cohort because we missed out 1 condition code for Beta thalassemia (concept_id = 4278669). 
  - 13 patients had the following procedure codes that were not included
  - We could include the false negative patients from CHOP and other site; however, this requires moving the study schema to a new version of the database. Need to check with Charlie for the optimal solutions. 
```{r false_negative_patients}
read.csv("../local/missing_px_codes.csv") %>%
  mutate(procedure_concept_id = as.character(procedure_concept_id)) %>% 
  prettify_kable()  
```
  
## True positives patients with more than one disease subgroup: 
```{r multi_disease_subgroup_ct_1}
# patients with more than 1 disease type
multi_subgroups <- results_tbl("scd_dx") %>% group_by(person_id) %>% 
                        summarise(subgroup_ct = n_distinct(scd_type)) %>% 
                        filter(subgroup_ct > 1) %>% ungroup() %>%
                        inner_join(cohort_covars %>% distinct(person_id, record_id, chart_completion) %>% copy_to_new(df=., name = "njh"), by = "person_id") 
multi_subgroups_resolved_ct <- read.csv("../local/multi_subgroup_resolved.csv") %>% distinct_ct()
```

- The true positives count only include patients in aim 2a & 2b. The overall cohort has `r cohort_covars %>% distinct_ct()` patients. 
- Only considered true positive patients in this table. 
  - `r multi_subgroups %>% filter(chart_completion) %>% distinct_ct("person_id")` true positive patients from chart review with more than 1 disease subgroup. 
  - `r multi_subgroups %>% distinct_ct("person_id")` true positive patients & patients from aim 3 with more than 1 disease subgroup. 
  - By filtering by problem lists, date of diagnosis, number of diagnosis, we were able to resolve `r multi_subgroups_resolved_ct` cases. 
```{r multi_disease_subgroup_ct_2}
# multi_subgroups %>% inner_join(results_tbl("scd_dx") , by = "person_id") %>% 
#                 distinct(record_id, scd_concept_name, scd_type) %>% 
#                 arrange(record_id, scd_concept_name) %>% prettify_kable()
```

## True positives patients with more than one disease diagnosis: 
```{r multi_disease_ct_1}
# patients with more than 1 disease type
multi_disease <- results_tbl("scd_dx") %>% group_by(person_id) %>% 
                        summarise(disease_ct = n_distinct(scd_concept_name)) %>% 
                        filter(disease_ct > 1) %>% ungroup() %>%
                        inner_join(cohort_covars %>% filter(chart_completion) %>% distinct(person_id, record_id) %>% copy_to_new(df=., name = "njh"), by = "person_id") 
```
- Only considered true positive patients in this table:
  -  `r multi_disease %>% distinct_ct("person_id")` patients with more than 1 disease diagnosis. 
```{r multi_disease_ct_2}
# multi_disease %>% inner_join(results_tbl("scd_dx") , by = "person_id") %>% 
#                 distinct(record_id, scd_concept_name, scd_type) %>% 
#                 arrange(record_id, scd_concept_name) %>% prettify_kable()
```

## VOD status:
- Refer to variable definition sections for the category code, the majority of VOD cases met the first definition. Introducing the other definition makes little impact. 
```{r vod_status}
vod <- results_tbl("covar_vod_dx") %>% 
        inner_join(results_tbl("analytics_dataset") %>%
                    filter(eligibility == "Yes", chart_completion) %>% 
                    select(person_id, transplant_date), by = "person_id") %>%
        mutate(vod_date := pmin(min_defi_abd_date, min_bili_abd_date, min_bili_abd_weight_date, na.rm = TRUE)) %>%
        collect() %>% filter(as.numeric(difftime(vod_date, transplant_date, units = "days")) <= 90 |
                                (vod_date >= transplant_date) | 
                                abs(as.numeric(difftime(abd_date, transplant_date, units = "days"))) <= 7) %>% 
        distinct(person_id, vod_cat_1, vod_cat_2, vod_cat_3, vod_cat_4, vod_cat_5, vod) 
vod %>% filter(!is.na(vod_cat_1)| !is.na(vod_cat_3)| !is.na(vod_cat_4)| !is.na(vod_cat_5)) %>%
group_by(vod_cat_1, vod_cat_2, vod_cat_3, vod_cat_4, vod_cat_5) %>% 
                 summarise(n_patients = n_distinct(person_id)) %>% ungroup() %>% 
                 mutate(across(starts_with("vod_cat"), ~if_else(!is.na(.x), "Yes", "No"))) %>% prettify_kable()
```

## match status, donor relation, and transplant type for aim_2a
```{r match_status}
results_tbl("analytics_dataset") %>%
        filter(aim_2a_1|aim_2a_2) %>%
        filter(eligibility == "Yes", chart_completion) %>%
        group_by(match_status, donor_relation, transplant_type) %>%
        summarise(n = n_distinct(person_id)) %>% 
        prettify_kable()
```
